### 2019-07-03 images/flow.tex
first commit

% Diagram of Android activity life cycle
% Author: Pavel Seda 
\documentclass[border=10pt]{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes, chains, arrows, shadows}

% Set up a few colours

\tikzset{%
  >={Latex[width=2mm,length=2mm]},
  % Specifications for style of nodes:
	base/.style = {draw, on chain, on grid, align=center, minimum height=4ex},
	Docs/.style = {base, fill=blue!30},
	screen/.style = {base, fill=orange!15},
	form/.style = {on chain, on grid, align=center, minimum height=4ex, fill=white},
    computer/.style = {base, fill=green!30},
    decision/.style={base, diamond, aspect=2.5, fill=red!30,text width=5em},
	process/.style = {base, minimum width=2.5cm, fill=orange!15,
                           font=\ttfamily},                       
    coord/.style={coordinate, on chain, on grid, node distance=6mm and 25mm},
	norm/.style={->, draw, black},  
	multidoc/.style={
		base,
		shape=tape,
		tape bend top=none,
		fill=white,
		double copy shadow,
		fill=blue!30
	},
}
\begin{document}    
% Drawing part, node distance is 1.5 cm and every node
% is prefilled with white background

\begin{tikzpicture}[
    >=triangle 60,              % Nice arrows; your taste may be different
start chain=going below,    % General flow is top-to-bottom
node distance=6mm and 50mm, % Global setup of box spacing
every join/.style={norm},   % Default linetype for connecting boxes
]
%    every node/.style={fill=white, font=\sffamily}, align=center]
  % Specification of nodes (position, etc.)
  \node (docs)  [multidoc]  {N possibly relevant documents};

  \node (sample)  [screen, join]  {Screen random documents};
  \node (model) [computer, join, left=of sample]  {Train model,\\ predict relevance};
  \node (screenrel)  [screen, below of =sample]  {Screen most \\ relevant documents};
  \node (stopscreen) [decision, join] {Irrelevant?};

  \node (screenr2) [screen, join]  {Screen random documents};

  \node (recall) [decision, join] {Recall $>$ threshold?};

	\node (sform) [form, right=of screenr2, xshift=-15mm] { Estimate $\hat{p}U$ };

  \node (rform) [form, right=of recall, xshift=-15mm] {$ 1 - \frac{\hat{p}U}{\hat{p}P} \frac{U}{P}$};

  \node (reldocs) [multidoc, below of=recall, xshift=20mm, yshift=-8mm]  {Relevant documents};
  \node (irreldocs) [multidoc, fill=blue!5, below of=recall, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
  %\node

  \draw [o->,norm] (model.south) |-  (screenrel.west);

  \node [coord, left=of stopscreen, xshift=-45mm] (c1)  {}; %\cmark{1} 
  \node [coord, above=of model, yshift=2em] (c2)  {}; %\cmark{1}

  \node [coord, left=of recall, xshift=-45mm] (c3)  {}; %\cmark{1}

 \path (stopscreen.west) to node [very near start, yshift=1em] {$n$} (c1);

 	\draw [o->,norm] (stopscreen.west) -- (c1) |-  (c2) -- (model.north);

  \path (recall.west) to node [very near start, yshift=1em] {$n$} (c3);

 %\draw [o->,norm] (recall.west) -- (c3) |-  (c2) -- (model.north);
 \draw [o->,norm] (recall.west) -- (c3) |-  (screenr2.west);

 \draw [o->,norm] (recall.south) -- (reldocs);
 \draw [o->,norm] (recall.south) -- (irreldocs);

  \end{tikzpicture}
\end{document}

### 2019-07-03 images/flow_basic.tex
first commit

--- 2019-07-03 images/flow.tex
@@
   \node (screenrel)  [screen, below of =sample]  {Screen most \\ relevant documents};
-  \node (stopscreen) [decision, join] {Irrelevant?};
+  %\node (stopscreen) [decision, join] {Irrelevant?};
 
-  \node (screenr2) [screen, join]  {Screen random documents};
+  %\node (screenr2) [screen, join]  {Screen random documents};
 
-  \node (recall) [decision, join] {Recall $>$ threshold?};
+  %\node (recall) [decision, join] {Recall $>$ threshold?};
 
-	\node (sform) [form, right=of screenr2, xshift=-15mm] { Estimate $\hat{p}U$ };
+	%\node (sform) [form, right=of screenr2, xshift=-15mm] { Estimate $pU$ };
 
-  \node (rform) [form, right=of recall, xshift=-15mm] {$ 1 - \frac{\hat{p}U}{\hat{p}P} \frac{U}{P}$};
+  %\node (rform) [form, right=of recall, xshift=-15mm] {$ 1 - \frac{pU}{pP} \frac{U}{P}$};
 
-  \node (reldocs) [multidoc, below of=recall, xshift=20mm, yshift=-8mm]  {Relevant documents};
-  \node (irreldocs) [multidoc, fill=blue!5, below of=recall, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
+  \node (reldocs) [multidoc, below of=screenrel, xshift=20mm, yshift=-8mm]  {Relevant documents};
+  \node (irreldocs) [multidoc, fill=blue!5, below of=screenrel, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
   %\node
@@
 
-  \node [coord, left=of stopscreen, xshift=-45mm] (c1)  {}; %\cmark{1} 
+  \node [coord, left=of screenrel, xshift=-45mm, yshift=-10mm] (c1)  {}; %\cmark{1} 
   \node [coord, above=of model, yshift=2em] (c2)  {}; %\cmark{1}
 
-  \node [coord, left=of recall, xshift=-45mm] (c3)  {}; %\cmark{1}
+  %\node [coord, left=of recall, xshift=-45mm] (c3)  {}; %\cmark{1}
 
- \path (stopscreen.west) to node [very near start, yshift=1em] {$n$} (c1);
+ %\path (screenrel.south) to node [very near start, yshift=-1em] {$n$} (c1);
 
- 	\draw [o->,norm] (stopscreen.west) -- (c1) |-  (c2) -- (model.north);
+ 	\draw [o->,norm] (screenrel.south) |- (c1) |-  (c2) -- (model.north);
 
-  \path (recall.west) to node [very near start, yshift=1em] {$n$} (c3);
+  %\path (recall.west) to node [very near start, yshift=1em] {$n$} (c3);
 
  %\draw [o->,norm] (recall.west) -- (c3) |-  (c2) -- (model.north);
- \draw [o->,norm] (recall.west) -- (c3) |-  (screenr2.west);
 
- \draw [o->,norm] (recall.south) -- (reldocs);
- \draw [o->,norm] (recall.south) -- (irreldocs);
+ %\draw [o->,norm] (screenrel.south) -- (reldocs);
+ %\draw [o->,norm] (screenrel.south) -- (irreldocs);
 

### 2019-07-03 pres/rapid-review.tex
first commit

\documentclass[9pt]{beamer}
\mode<presentation>
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usepackage{xmpmulti}
\usepackage{animate}
\usepackage{tcolorbox}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{csquotes}

\usetheme{Singapore}
%\usecolortheme{seahorse}

\usefonttheme{professionalfonts}

\title[Word \& Doc Embeddings]{Machine Assisted Rapid Review}
\author{Max Callaghan, Finn Müller-Hansen}
\institute[MCC]{
	\includegraphics[height=1cm,width=2cm]{../images/MCC_Logo_RZ_rgb.jpg}
}

\newif\ifframeinlbf
\frameinlbftrue
\makeatletter
\newcommand\listofframes{\@starttoc{lbf}}
\makeatother

\addtobeamertemplate{frametitle}{}{%
	\ifframeinlbf
	\addcontentsline{lbf}{section}{\protect\makebox[2em][l]{%
			\protect\usebeamercolor[fg]{structure}\insertframenumber\hfill}%
		\insertframetitle\par}%
	\else\fi
}

\newtheorem*{remark}{}

\bibliographystyle{apalike}

\begin{document}

\begin{frame}
	\titlepage
\end{frame}

\section{Introduction}

\begin{frame}{Screening}

In systematic review, we want to have a broad search query, to make sure we include as many relevant documents as possible

\bigskip

This means we need to screen all of the documents the search query returns, in order to find a clean set of relevant documents.

\end{frame}

\begin{frame}{Machine Assisted Screening}

\begin{itemize}

	\item Scientific literature is expanding in all subject areas, increasing the work needed to conduct systematic reviews.

	\item The process generates training data, making it an ideal setting for applying \textit{Active Learning} see \cite{Settles2009}

	\item An expanding literature exists on using active learning for screening in systematic reviews \cite{OMara-Eves2015}

\end{itemize}

\end{frame}

\section{Approach}

\begin{frame}{Active Learning}

\begin{figure}
	\includegraphics[width=0.8\linewidth]{../images/flow_basic.pdf}
\end{figure}

When do we stop? How can we describe when we stopped, and what we think this means about the literature we have selected and not selected?

$ Precision=1.0 $, but what about $Recall$?

\end{frame}

\begin{frame}{Missing stopping criteria}

Research on a Stopping criterion in AL literature suggest stopping when the confidence or performance of the classifier drops. These indicate when the ML part is no longer generating useful results, but do not relate to recall: Vlachos stress that ``the remaining instances should be considered redundant only for the given model and feature representation'' \cite{Vlachos2008}

\medskip

Stopping criteria in the systematic review literature are not well articulated:

\blockcquote{miwa2014}{Once a given stopping criterion is reached (for example, when all relevant studies have been identified, or when the reviewers have run out of time for manual screening), the process ceases, and the remainder of studies not yet screened manually is discarded}

\end{frame}

\begin{frame}{Missing stopping criteria}
Statistical approaches are hinted at but underdeveloped:

\blockcquote{Przybya2018}{Deciding when to stop can be left to the end-user, eg, if the prioritised references are consistently irrelevant, or heuristics or statistical approaches (based on an unbiased sample of the remaining references) can inform this decision. We intend to investigate this problem in the future}

\medskip

A reliable stopping criterion is still a research gap:

\blockcquote{bannach-brown2019}{The problem of choosing a cut-off threshold, equivalent to deciding when to stop when using a model for prioritising relevant documents, remains an open research question in information retrieval}

\end{frame}

\begin{frame}{Relevance of unseen documents is insufficient}
\begin{figure}
	\includegraphics[width=\linewidth]{../images/proportions_1}<1>
	\includegraphics[width=\linewidth]{../images/proportions_2}<2>
\end{figure}
\end{frame}

\begin{frame}{Relevance of unseen documents is insufficient}
\begin{columns}
	\begin{column}{0.6\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/proportions_1}

			\vspace{-1em}

			\includegraphics[width=\linewidth]{../images/proportions_2}
		\end{figure}
	\end{column}
	\begin{column}{0.4\linewidth}
			\only{It doesn't just matter what proportion of unseen documents is relevant, but also what proportion of documents is unseen
			}<1->

		\bigskip

		\only{We can calculate the recall with the relevance of the unseen docs, the relevance of the population and porportion of documents seen

			\[ 1 - \frac{pU}{pP} \frac{U}{P} \]

		}<2->
	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Uncertainty of pU}

 However, without perfect knowledge, we don't \textit{know} the relevance of unseen documents: but we can \textit{estimate} it through taking a sample. We do know \(U\), the number of unseen documents, \(S\), the number of seen documents, and \( pS \), the relevance of seen documents,

 \bigskip
\begin{tabular}{p{0.618\linewidth} p{0.382\linewidth}}
 For a given p-value, we can generate an Agresti-Coull interval & 
 \( CI_{AC} = \tilde{p} \pm \kappa(\tilde{p}\tilde{q})^{1/2}\tilde{n}^{-1/2} \) \\
  & \\
 We can use the outer limit \( \dot{p} = \max\tilde{p} \) to estimate the relevance of all documents 
  &  \( \dot{p}P = \frac{pS*S + \dot{p}U*U}{N}  \) \\
  & \\
 And plug both of these into the original equation to generate a minimum estimated recall & 
 \( \hat{R} = 1- \frac{\dot{p}U}{\dot{p}P} \frac{U}{P} \)

\end{tabular}

\bigskip

We can set the recall and confidence interval prior to the start of screening

\end{frame}

\begin{frame}{Uncertainty of pU}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/sample_status.png}
		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/sample_recall.png}
		\end{figure}
	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Uncertainty of pU}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/sample_status_2.png}
		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/sample_recall_2.png}
		\end{figure}
	\end{column}

\end{columns}

\bigskip

	\[   CI_{AC} = \tilde{p} \pm \kappa(\tilde{p}\tilde{q})^{1/2}\tilde{n}^{-1/2}  \]

\[ \hat{R} = 1- \frac{\dot{p}U}{\dot{p}P} \frac{U}{P} \]

\end{frame}

\begin{frame}{Active Learning with Robust Probabilistic Stopping Criteria}

\begin{figure}
	\includegraphics[width=0.8\linewidth]{../images/flow.pdf}
\end{figure}

\end{frame}

\section{Evaluation}

\begin{frame}{Evaluation on existing dataset(s)}
To be done - ours? Someone elses? Simulations with many types/specifications of ML models, showing reduction in work, predicted and actual recall

\bigskip

Necessary?
\end{frame}

\section{Question}

\begin{frame}{How big should the sample be?}

What confidence intervals do we get for different levels of relevance and different sample sizes?

\bigskip

Increasing N reduces unseen docs and size of confidence interval

\end{frame}

\begin{frame}{Do we need to take a sample?}

Could we just keep looking at the most relevant documents (this would make our estimates more conservative) or does this just make things complicated without saving time?

\end{frame}

\begin{frame}{When do we move from the most relevant documents to a random sample?}

Is this when we find no more relevant documents? When no more relevant documents are predicted? When model performance on new data drops below a certain level? Some other criterion or combination of criteria?

\end{frame}

\begin{frame}{Do we need to rate all documents in the sample?}

Can the system estimate recall as we rate documents, and stop as soon as we have reached the threshold?

\end{frame}

\begin{frame}{What if we prioritised irrelevant documents?}

In this way we could be certain about recall, and uncertain about precision.

\bigskip

We spend a lot of time looking at irrelevant documents (so learn less about the literature during the process), but sometimes exclusion decisions can be faster than inclusion decisions.

\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}

This is a new way to use machine learning in screening

\begin{itemize}
	\item<1-> Offers robust stopping criteria
	\item<2-> ML model performance does not determine accuracy of document selection, only influences work reduction
\end{itemize}

\end{frame}

\begin{frame}
\listofframes
\end{frame}

\begin{frame}{References}
\bibliography{../mendeley}
\end{frame}

\end{document}

### 2019-07-17 manuscript/abstract.tex
bmc style, background and disproofs written

\documentclass{article}
\usepackage{standalone}

\begin{document}

Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. Machines and humans  work together, with the human providing training data, and the machine optimising the documents a human screens. However, current approaches lack robust stopping criteria, meaning that such systems are hard to implement in live reviews. This paper introduces a workflow for working with robust and flexible statistical stopping criteria, that offer real work reductions on the basis of a given confidence level of reaching a given recall.

\end{document}

### 2019-07-17 manuscript/main.tex
bmc style, background and disproofs written

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{todonotes}
\usepackage{graphicx}
%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

	%%% Start of article front matter
	\begin{frontmatter}

		\begin{fmbox}
			\dochead{Research}

			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			%%                                          %%
			%% Enter the title of your article here     %%
			%%                                          %%
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

			\title{Statistical Stopping Criteria for Active Learning in Systematic Review Screening}

			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			%%                                          %%
			%% Enter the authors here                   %%
			%%                                          %%
			%% Specify information, if available,       %%
			%% in the form:                             %%
			%%   <key>={<id1>,<id2>}                    %%
			%%   <key>=                                 %%
			%% Comment or delete the keys which are     %%
			%% not used. Repeat \author command as much %%
			%% as required.                             %%
			%%                                          %%
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

			\author[
			addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
			corref={aff1},                       % id of corresponding address, if any
			%noteref={n1},                        % id's of article notes, if any
			email={callaghan@mcc-berlin.net}   % email address
			]{\inits{MW}\fnm{Max W} \snm{Callaghan}}
			\author[
			addressref={aff1},
			email={mueller-hansen@mcc-berlin.net}
			]{\inits{FMH}\fnm{Finn} \snm{M\"{u}ller-Hansen}}

			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			%%                                          %%
			%% Enter the authors' addresses here        %%
			%%                                          %%
			%% Repeat \address commands as much as      %%
			%% required.                                %%
			%%                                          %%
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

			\address[id=aff1]{%                           % unique id
				\orgname{Mercator Research Institute on Global Commons and Climate Change}, % university, etc
				\street{Torgauer Straße},                     %
				\postcode{10829}                                % post or zip code
				\city{Berlin},                              % city
				\cny{Germany}                                    % country
			}
			\address[id=aff2]{%
				\orgname{Priestley International Centre for Climate, University of Leeds, Leeds },
				%\street{Dsternbrooker Weg 20},
				\postcode{LS2 9JT}
				\city{Leeds},
				\cny{United Kingdom}
			}

			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			%%                                          %%
			%% Enter short notes here                   %%
			%%                                          %%
			%% Short notes will be after addresses      %%
			%% on first page.                           %%
			%%                                          %%
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

			\begin{artnotes}
				%\note{Sample of title note}     % note to the article
				%\note[id=n1]{Equal contributor} % note, connected to author
			\end{artnotes}

		\end{fmbox}% comment this for two column layout

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		%%                                          %%
		%% The Abstract begins here                 %%
		%%                                          %%
		%% Please refer to the Instructions for     %%
		%% authors on http://www.biomedcentral.com  %%
		%% and include the section headings         %%
		%% accordingly for your article type.       %%
		%%                                          %%
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		\begin{abstractbox}

			\begin{abstract} % abstract
				\parttitle{First part title} %if any
				Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. 
				Machines and humans  work together, with the human providing training data, and the machine optimising the documents a human screens. 
				However, current approaches lack robust stopping criteria, meaning that such systems are hard to implement in live reviews. 
				This paper introduces a workflow for working with robust and flexible statistical stopping criteria, that offer real work reductions on the basis of a given confidence level of reaching a given recall.

			\end{abstract}

			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			%%                                          %%
			%% The keywords begin here                  %%
			%%                                          %%
			%% Put each keyword in separate \kwd{}.     %%
			%%                                          %%
			%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

			\begin{keyword}
				\kwd{sample}
				\kwd{article}
				\kwd{author}
			\end{keyword}

			% MSC classifications codes, if any
			%\begin{keyword}[class=AMS]
			%\kwd[Primary ]{}
			%\kwd{}
			%\kwd[; secondary ]{}
			%\end{keyword}

		\end{abstractbox}
		%
		%\end{fmbox}% uncomment this for twcolumn layout

	\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%

	%%%%%%%%%%%%%%%%%%%%%%%%
	%% Introduction
	\section*{Background}

	Machine learning for evidence synthesis is a growing field which promises to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
	A major strand of the literature works on screening,  
	where several papers have developed or evaluated active learning. 
	Active learning is an iterative process where documents screened by humans are used to train a machine learning model to predict the relevance of unseen papers.
	The most relevant studies are rated by a human, which generates more labels to feed back to the machine. 
	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or above a given threshold proportion of relevant studies - before having seen all the documents in the corpus. 
	The documents not yet seen by the human when they reach the given recall threshold have been represented as the work saved.

	In live review settings, however, the recall - or proportion of relevant studies screened - remains an unknown until all documents have been screened. In order for work to really be saved, reviewers have to stop screening while uncertain about the recall. The lack of appropriate stopping criteria has been identified as a research gap \cite{bannach-brown2019}, and existing approaches have fallen into the following categories
	\begin{itemize}
		\item \textbf{Sampling criteria:} Reviewers estimate the number of relevant documents by taking a random sample at the start of the process. They stop when this number, or a given proportion of it, has been reached \cite{Shemilt2014}
		\item \textbf{Heuristics:} Reviewers stop when a given number of irrelevant articles are seen in a row \cite{Przybya2018}. 
		\item \textbf{Pragmatic criteria:} Reviewers stop when they run out of time \cite{miwa2014}. 
	\end{itemize}

	We show in this paper how first two criteria are inadequate.
	This lack of adequate criteria challenges users of AL systems in live reviews, as authors are uncertain when to stop, and have no reliable way of reporting when they stopped and what that may say about the recall they reached.

	This study proposes a system for estimating the recall based on random sampling of remaining documents. We draw on the literature on binomial distributions \cite{Brown2001}, to illustrate how AL users can predefine a threshold in terms of uncertainty and recall, and use this to transparently save work with machine learning, while making a statement like ``There is a <5\% chance that we achieve a recall under 95\%''.

	We evaluate this stopping criteria on real-world systematic review datasets on which active learning systems have previously been tested.

	\section*{Methods}

	\subsection*{Existing Stopping Criteria for Active Learning}

	We start by explaining the sampling and heuristic based stopping criteria before showing with toy data how the criteria falls short. Then we introduce our own suggested stopping criteria, and show its benefits with toy data, before testing all criteria on real world datasets.

	\subsubsection*{Sampling Based Stopping Criteria}

	The stopping criterion suggested by \cite{Shemilt2014} involves establishing the Baseline Inclusion Rate (BIR), by taking a random sample at the beginning of screening. This is used to estimate the number of relevant documents in the whole dataset. Reviewers continue to screen until this number, or a proportion of it corresponding to the desired level of recall, is reached.

	However, the estimation of the BIR fails to take into account the sampling uncertainty. 
	Figure \ref{bir-sampling}.a shows for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant), the predicted number of documents after each document drawn. Depending on the luck of the draw, one might estimate the true number of documents as being
	\begin{itemize}
		\item much higher than the true value, meaning that the stopping criterion would never be reached - and no work could be saved
		\item much lower than the true value, meaning that the stopping criterion would be reached before the desired level of recall was actually achieved
	\end{itemize}

	We can estimate the uncertainty of a binomial outcome using an Agresti-Coull confidence interval (which has been shown to be conservative at levels near 0 \todo{this is less relevant here, go back to literature and see if another is better} \cite{Brown2001}). Figure \ref{bir-sampling}.b shows the 95\% confidence interval (brown) and its upper bound (green) in the same set of random draws. 
	Where the green line is above the above the lower (upper) grey line, identification of 100\% (95\%) of the estimated number of studies would be impossible, and no work would be saved.

	Figure \ref{bir-variation} explores different scenarios of true relevance and probability thresholds, showing that the higher the true proportion of relevant documents, the fewer documents one would have to sample to make the identification of a given fraction (the true number of relevant results divided by the upper estimate) possible. \todo{Why does it behave like this, is there a theoretical maximum for each? something to do with absolute values? a Bug? Is uncertainty a function of being near to 0?} Even with a relatively high proportion of relevant documents, and a sample of 500, all relevant documents would have to be identified to have a 95\% likelihood of achieving 95\% recall. With low proportions of relevant documents, only an 80\% recall can be estimated with confidence, even after a sample of 2000, and subsequently identifying all relevant documents. N.B, these results ignore the random element of sample order, which explains the variation on figure \ref{bir-sampling}. They show upper confidence bounds for accurate estimates of relevance.

	\begin{figure}
	\includegraphics[width=\linewidth]{../images/bir_sampling.png}
	\caption{Uncertainty in estimating the baseline inclusion rate. a. shows the estimated BIR after drawing each document, for 50 random samples of from 20,000 documents, after each document is drawn. b. Shows the 95\% Confidence interval (brown) and the upper bound (green) for each of the same samples. The two lines correspond to 100\% of relevant documents, and $\frac{100}{0.95}$\% of relevant documents.}
	\label{bir-sampling}
\end{figure}

\begin{figure}
	\includegraphics[width=0.5\linewidth]{../images/bir_variation.png}
	\caption{Maximum indentifiable fraction of relevant results under different levels of true relevance $p$ and confidence thresholds $\alpha$. Maximum identifiable fraction is the true number of relevant documents divided by the upper estimate.}
	\label{bir-variation}
\end{figure}

\subsubsection*{Heuristic Based Stopping Criteria}

Some studies give the example of heuristic based stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Przybya2018} [+REFS]. We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low. We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; and 2) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.

Figure \ref{unseen-proportions} illustrates this second point. We show two scenarios with identical low proportions of relevant documents (note that estimating a proportion lower than 0.02 with 95\% confidence requires 160 consecutive irrelevant documents)\todo{demonstrate?}. Recall is not simply a function of the relevance of unseen documents, but also of the number of unseen documents. This also means that where machine learning has performed well (as in the top figure), low proportions of irrelevant documents in those that remain are indicative of lower recall than where ML has performed less well.

\begin{figure}
	\includegraphics[width=\linewidth]{../images/proportions_1.png}
	\includegraphics[width=\linewidth]{../images/proportions_2.png}
	\caption{Similar low proportions of relevant documents in unseen documents with different consequences for recall. The top bar shows a random distribution of relevant documents (green) and irrelevant documents (red) at a given proportion of relevance. The bottom bar  shows distributions of relevant and irrelevant documents in hypothetical sets of seen (right) and unseen (left - transparent) documents.}
	\label{unseen-proportions}
\end{figure}

\subsection*{A Statistical Stopping Criterion for Active Learning}
	Figure \ref{flow} shows a workflow for the approach proposed in this paper. The random sampling occurs at the end of the process, and is used to estimate the number of relevant documents remaining and the total number of relevant documents. The rationale is to limit the uncertainty to a subsection of the dataset: that which has not yet been screened. As reviewers continue to draw random documents, the uncertainty range decreases, and the proportion of the data about which one is uncertain also decreases.

	\begin{figure}
		\includegraphics[width=0.5\linewidth]{../images/flow}
		\caption{A workflow for active learning in screening with a statistical stopping criterion}
		\label{flow}
	\end{figure}

	\begin{table}
		\begin{tabular}{p{0.15\linewidth} p{0.425\linewidth} p{0.425\linewidth}}
			Parameter & Description & Estimation \\
			\hline
			$ N  $ &  Total number of studies & \textit{Observed} \\
			$ S $ & Number of studies coded by humans & \textit{Observed} \\
			$ U $ & Number of studies not yet coded by humans & \textit{Observed} ($N - S$) \\
			$ p_S $ & relevance of documents rated by humans & \textit{Observed} \\
			$ \alpha $ & Acceptable uncertainty level & \textit{Given} \\
			$ \kappa $ & & \\
			$ n $ & Number of randomly drawn documents & \textit{Observed} \\
			$ \tilde{n} $ & & \\
			$ \kappa $ & & \\
			$ \hat{R} $ & Estimated recall & \\
			$ \dot{R} $ & Minimum estimated recall at a given confidence threshold & \\
		\end{tabular}
		\caption{Parameters}
		\label{parameters}
	\end{table}

	Table \ref{parameters} shows the parameters, known, estimated, and given, available during the random sampling process and required to estimate recall. Expanding on what was stated before, recall - at a given confidence threshold is a function of 1) the upper estimate of the relevance of remaining documents, 2) The estimated relevance of all documents in the dataset, 3) The proportion of documents not yet seen.

	Figure \ref{sample-recall} shows the minimum estimated recall for a set of confidence intervals, along with the actual recall (in grey) for a case where 800 out of 2,000 documents have been reviewed, 2\% of remaining documents are relevant, and 20\% of all documents are relevant. We see that after 95\% recall has actually been achieved, but before 100\% of documents have been been seen, we can be confident at each of the given confidence levels that 95\% of relevant documents habe identified. All estimates of recall only reach 100\% after all documents have been seen, as it is not possible to exclude the possibility, at any given confidence level, that the proportion of relevant documents is greater than 0 \todo{Actually maybe this is possible if we work with whole numbers? How does that change things?}. Figure \ref{sample-recall-variation} shows the various trajectories of a 95\% minimum recall (blue) and actual recall (grey) given the same paramaters for 200 random samples.

	\begin{figure}
		\includegraphics[width=\linewidth]{../images/sample_recall}
		\caption{Sample recall}
		\label{sample-recall}
	\end{figure}

	\begin{figure}
		\includegraphics[width=\linewidth]{../images/sample_recall_2}
		\caption{Sample recall variation}
		\label{sample-recall-variation}
	\end{figure}

	That the blue lines consistently meet a 95\% threshold to the right of the grey lines indicates that reviewers have to see more than 95\% of relevant documents in order to be confident that they have achieved their threshold. We call the number of documents it is necessary to review to establish with confidence that reviewers have achieved 95\% recall after they have already achieved 95\% recall the \textit{additional burden}.

	In figure [], we investigate how the total number of documents and the proportion of relevant documents affect the additional burden \todo{Use the number of documents seen instead of p}.

	\begin{figure}
		\includegraphics[width=\linewidth]{../images/additional_burden}
		\caption{Additional burden for 20 random samples where the 95\% threshold has been reached, and the remaining documents are 1\%, 2\% or 5\% relevant, for different total numbers of documents.}
		\label{additional-burden}
	\end{figure}

	\subsection*{Evaluation}

	\subsubsection*{Evaluation Data}

	\section*{Results}

	\section*{Discussion}

	\section*{Conclusion}

\begin{backmatter}

	\section*{Competing interests}
	The authors declare that they have no competing interests.

	\section*{Author's contributions}
	Text for this section \ldots

	\section*{Acknowledgements}
	Max Callaghan is supported by a PhD scholarship from the Heinrich Böll Foundation

	\bibliography{../mendeley}
	\bibliographystyle{bmc_article}

\end{backmatter}
\end{document}

### 2019-07-18 manuscript/main.tex
additional burden - scrape review

@@
 		\begin{fmbox}
-			\dochead{Research}
+			\dochead{Methodology}
 
@@
 			\begin{abstract} % abstract
-				\parttitle{First part title} %if any
+				%\parttitle{First part title} %if any
 				Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. 
@@
 			\begin{keyword}
-				\kwd{sample}
-				\kwd{article}
-				\kwd{author}
+				\kwd{Systematic Review}
+				\kwd{Machine Learning}
+				\kwd{Active Learning}
+				\kwd{Stopping Criteria}
 			\end{keyword}
@@
 	Figure \ref{bir-variation} explores different scenarios of true relevance and probability thresholds, showing that the higher the true proportion of relevant documents, the fewer documents one would have to sample to make the identification of a given fraction (the true number of relevant results divided by the upper estimate) possible.
-\todo{Why does it behave like this, is there a theoretical maximum for each?
-something to do with absolute values?
-a Bug?
-Is uncertainty a function of being near to 0?} Even with a relatively high proportion of relevant documents, and a sample of 500, all relevant documents would have to be identified to have a 95\% likelihood of achieving 95\% recall.
+	\todo{Why does it behave like this, is there a theoretical maximum for each? something to do with absolute values? a Bug? Is uncertainty a function of being near to 0?} 
+	Even with a relatively high proportion of relevant documents, and a sample of 500, all relevant documents would have to be identified to have a 95\% likelihood of achieving 95\% recall. 
 	With low proportions of relevant documents, only an 80\% recall can be estimated with confidence, even after a sample of 2000, and subsequently identifying all relevant documents. 
@@
 	Figure \ref{sample-recall} shows the minimum estimated recall for a set of confidence intervals, along with the actual recall (in grey) for a case where 800 out of 2,000 documents have been reviewed, 2\% of remaining documents are relevant, and 20\% of all documents are relevant.
-We see that after 95\% recall has actually been achieved, but before 100\% of documents have been been seen, we can be confident at each of the given confidence levels that 95\% of relevant documents habe identified.
+	We see that after 95\% recall has actually been achieved, but before 100\% of documents have been been seen, we can be confident at each of the given confidence levels that 95\% of relevant documents have been identified.
 	All estimates of recall only reach 100\% after all documents have been seen, as it is not possible to exclude the possibility, at any given confidence level, that the proportion of relevant documents is greater than 0 \todo{Actually maybe this is possible if we work with whole numbers?
@@
 
-	In figure [], we investigate how the total number of documents and the proportion of relevant documents affect the additional burden \todo{Use the number of documents seen instead of p}.
+	In figure \ref{additional-burden}, we investigate how the total number of documents and the proportion of relevant documents affect the additional burden \todo{Use the number of documents seen instead of p?}.
+We demonstrate the additional burden decreases \todo{how?
+Non-linearly?} with increasing total numbers of documents.
+With decreasing relevance of the remaining documents (holding the recall constant at 95\%, this means that fewer documents remain unseen), the additional burden increases.
+This implies that where the machine learning has more successfully identified documents and reduced burden, the additional burden required to establish the level of recall is higher.
+This is because the proportion of documents not seen is greater, so uncertainty around the relevance of that proportion has a greater impact on the estimation of recall.
 
@@
 
+	When to move to random sampling remains an open question.
+Our current approach is to use the predicted relevance of the machine learning model as a guide.
+We calculate the [minimum/maximum] estimated recall based on observing the predicted relevance level in [x] random documents.
+If the estimated recall would be below the threshold, then we begin drawing random documents.
+
+	The stopping criteria explained here is
+
 	\subsection*{Evaluation}
+
+	We evaluate each of the three criteria discussed on real world test data.
+With each method, we run 100 runs and record the recall error and actual work saved each time.
+We operationalise the heuristic stopping criteria with 10, 20, and 50 irrelevant records in a row.
+For simplicity, we use a basic SVM model [REFS], with 1-2 word n-grams in the document abstracts used as input data.
+We start with random samples of [100, 200, and 500] documents.
 

### 2019-07-18 analysis/get_data/scrape_pbr.py

@@ -1,0 +1,3 @@
+#!/usr/bin/python3
+
+import requests

### 2019-07-18 analysis/rapid_review.py

@@ -1,0 +1,30 @@
+import pandas as pd
+import xml.etree.ElementTree as ET
+
+class Screen:
+    '''
+    A class to record and advance the state of ml-aided document screening
+    '''
+    def __init__(self, N):
+        self.N = N
+
+
+def get_field(a, f):
+    try:
+        return list(a.iter(f))[0].text
+    except:
+        return None
+    
+def parse_pmxml(path):
+    tree = ET.parse(path)
+    root = tree.getroot()
+    docs = []
+    for a in root.iter('PubmedArticle'):
+        docs.append({
+            "ab": get_field(a, 'AbstractText'),
+            "pmid": get_field(a, 'PMID'),
+            "ti": get_field(a, 'ArticleTitle')
+        })
+    df = pd.DataFrame.from_dict(docs)
+    return df
+    

### 2019-07-18 analysis/get_data/get_cohen_xmls.py
first evaluations + cohen data

@@ -1,0 +1,24 @@
+import requests
+import time
+from itertools import zip_longest
+def grouper(n, iterable, padvalue=None):
+    "grouper(3, 'abcdefg', 'x') --> ('a','b','c'), ('d','e','f'), ('g','x','x')"
+    return zip_longest(*[iter(iterable)]*n, fillvalue=padvalue)
+
+pmids = []
+with open('data/epc-ir.clean.tsv','r') as f:
+    for l in f:
+        pmids.append(l.split('\t')[2])
+
+pmids = set(pmids)
+
+for i, group in enumerate(grouper(200, pmids)):
+    group = [x for x in group if x is not None]
+    if os.path.exists(f'data/cohen_all_{i}.xml'):
+        continue
+    with open(f'data/cohen_all_{i}.xml','w') as f:
+        pm_url = 'https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi?db=pubmed&retmode=xml&id='
+        pm_url = pm_url+",".join(group)
+        result = requests.post(pm_url)
+        f.write(str(result.content, "utf-8"))
+        time.sleep(5)

### 2019-07-18 analysis/rapid_review.py

@@ -1,12 +1,115 @@
 import pandas as pd
 import xml.etree.ElementTree as ET
+from nltk import wordpunct_tokenize
+from nltk import WordNetLemmatizer
+from nltk import sent_tokenize
+from nltk import pos_tag
+from nltk.corpus import stopwords as sw
+import string
+punct = set(string.punctuation)
+from nltk.corpus import wordnet as wn
+from sklearn.feature_extraction.text import CountVectorizer, TfidfVectorizer
+stopwords = set(sw.words('english'))
+import random
 
-class Screen:
+def lemmatize(token, tag):
+        tag = {
+            'N': wn.NOUN,
+            'V': wn.VERB,
+            'R': wn.ADV,
+            'J': wn.ADJ
+        }.get(tag[0], wn.NOUN)
+        return WordNetLemmatizer().lemmatize(token, tag)
+
+def tokenize(X):
+    for sent in sent_tokenize(X):
+        for token, tag in pos_tag(wordpunct_tokenize(sent)):
+            token = token.lower().strip()
+            if token in stopwords:
+                continue
+            if all(char in punct for char in token):
+                continue
+            if len(token) < 3:
+                continue
+            if all(char in string.digits for char in token):
+                continue
+            lemma = lemmatize(token,tag)
+            yield lemma
+
+class ScreenScenario:
     '''
     A class to record and advance the state of ml-aided document screening
     '''
-    def __init__(self, N):
-        self.N = N
+    def __init__(self, df, models, samples):
+        self.df = df
+        self.models = models
+        self.N = df.shape[0]
+        self.r_docs = df[df["relevant"]==1].shape[0]
+        self.p = self.r_docs/self.N
+        self.bir = None
+        self.bir_upperbound = None
+        self.df['seen'] = 0
+        self.seen_docs = 0
+        self.r_seen = 0
+        self.samples = samples
+        self.results = []
+        self.iteration_size = 20
+        self.recall = 0
+        
+        self.X = TfidfVectorizer(
+            ngram_range=(1,2),
+            min_df=5, max_df=0.6, strip_accents='unicode', 
+            max_features=10000,
+            use_idf=1,
+            smooth_idf=1, sublinear_tf=1,
+            stop_words="english",tokenizer=tokenize
+        ).fit_transform(df['ab'])
+
+    def reset(self):
+        self.df['seen'] = 0
+        self.seen_docs = 0
+
+    def screen_bir(self):
+        for s in self.samples:
[…]

### 2019-07-19 analysis/get_data/get_cohen_xmls.py
setup for scenario runs

@@ -10,7 +10,8 @@ with open('data/epc-ir.clean.tsv','r') as f:
     for l in f:
         pmids.append(l.split('\t')[2])
 
-pmids = set(pmids)
+pmids = list(set(pmids))
+pmids.sort()
 
 for i, group in enumerate(grouper(200, pmids)):
     group = [x for x in group if x is not None]

### 2019-07-19 analysis/get_data/scrape_pbr.py

@@ -1,3 +1,32 @@
 #!/usr/bin/python3
 
 import requests
+from bs4 import BeautifulSoup
+
+
+def parse_ncbi(url):
+    result = requests.get(url)
+    c = result.content
+    pmids = []
+    for row in soup.find('table').find_all('tr'):
+        if not row.find_all('td'):
+            continue
+        if row.find('td').attrs['colspan']=='2':
+            continue
+        if row.find_all('td')[1].string:
+            pmids.append(row.find_all('td')[1].string)
+
+    pm_url = 'https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi?db=pubmed&retmode=xml&id='
+    pm_url = pm_url+",".join(pmids)
+
+    result = requests.post(pm_url)
+    return result
+
+result = parse_ncbi('https://www.ncbi.nlm.nih.gov/books/NBK44537/')
+with open('data/pbr_included.xml','w') as f:
+    f.write(str(result.content, "utf-8"))
+
+result = parse_ncbi('https://www.ncbi.nlm.nih.gov/books/NBK44542/')
+with open('data/pbr_excluded.xml','w') as f:
+    f.write(str(result.content, "utf-8"))
+

### 2019-07-19 analysis/rapid_review.py

@@ -11,15 +11,20 @@ from nltk.corpus import wordnet as wn
 from sklearn.feature_extraction.text import CountVectorizer, TfidfVectorizer
 stopwords = set(sw.words('english'))
 import random
+import scipy.stats as st
+import numpy as np
+import time, sys
+from IPython.display import clear_output
+import math
 
 def lemmatize(token, tag):
-        tag = {
-            'N': wn.NOUN,
-            'V': wn.VERB,
-            'R': wn.ADV,
-            'J': wn.ADJ
-        }.get(tag[0], wn.NOUN)
-        return WordNetLemmatizer().lemmatize(token, tag)
+    tag = {
+        'N': wn.NOUN,
+        'V': wn.VERB,
+        'R': wn.ADV,
+        'J': wn.ADJ
+    }.get(tag[0], wn.NOUN)
+    return WordNetLemmatizer().lemmatize(token, tag)
 
 def tokenize(X):
     for sent in sent_tokenize(X):
@@ -36,12 +41,29 @@ def tokenize(X):
             lemma = lemmatize(token,tag)
             yield lemma
 
+def ci_ac(X, n, a):
+    def get_k(a):
+        a = (1-(1-a)/2)
+        return st.norm.ppf(a)
+
+    k = get_k(a)
+
+    X_tilde = X + k**2/2
+    n_tilde = n + k**2
+    p_tilde = X_tilde / n_tilde
+    q_tilde = 1-p_tilde
+
+    ci = k* np.sqrt(p_tilde*q_tilde/n_tilde) 
+
+    return p_tilde, ci
+
 class ScreenScenario:
     '''
     A class to record and advance the state of ml-aided document screening
     '''
-    def __init__(self, df, models, samples):
+    def __init__(self, df, models, sample, irrelevant_heuristic, dataset):
         self.df = df
+        self.dataset = dataset
         self.models = models
         self.N = df.shape[0]
         self.r_docs = df[df["relevant"]==1].shape[0]
@@ -51,10 +73,25 @@ class ScreenScenario:
         self.df['seen'] = 0
         self.seen_docs = 0
         self.r_seen = 0
-        self.samples = samples
+        self.r_predicted = None
+        self.r_predicted_upperbound = None
+        self.s = sample
+        self.irrelevant_heuristic = irrelevant_heuristic # how many irrelevant do we have to see in a row in order to stop
         self.results = []
-        self.iteration_size = 20
-        self.recall = 0
+        self.iteration_size = 50
[…]

### 2019-07-19 analysis/run_scenarios.py

@@ -1,0 +1,70 @@
+import pandas as pd
+import numpy as np
+import rapid_review as rr
+import os
+import matplotlib.pyplot as plt
+import importlib
+from sklearn.svm import SVC, OneClassSVM
+
+##########################################
+## Pull the document metadata from the xml files from the Pubmed API
+document_index = None
+
+for fpath in os.listdir('../data/'):
+    if "cohen_all" in fpath:
+        ndf = rr.parse_pmxml(f'../data/{fpath}')
+    if document_index is None:
+        document_index = ndf
+    else:
+        document_index = pd.concat([document_index,ndf])
+        
+document_index = document_index.drop_duplicates()
+print(document_index.shape)
+document_index.head()
+
+#############################
+## Load the cohen database of SRs
+https://dmice.ohsu.edu/cohenaa/systematic-drug-class-review-data.html
+cohen_db = pd.read_csv(
+    '../data/epc-ir.clean.tsv',
+    sep='\t',header=None,
+    names=["review","EID","PMID","relevant","fulltext_relevant"]
+)
+
+cohen_db['relevant'] = np.where(cohen_db['relevant']=="I",1,0)
+cohen_db = cohen_db[["review","PMID","relevant"]]
+
+cohen_db.head()
+
+models = [
+    SVC(kernel='linear',C=5,probability=True)
+]
+iterations = 500
+
+results = []
+rs_results = []
+paths = []
+for name, group in cohen_db.groupby('review'):
+    df = pd.merge(
+        group,
+        document_index,
+    )
+    df = df.dropna().reset_index(drop=True)
+    for s in [100, 200, 500]:
+        ss = rr.ScreenScenario(
+            df, models, s, 50, name
+        )
+        for i in range(iterations):
+            print(i)
+            rs_results.append(ss.screen(i, True))
+            results.append(ss.screen(i))
+            paths.append({
+                "dataset": name,
+                "work_path": ss.work_track,
+                "recall_path": ss.recall_track
+            })
+
+results_df = pd.DataFrame.from_dict(results)
+rs_results_df = pd.DataFrame.from_dict(rs_results)
+rs_results_df.to_csv('../results/rs_results.csv', index=False)
+results_df.to_csv('../results/results.csv', index=False)

### 2019-07-19 analysis/run_scenarios.py
comment link

@@ -24,7 +24,8 @@ document_index.head()
 
 #############################
 ## Load the cohen database of SRs
-https://dmice.ohsu.edu/cohenaa/systematic-drug-class-review-data.html
+#https://dmice.ohsu.edu/cohenaa/systematic-drug-class-review-data.html
+
 cohen_db = pd.read_csv(
     '../data/epc-ir.clean.tsv',
     sep='\t',header=None,

### 2019-07-19 analysis/run_scenarios.py
load data typo

@@ -13,10 +13,10 @@ document_index = None
 for fpath in os.listdir('../data/'):
     if "cohen_all" in fpath:
         ndf = rr.parse_pmxml(f'../data/{fpath}')
-    if document_index is None:
-        document_index = ndf
-    else:
-        document_index = pd.concat([document_index,ndf])
+        if document_index is None:
+            document_index = ndf
+        else:
+            document_index = pd.concat([document_index,ndf])
         
 document_index = document_index.drop_duplicates()
 print(document_index.shape)

### 2019-07-22 analysis/run_scenarios.py
first results

@@ -40,7 +40,7 @@ cohen_db.head()
 models = [
     SVC(kernel='linear',C=5,probability=True)
 ]
-iterations = 500
+iterations = 5
 
 results = []
 rs_results = []
@@ -51,14 +51,18 @@ for name, group in cohen_db.groupby('review'):
         document_index,
     )
     df = df.dropna().reset_index(drop=True)
-    for s in [100, 200, 500]:
+    for s in [200, 500]:
         ss = rr.ScreenScenario(
             df, models, s, 50, name
         )
         for i in range(iterations):
             print(i)
-            rs_results.append(ss.screen(i, True))
-            results.append(ss.screen(i))
+            r = ss.screen(i, True) 
+            if r is not None:
+                rs_results.append(r)
+            r = ss.screen(i)
+            if r is not None:
+                results.append(r)
             paths.append({
                 "dataset": name,
                 "work_path": ss.work_track,

### 2019-07-22 manuscript/main.tex
adjust random sample threshold

@@
 \usepackage{graphicx}
+\usepackage{url}
 %\def\includegraphic{}
@@
 
-	We start by explaining the sampling and heuristic based stopping criteria before showing with toy data how the criteria falls short. 
+	We start by explaining the sampling and heuristic based stopping criteria before showing with toy data how the criteria fall short. 
 	Then we introduce our own suggested stopping criteria, and show its benefits with toy data, before testing all criteria on real world datasets.
@@
 \end{figure}
+
+\subsubsection*{Other stopping criteria}
+
+\cite{Wallace2010a} Develop a ``simple, operational stopping criterion'': stopping after half the documents have been screened.
+Although the criterion worked in their experiment, it is unclear how this could be generalised, and its development depended on knowledge of the true relevance values.
+\cite{Jonnalagadda2013} note that ``the reviewer can elect to end the process of classifying documents at any point, recognizing that stopping before reviewing all documents involves a trade-off of lower recall for reduced workload'', although clearly the reviewer lacks information about probable recall.
+\cite{Yu2019} adopt a more complicated stopping criterion \todo{How does it actually work??} which allows the user to target a specific level of recall.
+However, reviewers are not given the opportunity to specify a confidence level, and for two of the four datasets in which they tested their criteria, the median achieved recall at a stopping criteria targeting 95\% recall was below 95\%.
 
@@
 
-	We evaluate each of the three criteria discussed on real world test data.
-With each method, we run 100 runs and record the recall error and actual work saved each time.
+	We evaluate each of the three criteria discussed on real world test data. With each method, we run 100 runs and record 
+	\begin{itemize}
+		\item \textbf{Actual Recall}: The recall when the stopping criteria is met
+		\item \textbf{WS-SC}: Work saved when the stopping criteria is met \footnote{we do not use work saved over sampling, is it is not axiomatic that seeing 95\% of a random sample would achieve 95\% recall}
+		\item \textbf{Additional Burden}: The proportion of the literature that need to be screened in order to \textit{be confident} that 95\% recall has been achieved, after 95\% of documents have actually been identified.
+	\end{itemize}
 	We operationalise the heuristic stopping criteria with 10, 20, and 50 irrelevant records in a row.
@@
 
-	\subsubsection*{Evaluation Data}
+	%\subsubsection*{Evaluation Data}
+
+	We use the collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}.
+This study was the first to apply machine-learning to classifying the relevance of documents in systematic reviews, and the underlying data was made public.
+It contains 15 sets of documents from comprehensive literature reviews of classes of drugs used in treating specific conditions.
+For each document, the original authors' decision to exclude or include at abstract and full-text level is given.
+We try to predict the relevance at abstract level.
 
@@
 	\bibliography{../mendeley}
-	\bibliographystyle{bmc_article}
+	\bibliographystyle{vancouver}
 

### 2019-07-22 analysis/rapid_review.py

@@ -16,6 +16,7 @@ import numpy as np
 import time, sys
 from IPython.display import clear_output
 import math
+import copy
 
 def lemmatize(token, tag):
     tag = {
@@ -113,6 +114,8 @@ class ScreenScenario:
         self.wss95_pf = None
         self.wss95_ih = None
         self.wss95_rs = None
+        for ih in self.irrelevant_heuristic:
+            setattr(self, f'wss95_ih_{ih}', None)
 
     def screen(self, i, rs=False):
         s = self.s
@@ -144,14 +147,17 @@ class ScreenScenario:
                     learning=False
                     break
                 x = self.X[index]
+                y = self.df.loc[index,'relevant']
                 y = self.df.query('seen==1')['relevant']
-                if len(set(y))<2:
+                last_iteration = self.ratings[-self.iteration_size:]
+                last_iteration_relevance = np.sum(last_iteration)/len(last_iteration)
+                if len(set(y))<2: # if we have a single class - just keep sampling
                     y_pred = [random.random() for x in unseen_index]
                 else:
                     clf.fit(x,y)
                     y_pred = clf.predict_proba(self.X[unseen_index])[:,1]
                     if rs:
-                        if max(y_pred) < 0.5:
+                        if max(y_pred) < 0.05 and last_iteration_relevance < 0.05:
                             r = self.sample_threshold()
                             return r
                 # These are the next documents
@@ -161,32 +167,35 @@ class ScreenScenario:
                     self.ratings.append(self.df.loc[i,'relevant'])
                     self.seen_docs = self.df.query('seen==1').shape[0]
                     self.r_seen = self.df.query('seen==1 & relevant==1').shape[0]
-                    self.recall_track.append(self.r_seen / self.r_docs)
+                    
+                    self.recall_track.append(self.get_recall())
                     self.work_track.append(self.seen_docs / self.N)
-                    last_ratings = self.ratings[-self.irrelevant_heuristic:]
-                    self.recall = self.r_seen / self.r_docs
-                    if len(last_ratings) == self.irrelevant_heuristic and np.sum(last_ratings)==0:
-                        if self.wss95_ih is None:
-                            self.wss95_ih = 1 - self.seen_docs / self.N
-                            self.recall_ih = self.r_seen / self.r_docs
[…]

### 2019-07-22 analysis/run_scenarios.py

@@ -40,7 +40,7 @@ cohen_db.head()
 models = [
     SVC(kernel='linear',C=5,probability=True)
 ]
-iterations = 5
+iterations = 10
 
 results = []
 rs_results = []
@@ -66,10 +66,14 @@ for name, group in cohen_db.groupby('review'):
             paths.append({
                 "dataset": name,
                 "work_path": ss.work_track,
-                "recall_path": ss.recall_track
+                "recall_path": ss.recall_track,
+                "estimated_recall_path": ss.estimated_recall_path
             })
 
+paths = [x for x in paths if x is not None]
+path_df = pd.DataFrame.from_dict(paths)
 results_df = pd.DataFrame.from_dict(results)
 rs_results_df = pd.DataFrame.from_dict(rs_results)
 rs_results_df.to_csv('../results/rs_results.csv', index=False)
 results_df.to_csv('../results/results.csv', index=False)
+paths_df.to_csv('../results/paths.csv', index=False)

### 2019-07-23 analysis/run_scenarios.py
new small results

@@ -40,7 +40,7 @@ cohen_db.head()
 models = [
     SVC(kernel='linear',C=5,probability=True)
 ]
-iterations = 10
+iterations = 20
 
 results = []
 rs_results = []
@@ -50,28 +50,30 @@ for name, group in cohen_db.groupby('review'):
         group,
         document_index,
     )
+    if df.shape[0] > 1000000:
+        continue
     df = df.dropna().reset_index(drop=True)
     for s in [200, 500]:
         ss = rr.ScreenScenario(
-            df, models, s, 50, name
+            df, models, s, [50, 100, 200], name
         )
         for i in range(iterations):
             print(i)
             r = ss.screen(i, True) 
             if r is not None:
                 rs_results.append(r)
-            r = ss.screen(i)
-            if r is not None:
-                results.append(r)
             paths.append({
                 "dataset": name,
                 "work_path": ss.work_track,
                 "recall_path": ss.recall_track,
                 "estimated_recall_path": ss.estimated_recall_path
             })
+            r = ss.screen(i)
+            if r is not None:
+                results.append(r)
 
 paths = [x for x in paths if x is not None]
-path_df = pd.DataFrame.from_dict(paths)
+paths_df = pd.DataFrame.from_dict(paths)
 results_df = pd.DataFrame.from_dict(results)
 rs_results_df = pd.DataFrame.from_dict(rs_results)
 rs_results_df.to_csv('../results/rs_results.csv', index=False)

### 2019-07-23 manuscript/main.tex
manuscript

@@
 However, reviewers are not given the opportunity to specify a confidence level, and for two of the four datasets in which they tested their criteria, the median achieved recall at a stopping criteria targeting 95\% recall was below 95\%.
+\cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold.
+These last examples are promising developments, but [are not ready for live reviews because hard to understand/communicate?? do not give uncertainty about reaching a threshold]
 
@@
 
+	We argue that the stopping criteria proposed here has several important advantages.
+
+	\begin{itemize}
+		\item The accuracy of results is independent of modelling decisions. When a model fits badly, you have to work more, but you don't miss results without knowing it.
+		\item A minimum threshold with uncertainty is understandable and communicable in a live review. It is based solely on arithmetic and simple statistics.
+		\item A minimum threshold better takes into account the undesirability of missed recall.
+	\end{itemize}
+
 \begin{backmatter}

### 2019-07-23 analysis/run_scenarios.py

@@ -53,7 +53,7 @@ for name, group in cohen_db.groupby('review'):
     df = df.dropna().reset_index(drop=True)
     for s in [200, 500]:
         ss = rr.ScreenScenario(
-            df, models, s, 50, name
+            df, models, s, [50, 100, 200], name
         )
         for i in range(iterations):
             print(i)

### 2019-07-23 manuscript/main.tex
add paragraph on information retrieval stopping criteria literature

--- 2019-07-22 manuscript/main.tex
@@
 
+	The information retrieval literature discusses similar stopping criteria for ranking algorithms like BM25 and variants \cite{DiNunzio2018, Yu2019}.
+However, the estimators they use to determine the recall rely on the specific ranking functions and depend on their search input.
+Therefore, the quality of the estimation depends on the adequacy of the model.
+Our approach, on the contrary, is model-free.
+
 	We evaluate this stopping criteria on real-world systematic review datasets on which active learning systems have previously been tested.

### 2019-07-23 analysis/rapid_review.py
improve ML performance a bit

@@ -17,6 +17,8 @@ import time, sys
 from IPython.display import clear_output
 import math
 import copy
+from sklearn.ensemble import IsolationForest
+from sklearn.neural_network import MLPClassifier
 
 def lemmatize(token, tag):
     tag = {
@@ -79,7 +81,8 @@ class ScreenScenario:
         self.s = sample
         self.irrelevant_heuristic = irrelevant_heuristic # how many irrelevant do we have to see in a row in order to stop
         self.results = []
-        self.iteration_size = 50
+        self.iteration_size = 20
+        self.iterations = 0
         self.recall_track = []
         self.work_track = []
         self.ratings = []
@@ -96,13 +99,17 @@ class ScreenScenario:
         self.unseen_p = None
         
         self.X = TfidfVectorizer(
-            ngram_range=(1,2),
-            min_df=5, max_df=0.6, strip_accents='unicode', 
+            ngram_range=(1,1),
+            min_df=2, max_df=0.9, strip_accents='unicode', 
             max_features=10000,
             use_idf=1,
             smooth_idf=1, sublinear_tf=1,
-            stop_words="english",tokenizer=tokenize
-        ).fit_transform(df['ab'])
+            #stop_words="english",tokenizer=tokenize
+        ).fit_transform(df['mesh'])
+
+        clf = IsolationForest()
+        clf.fit(self.X)
+        self.df['outlying'] = clf.predict(self.X)
 
     def reset(self):
         self.df['seen'] = 0
@@ -128,6 +135,8 @@ class ScreenScenario:
             print(f"skipping sample {s}, as it is more than 50% of the data")
             return
         sids = random.sample(list(self.df.index), s)
+        if rs:
+            sids = self.df.sort_values('outlying').index[:s]
         self.df.loc[sids,'seen'] = 1
         self.seen_docs = s
         self.r_seen = self.df.query('seen==1 & relevant==1').shape[0]
@@ -141,8 +150,9 @@ class ScreenScenario:
         for clf in self.models:
             learning = True
             while learning:
+                self.iterations +=1
                 clear_output(wait=True)
-                print(f"Dataset: {self.dataset}, iteration {self.iteration}.  {self.seen_docs} out of {self.N} documents seen")
+                print(f"Dataset: {self.dataset}, iteration {self.iteration}.  {self.seen_docs} out of {self.N} documents seen ({self.seen_docs/self.N:.0%}) - recall: {self.get_recall():.0%}")
                 index = self.df.query('seen==1').index
[…]

### 2019-07-24 analysis/rapid_review.py
forking paths

@@ -86,6 +86,8 @@ class ScreenScenario:
         self.recall_track = []
         self.work_track = []
         self.ratings = []
+        self.random_recall_track = []
+        self.random_work_track = []
         self.estimated_recall_path = []
         self.recall = None
         self.recall_pf = None
@@ -94,8 +96,6 @@ class ScreenScenario:
         self.wss95_bir = None # Work saved at 95\% recall
         self.wss95_bir_ci = None
         self.wss95_pf = None # Work saved at 95\% recall with perfect knowledge
-        self.wss95_ih = None
-        self.recall_ih = None
         self.unseen_p = None
         
         self.X = TfidfVectorizer(
@@ -117,6 +117,8 @@ class ScreenScenario:
         self.ratings = []
         self.recall_track = []
         self.work_track = []
+        self.random_recall_track = []
+        self.random_work_track = []
         self.estimated_recall_path = []
         self.wss95_bir = None
         self.wss95_bir_ci = None
@@ -135,8 +137,6 @@ class ScreenScenario:
             print(f"skipping sample {s}, as it is more than 50% of the data")
             return
         sids = random.sample(list(self.df.index), s)
-        if rs:
-            sids = self.df.sort_values('outlying').index[:s]
         self.df.loc[sids,'seen'] = 1
         self.seen_docs = s
         self.r_seen = self.df.query('seen==1 & relevant==1').shape[0]
@@ -145,6 +145,14 @@ class ScreenScenario:
         bir, ci = ci_ac(self.r_seen, self.seen_docs, 0.95)
         self.bir_upperbound = bir + ci
         self.r_predicted_upperbound = round(self.bir_upperbound*self.N)
+
+        outliers = False
+            if outliers:
+            self.df['seen'] = 0
+            sids = self.df.sort_values('outlying').index[:s]
+            self.df.loc[sids,'seen'] = 1
+            self.seen_docs = s
+            self.r_seen = self.df.query('seen==1 & relevant==1').shape[0]
 
         # Do some machine learning
         for clf in self.models:
@@ -168,26 +176,12 @@ class ScreenScenario:
                 else:
                     clf.fit(x,y)
                     y_pred = clf.predict_proba(self.X[unseen_index])[:,1]
-                    if rs and self.iterations > 2:
-                        # n = round((self.N - self.seen_docs)*0.1)
-                        # X = round(last_iteration_relevance*n)
-                        # print(last_iteration_relevance)
-                        # print(X)
-                        # print(n)
[…]

### 2019-07-24 analysis/run_scenarios.py

@@ -5,6 +5,11 @@ import os
 import matplotlib.pyplot as plt
 import importlib
 from sklearn.svm import SVC, OneClassSVM
+import argparse
+
+parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
+parser.add_argument('iterations', type=int)
+args = parser.parse_args()
 
 ##########################################
 ## Pull the document metadata from the xml files from the Pubmed API
@@ -40,7 +45,7 @@ cohen_db.head()
 models = [
     SVC(kernel='linear',C=5,probability=True)
 ]
-iterations = 20
+iterations = args.iterations
 
 results = []
 rs_results = []
@@ -59,23 +64,9 @@ for name, group in cohen_db.groupby('review'):
         )
         for i in range(iterations):
             print(i)
-            r = ss.screen(i, True) 
-            if r is not None:
-                rs_results.append(r)
-            paths.append({
-                "dataset": name,
-                "work_path": ss.work_track,
-                "recall_path": ss.recall_track,
-                "estimated_recall_path": ss.estimated_recall_path
-            })
-            r = ss.screen(i)
+            r = ss.screen(i, True)
             if r is not None:
                 results.append(r)
 
-paths = [x for x in paths if x is not None]
-paths_df = pd.DataFrame.from_dict(paths)
 results_df = pd.DataFrame.from_dict(results)
-rs_results_df = pd.DataFrame.from_dict(rs_results)
-rs_results_df.to_csv('../results/rs_results.csv', index=False)
 results_df.to_csv('../results/results.csv', index=False)
-paths_df.to_csv('../results/paths.csv', index=False)

### 2019-07-24 analysis/rapid_review.py
mistaken indent

@@ -147,7 +147,7 @@ class ScreenScenario:
         self.r_predicted_upperbound = round(self.bir_upperbound*self.N)
 
         outliers = False
-            if outliers:
+        if outliers:
             self.df['seen'] = 0
             sids = self.df.sort_values('outlying').index[:s]
             self.df.loc[sids,'seen'] = 1

### 2019-07-25 analysis/rapid_review.py
run on new data

@@ -105,7 +105,7 @@ class ScreenScenario:
             use_idf=1,
             smooth_idf=1, sublinear_tf=1,
             #stop_words="english",tokenizer=tokenize
-        ).fit_transform(df['mesh'])
+        ).fit_transform(df['x'])
 
         clf = IsolationForest()
         clf.fit(self.X)
@@ -160,7 +160,7 @@ class ScreenScenario:
             while learning:
                 self.iterations +=1
                 clear_output(wait=True)
-                print(f"Dataset: {self.dataset}, iteration {self.iteration}.  {self.seen_docs} out of {self.N} documents seen ({self.seen_docs/self.N:.0%}) - recall: {self.get_recall():.0%}")
+                print(f"Dataset: {self.dataset}, iteration {self.iteration}.  {self.seen_docs} out of {self.N} documents seen ({self.seen_docs/self.N:.0%}) - recall: {self.get_recall():.2%}")
                 index = self.df.query('seen==1').index
                 unseen_index = self.df.query('seen==0').index
                 if len(unseen_index) == 0:
@@ -169,21 +169,27 @@ class ScreenScenario:
                 x = self.X[index]
                 y = self.df.loc[index,'relevant']
                 y = self.df.query('seen==1')['relevant']
-                last_iteration = self.ratings[-self.iteration_size*4:]
+                n_last = self.iteration_size*4
+                if n_last > 0.05*self.N:
+                    n_last = int(round(0.05*self.N))
+                last_iteration = self.ratings[-n_last:]
                 last_iteration_relevance = np.sum(last_iteration)/len(last_iteration)
                 if len(set(y))<2: # if we have a single class - just keep sampling
-                    y_pred = [random.random() for x in unseen_index]
+                    y_pred = np.array([random.random() for x in unseen_index])
                 else:
-                    clf.fit(x,y)
-                    y_pred = clf.predict_proba(self.X[unseen_index])[:,1]
-                    if rs and self.iterations > 2 and self.random_work_track == []:                        
-                        if max(y_pred) < 0.02 and last_iteration_relevance < self.bir*0.5:
-                            self.last_iteration_relevance=last_iteration_relevance
-                            tdf = copy.deepcopy(self.df)
-                            r = self.sample_threshold()
-                            self.df = tdf
-                # These are the next documents
-                next_index = unseen_index[(-y_pred).argsort()[:self.iteration_size]]
[…]

### 2019-07-25 analysis/run_scenarios.py

@@ -43,7 +43,7 @@ cohen_db = cohen_db[["review","PMID","relevant"]]
 cohen_db.head()
 
 models = [
-    SVC(kernel='linear',C=5,probability=True)
+    SVC(kernel='linear',class_weight='balanced',probability=True)
 ]
 iterations = args.iterations
 
@@ -58,6 +58,7 @@ for name, group in cohen_db.groupby('review'):
     if df.shape[0] > 1000000:
         continue
     df = df.dropna().reset_index(drop=True)
+    df['x'] = df['mesh']
     for s in [200, 500]:
         ss = rr.ScreenScenario(
             df, models, s, [50, 100, 200], name

### 2019-07-25 analysis/run_scenarios_copd_pb.py

@@ -1,0 +1,66 @@
+import pandas as pd
+import numpy as np
+import rapid_review as rr
+import os
+import matplotlib.pyplot as plt
+import importlib
+from sklearn.svm import SVC, OneClassSVM
+import argparse
+
+parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
+parser.add_argument('iterations', type=int)
+args = parser.parse_args()
+
+importlib.reload(rr)
+
+dfs = []
+for path in [("ProtonBeam","all"),("COPD","copd")]:
+    document_index = rr.parse_pb_xml(f'../data/{path[0]}/{path[1]}.xml')
+    document_index = document_index.drop_duplicates()
+    document_index['rec-number'] = document_index['rec-number'].astype(int)
+    
+    relevant_index = pd.read_csv(
+    f'../data/{path[0]}/relevant.txt',header=None,
+    names=["rec-number"])
+    
+    relevant_index['relevant'] = 1
+    
+    df = pd.merge(
+        document_index,
+        relevant_index,
+        how="left"
+    )
+    
+    df['review'] = path[0]
+    dfs.append(df)
+    
+df = pd.concat(dfs)
+df['relevant'] = df['relevant'].fillna(0)
+
+importlib.reload(rr)
+
+results = []
+iterations = args.iterations
+
+df['x'] = df['ab']
+
+models = [
+    SVC(kernel='linear',class_weight='balanced',probability=True)
+]
+
+for name, group in df.groupby('review'):
+    group = group.dropna().reset_index(drop=True)
+    if group.shape[0] < 2000:
+        continue
+    for s in [200, 500]:
+        ss = rr.ScreenScenario(
+            group, models, s, [50, 100, 200], name
+        )
+        for i in range(iterations):
+            print(i)
+            r = ss.screen(i, True)
+            if r is not None:
+                results.append(r)
+
+results_df = pd.DataFrame.from_dict(results)
+results_df.to_csv('../results/results_pb_copd.csv', index=False)

### 2019-09-10 manuscript/main.tex
plots and writing

@@
 				Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. 
-				Machines and humans  work together, with the human providing training data, and the machine optimising the documents a human screens. 
-				However, current approaches lack robust stopping criteria, meaning that such systems are hard to implement in live reviews. 
+				Machines and humans  work together, with the human providing training data, and the machine optimising the documents a human screens, so that they can identify all relevant documents after viewing only a fraction of the total documents.
+				However, current approaches lack robust stopping criteria, so that reviewers do not know when they have seen all or a certain proportion of relevant documents.
+This means that such systems are hard to implement in live reviews.
 				This paper introduces a workflow for working with robust and flexible statistical stopping criteria, that offer real work reductions on the basis of a given confidence level of reaching a given recall.
+				The stopping criteria is shown on test datasets to achieve a reliable level of recall, while still providing consistent work reductions, while other methods proposed are shown to provide inconsistent recall and work reductions across datasets.
 
@@
 
-	Machine learning for evidence synthesis is a growing field which promises to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
-	A major strand of the literature works on screening,  
-	where several papers have developed or evaluated active learning. 
-	Active learning is an iterative process where documents screened by humans are used to train a machine learning model to predict the relevance of unseen papers.
-	The most relevant studies are rated by a human, which generates more labels to feed back to the machine. 
-	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or above a given threshold proportion of relevant studies - before having seen all the documents in the corpus.
-	The documents not yet seen by the human when they reach the given recall threshold have been represented as the work saved.
+	Machine learning for evidence synthesis is a growing field, where machine learning enabled interventions are introduced at various points in the workflow, in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
+	A major strand of the literature works on screening - the identification of relevant documents in a set of documents whose relevance is uncertain \cite{OMara-Eves2015}. 
+	For this task, several papers have developed or evaluated active learning.
 
-	In live review settings, however, the recall - or proportion of relevant studies screened - remains an unknown until all documents have been screened. 
+	Active learning is an iterative process where documents screened by humans are used to train a machine learning model to predict the relevance of unseen papers \cite{Settles2009}.
+	The most relevant studies are passed back to the human and rated, generating more labels to feed back to the machine. 
+	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (recall) - before having seen all the documents in the corpus.
+	The proportion of documents not yet seen by the human when they reach the given recall threshold are referred to as the work saved.
+
+	In live review settings, however, the recall remains unknown until all documents have been screened. 
 	In order for work to really be saved, reviewers have to stop screening while uncertain about the recall.
-	The lack of appropriate stopping criteria has been identified as a research gap \cite{bannach-brown2019}, and existing approaches have fallen into the following categories
+This is particularly problematic in systematic reviews because low recall increases the risk of publication bias %\cite{Lefebrve in Cochrane}
+	The lack of appropriate stopping criteria has therefore been identified as a research gap \cite{bannach-brown2019}, although some approaches have been suggested.
+These fall into the following categories
 	\begin{itemize}
@@
 
-	We show in this paper how first two criteria are inadequate.
-	This lack of adequate criteria challenges users of AL systems in live reviews, as authors are uncertain when to stop, and have no reliable way of reporting when they stopped and what that may say about the recall they reached.
+	We show in this paper how the first two criteria are inadequate.
+	We argue that the inadequacy lies in the unreliability - particularly across different domains, or datasets with different properties \todo{ref on cross-domain reliability} - both of the work saved and the recall achieved.
+Not achieving the desired level of recall is a particular challenge for systematic review screening.
+In systematic reviews, the larger the proportion of the relevant literature is not considered, the greater the risk of publication bias \todo{ref cochrane handbook}.
+	Without the reliable or reportable achievement of a desired level of recall, AL systems in live reviews remain challenging.
 
@@
 Therefore, the quality of the estimation depends on the adequacy of the model.
-Our approach, on the contrary, is model-free.
+Our approach, on the contrary, is independent of model choice or model performance.
 
@@
 		\item \textbf{Additional Burden}: The proportion of the literature that need to be screened in order to \textit{be confident} that 95\% recall has been achieved, after 95\% of documents have actually been identified.
+		\item \textbf{Missed Recall}: The amount by which, if any, the recall target was missed - in percentage points
 	\end{itemize}
-	We operationalise the heuristic stopping criteria with 10, 20, and 50 irrelevant records in a row.
+	We operationalise the heuristic stopping criteria with 50, 100, and 200 irrelevant records in a row.
 For simplicity, we use a basic SVM model [REFS], with 1-2 word n-grams in the document abstracts used as input data.
@@
 
+	As the intention is to evaluate stopping criteria, rather than machine learning approaches, we focus on additional burden and missed recall.
+These parameters correspond to the tradeoff inherent in stopping criteria: stopping too early means missing the recall target, stopping too late means undergoing additional burden than was actually required to meet the target.
+
 	%\subsubsection*{Evaluation Data}
 
-	We use the collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}.
-This study was the first to apply machine-learning to classifying the relevance of documents in systematic reviews, and the underlying data was made public.
-It contains 15 sets of documents from comprehensive literature reviews of classes of drugs used in treating specific conditions.
-For each document, the original authors' decision to exclude or include at abstract and full-text level is given.
-We try to predict the relevance at abstract level.
+	We use the collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}, along with PB COPD, and FASTREAD \todo{fill in gaps}.
+We feel that testing on datasets with different properties and from different domains is key to establishing criteria appropriate for general use.
+In this way we can show how well the criteria perform even when the model performs badly.
 
 	\section*{Results}
+
+	\begin{figure}
+		\includegraphics[width=\linewidth]{../images/burden_vs_misrecall}
+		\caption{Additional burden ($y > 0$) and missed recall ($ y < 0 $) for the range of stopping criteria tested.}
+		\label{additional-burden-missed-recall}
+	\end{figure}
+
+	\begin{figure}
+	\includegraphics[width=\linewidth]{../images/wss_recall_burden_boxplot}
+	\caption{Work saved, recall and additional burden across a variety of stopping criteria}
+	\label{additional-burden-missed-recall}
+\end{figure}
 

### 2019-09-10 analysis/rapid_review.py

@@ -136,7 +136,10 @@ class ScreenScenario:
         for ih in self.irrelevant_heuristic:
             setattr(self, f'wss95_ih_{ih}', None)
 
-    def screen(self, i, rs=False, nrs=True):
+    def screen(self, i, rs=False, nrs=True, alpha=0.95, recall_target=0.95):
+
+        self.recall_target = recall_target
+        self.alpha = alpha
         s = self.s
         self.iteration = i
         self.reset()
@@ -150,7 +153,7 @@ class ScreenScenario:
         self.r_seen = self.df.query('seen==1 & relevant==1').shape[0]
         self.bir = self.r_seen/self.seen_docs
         self.r_predicted = round(self.bir*self.N)
-        bir, ci = ci_ac(self.r_seen, self.seen_docs, 0.95)
+        bir, ci = ci_ac(self.r_seen, self.seen_docs, self.alpha)
         self.bir_upperbound = bir + ci
         self.r_predicted_upperbound = round(self.bir_upperbound*self.N)
 
@@ -203,11 +206,11 @@ class ScreenScenario:
                     clf.fit(x,y)
                     y_pred = clf.predict_proba(self.X[unseen_index])[:,1]
                     if rs and self.iterations > 2 and self.random_work_track == []:
-                        if (max(y_pred) < 0.2 and (self.max_prob_recall > 0.95 or last_n_relevant==0)) or self.seen_docs/self.N > 0.9:
+                        if (max(y_pred) < 0.2 and (self.max_prob_recall > self.recall_target or last_n_relevant==0)) or self.seen_docs/self.N > 0.2:
                             self.last_iteration_relevance=last_iteration_relevance
                             tdf = copy.deepcopy(self.df)
                             r = self.sample_threshold()
-                            
+                            break
                             self.df = tdf
                     # These are the next documents
                     next_index = unseen_index[(-y_pred).argsort()[:self.iteration_size]]
@@ -229,7 +232,7 @@ class ScreenScenario:
                     if self.r_seen > self.r_predicted_upperbound and self.wss95_bir_ci is None:
                         self.wss95_bir_ci = 1 - self.seen_docs / self.N
                         self.recall_bir_ci = self.get_recall()
-                    if self.get_recall() > 0.95 and self.wss95_pf is None:
+                    if self.get_recall() > self.recall_target and self.wss95_pf is None:
                         self.wss95_pf = 1 - self.seen_docs / self.N
                         self.recall_pf = self.get_recall()
[…]

### 2019-10-08 manuscript/main.tex
additional graphs
equations

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 \usepackage{url}
@@
 \subsection*{A Statistical Stopping Criterion for Active Learning}
+
+	\subsection*{Random Sampling}
+
+	Using the cumulative distribution function of the hypergeometric distribution, we can estimate the probability of observing $k$ or fewer relevant documents after drawing $N$ documents without replacement from a pool of $M$ documents, of which $n$ are relevant.
+
+	The probability mass function is defined as,
+
+	\begin{equation}	
+		 p(k, M, n, N) = \frac{\binom{n}{k} \binom{M - n}{N - k}} {\binom{M}{N}}	
+		 \label{pfunc}
+	\end{equation}
+
+	After seeing a proportion of the documents in a query, we can set $M$ to the number of documents remaining, and begin drawing documents at random from the remaining unseen documents.
+After each draw, $k$ is the number of relevant documents seen since sampling began, and $N$ is the number of documents sampled.
+We can generate a counterfactual $\hat{n}$, corresponding to the largest integer greater than the number of relevant documents seen $\rho$ divided by the recall target $\tau$ minus the relevant documents seen before random sampling began $\hat{\rho}$
+
+	\begin{equation}
+	\hat{n} = \lceil \frac{\rho}{\tau} - \hat{\rho} \rceil
+	\end{equation}
+
+	Substituting $\hat{n}$ into equation \ref{pfunc}, we can generate the $p$, the probability of observing $k$ or fewer documents if our recall target had been reached.
+If this is below $1-\alpha$, our confidence level, we can be $\alpha$ confident that our target has been reached.
+
+	\subsection*{Pseudo-random sampling}
+	In order to decide when to begin a random sample, we employ pseudo-random sampling, where we treat previously screened documents as a random sample.
+The distribution of relevant documents among previously screened documents is clearly not random, as documents predicted to be relevant are prioritised.
+It is reasonable to assume, though, that the distribution of relevant documents is greater among previously screened documents than among remaining unseen documents.
+This would make the following estimates conservative.
+
+	After reviewing each document, $S$ documents are screened, and $U$ documents are yet to be seen.
+We treat $i = 1 \dots S$ of the previously screened documents as a random sample, and calculate $p$ as before for each sample, taking the minimum across all samples $\dot{p}$.
+If $\dot{p}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling.
+We also calculate $\dot{p}$ for the remaining values as if we had not switched to random sampling and present these in the results below.
+
+	\medskip
+
+	\subsection*{Old description}
+
 	Figure \ref{flow} shows a workflow for the approach proposed in this paper. 
@@
 			\hline
-			$ N  $ &  Total number of studies & \textit{Observed} \\
+			$ P  $ &  Total number of studies & \textit{Observed} \\
 			$ S $ & Number of studies coded by humans & \textit{Observed} \\
-			$ U $ & Number of studies not yet coded by humans & \textit{Observed} ($N - S$) \\
-			$ p_S $ & relevance of documents rated by humans & \textit{Observed} \\
+			$ U $ & Number of studies not yet coded by humans & \textit{Observed} ($P - S$) \\
 			$ \alpha $ & Acceptable uncertainty level & \textit{Given} \\
-			$ \kappa $ & & \\
-			$ n $ & Number of randomly drawn documents & \textit{Observed} \\
-			$ \tilde{n} $ & & \\
-			$ \kappa $ & & \\
-			$ \hat{R} $ & Estimated recall & \\
-			$ \dot{R} $ & Minimum estimated recall at a given confidence threshold & \\
+			$ \tau $ & Target recall threshold & \textit{Given} \\
+			$ k $ & Number of relevant documents drawn & \\
+			$ N $ & Number of randomly drawn documents & \textit{Observed} \\
+			$ M $ & Number of remaining unseen documents before sampling & \\
+			$ n $ & Number of relevant documents in sample & \\
+			$ \hat{n} $ & Minimum number of relevant documents in sample that would mean the recall threshold had been reached  & \\
 		\end{tabular}
@@
 		\caption{Work saved, recall and additional burden across a variety of stopping criteria}
-	\label{additional-burden-missed-recall}
+		\label{additional-burden-missed-recall-boxplots}
+	\end{figure}
+
+	\begin{figure}
+	\includegraphics[width=\linewidth]{../images/results_hex}
+	\caption{Work saved, recall across a variety of stopping criteria}
+	\label{results-hex}
+\end{figure}
+
+	\begin{figure}
+	\includegraphics[width=\linewidth]{../images/ab_recall_studies}
+	\caption{Work saved, recall across a variety of stopping criteria}
+	\label{ab_recall_studies}
 \end{figure}

### 2019-10-08 analysis/rapid_review.py

@@ -82,7 +82,7 @@ class ScreenScenario:
         self.s = sample
         self.irrelevant_heuristic = irrelevant_heuristic # how many irrelevant do we have to see in a row in order to stop
         self.results = []
-        self.iteration_size = 20
+        self.iteration_size = 40
         self.iterations = 0
         self.recall_track = []
         self.work_track = []
@@ -94,9 +94,9 @@ class ScreenScenario:
         self.recall_pf = None
         self.recall_bir = None
         self.recall_bir_ci = None
-        self.wss95_bir = None # Work saved at 95\% recall
-        self.wss95_bir_ci = None
-        self.wss95_pf = None # Work saved at 95\% recall with perfect knowledge
+        self.wss_bir = None # Work saved 
+        self.wss_bir_ci = None
+        self.wss_pf = None # Work saved
         self.unseen_p = None
         
         self.X = TfidfVectorizer(
@@ -121,23 +121,32 @@ class ScreenScenario:
         self.random_recall_track = []
         self.random_work_track = []
         self.estimated_recall_path = []
-        self.estimated_p_path = []
         self.hyper_hypo_path = []
         self.prob_recall_path = []
         self.max_min_recall_path= []
-        self.wss95_bir = None
-        self.wss95_bir_ci = None
-        self.wss95_pf = None
-        self.wss95_ih = None
-        self.wss95_rs = None
-        self.wss95_nrs = None
-        self.wss95_hyper = None
+        self.prob_target_path = []
+        self.nr_prob_target_path = []
+        self.prob_target = None
+        self.wss_bir = None
+        self.wss_bir_ci = None
+        self.wss_pf = None
+        self.wss_ih = None
+        self.wss_rs = None
+        self.wss_nrs = None
+        self.wss_hyper = None
         self.max_prob_recall = 0
+        self.min_prob_target = 1
         for ih in self.irrelevant_heuristic:
-            setattr(self, f'wss95_ih_{ih}', None)
+            setattr(self, f'wss_ih_{ih}', None)
 
-    def screen(self, i, rs=False, nrs=True, alpha=0.95, recall_target=0.95):
 
+    def screen(self, i, rs=False, nrs=True, alpha=0.95, recall_target=0.95, max_switch=0.8):
+
+        def gen_hypo(X, r_seen, r_target):
+            return math.ceil(r_seen/r_target+0.01) - r_seen + X
+
+        vgen_hypo = np.vectorize(gen_hypo)
+        
[…]

### 2019-10-14 manuscript/main.tex
alternative version of stopping criterion description

@@
 
-	Using the cumulative distribution function of the hypergeometric distribution, we can estimate the probability of observing $k$ or fewer relevant documents after drawing $N$ documents without replacement from a pool of $M$ documents, of which $n$ are relevant.
-
-	The probability mass function is defined as,
-
+	Our aim is to formulate a stopping criterion such that a recall of $\tau$ has been reached with confidence level $\alpha$.
+	To achieve this, we formulate a hypothesis test:
+	We first calculate the number $\hat{n}$ of relevant documents that can maximally remain in the unseen documents to achieve a given recall target.
+The recall is defined as the ratio of seen relevant documents over all relevant documents
 	\begin{equation}
-		 p(k, M, n, N) = \frac{\binom{n}{k} \binom{M - n}{N - k}} {\binom{M}{N}}	
-		 \label{pfunc}
+		\tau = \frac{\hat{\rho}}{\rho + \tilde{n}}.
 	\end{equation}
-
-	After seeing a proportion of the documents in a query, we can set $M$ to the number of documents remaining, and begin drawing documents at random from the remaining unseen documents.
-After each draw, $k$ is the number of relevant documents seen since sampling began, and $N$ is the number of documents sampled.
-We can generate a counterfactual $\hat{n}$, corresponding to the largest integer greater than the number of relevant documents seen $\rho$ divided by the recall target $\tau$ minus the relevant documents seen before random sampling began $\hat{\rho}$
-
+	Reorganization of this equation gives
+	\begin{equation}
+		\tilde{n} = \frac{\hat{\rho}}{\tau} - \rho.
+	\end{equation}
+	Because we need to use integers and want a conservative estimation, we are using the next largest integer given by the ceiling function:
 	\begin{equation}
@@
 	\end{equation}
+	The null-hypothesis to reject can then be formulated as:
+	\begin{quote}
+	$H_0$: There are $\hat{n}$ or more relevant documents remaining in the set of unseen documents.
+	\end{quote}
 
-	Substituting $\hat{n}$ into equation \ref{pfunc}, we can generate the $p$, the probability of observing $k$ or fewer documents if our recall target had been reached.
-If this is below $1-\alpha$, our confidence level, we can be $\alpha$ confident that our target has been reached.
+	% notation here follows notation for scipy description of hypergeometric distribution
+	% better use another notation more in line with the literature? -> literature seems to use very different notations
+
+	In theory, we draw a random sample of $N$ documents, evaluate how many of them are relevant, and denote this number with $k$.
+We can estimate the probability that we observe $k$ or fewer relevant documents in a random draw of $N$ documents without replacement from a pool of $M$ documents, of which $n$ are relevant using the cumulative distribution function of the hypergeometric distribution.
+This function is defined as the sum over the probability mass function of the hypergeometric distribution $p(x, M, n, N)$:
+	\begin{equation}	
+		 cdf(k, M, n, N) = \sum_{x \leq k} p(x, M, n, N) = \sum_{x \leq k} \frac{\binom{n}{x} \binom{M - n}{N - x}} {\binom{M}{N}},
+		 \label{pfunc}
+	\end{equation}
+	where $\binom{a}{b}$ denote binomial coefficients. In our case, $M$ is the number of unseen documents before we start random sampling.
+
+	We can reject the null hypothesis if the p-value $p = cdf(k,M,\hat{n},N)$ is smaller than $1 - \alpha$ [references to hypergeometric testing?, needs further argumentation?].
+Because we are testing for the lower tail, this implies that we can be confident with a level $\alpha$ that the number of unseen relevant documents is lower than $\hat{n}$.
+
+	In practice, we can draw random unseen documents repeatedly. After each draw, we calculate the p-value using the number of relevant documents $k$ and the number of randomly sampled documents $N$.
+	We can stop the screening once a p-value $p < 1 - \alpha$ is reached.
 
@@
 The distribution of relevant documents among previously screened documents is clearly not random, as documents predicted to be relevant are prioritised.
-It is reasonable to assume, though, that the distribution of relevant documents is greater among previously screened documents than among remaining unseen documents.
+It is reasonable to assume, though, that the density of relevant documents is greater among previously screened documents than among remaining unseen documents.
 This would make the following estimates conservative.
@@
 	After reviewing each document, $S$ documents are screened, and $U$ documents are yet to be seen.
-We treat $i = 1 \dots S$ of the previously screened documents as a random sample, and calculate $p$ as before for each sample, taking the minimum across all samples $\dot{p}$.
-If $\dot{p}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling.
-We also calculate $\dot{p}$ for the remaining values as if we had not switched to random sampling and present these in the results below.
+We treat $i = 1 \dots S$ of the previously screened documents as a random sample, and calculate $p$ as before for each sample, taking the minimum across all samples $p_{min}$.
+If $p_{min}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling.
+We also calculate $p_{min}$ for the remaining values as if we had not switched to random sampling and present these in the results below.
 

### 2019-10-14 manuscript/main.tex
update minor mistakes in stopping criterion description

@@
 	\begin{equation}
-		\tau = \frac{\hat{\rho}}{\rho + \tilde{n}}.
+		\tau = \frac{\rho}{\hat{\rho} + \tilde{n}},
 	\end{equation}
+	with the number total number of seen relevant documents $\rho$, the number of seen relevant documents at the time we start random sampling $\hat{\rho}$, and the remaining number of relevant documents in the unseen sample at the time random sampling starts $\tilde{n}$.
 	Reorganization of this equation gives
 	\begin{equation}
-		\tilde{n} = \frac{\hat{\rho}}{\tau} - \rho.
+		\tilde{n} = \frac{\rho}{\tau} - \hat{\rho}.
 	\end{equation}
-	Because we need to use integers and want a conservative estimation, we are using the next largest integer given by the ceiling function:
+	Because we need to use integers, we get the next largest integer given by the ceiling function $\lceil \cdot \rceil$:
 	\begin{equation}

### 2019-10-17 manuscript/main.tex
jointplots

--- 2019-10-08 manuscript/main.tex
@@
 \usepackage{url}
+\usepackage{subcaption}
 %\def\includegraphic{}
@@
 
-	\begin{figure}
-		\includegraphics[width=\linewidth]{../images/additional_burden}
-		\caption{Additional burden for 20 random samples where the 95\% threshold has been reached, and the remaining documents are 1\%, 2\% or 5\% relevant, for different total numbers of documents.}
-		\label{additional-burden}
-	\end{figure}
-
 	When to move to random sampling remains an open question.
@@
 
-	\begin{figure}
-		\includegraphics[width=\linewidth]{../images/burden_vs_misrecall}
-		\caption{Additional burden ($y > 0$) and missed recall ($ y < 0 $) for the range of stopping criteria tested.}
-		\label{additional-burden-missed-recall}
-	\end{figure}
-
-	\begin{figure}
-		\includegraphics[width=\linewidth]{../images/wss_recall_burden_boxplot}
-		\caption{Work saved, recall and additional burden across a variety of stopping criteria}
-		\label{additional-burden-missed-recall-boxplots}
-	\end{figure}
-
-	\begin{figure}
-	\includegraphics[width=\linewidth]{../images/results_hex}
-	\caption{Work saved, recall across a variety of stopping criteria}
-	\label{results-hex}
-\end{figure}
-
-	\begin{figure}
-	\includegraphics[width=\linewidth]{../images/ab_recall_studies}
-	\caption{Work saved, recall across a variety of stopping criteria}
-	\label{ab_recall_studies}
-\end{figure}
+    \begin{figure*}
+	\centering
+	\begin{subfigure}[b]{0.475\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{../images/jointplot_bir}
+		\caption[Network2]%Example-Image
+		{{\small Baseline Sampling}}    
+		\label{fig:mean and std of net14}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}  
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_ih_50}
+		\caption[]%
+		{{\small 50 consecutive irrelevant results}}    
+		\label{fig:mean and std of net24}
+	\end{subfigure}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_hyper}
+		\caption[]%
+		{{\small Hypergeometric sampling \\}}    
+		\label{fig:mean and std of net34}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_nrs}
+		\caption[]%
+		{{\footnotesize Pseudorandom hypergeometric sampling}}    
+		\label{fig:mean and std of net44}
+	\end{subfigure}
+	\caption[ The average and standard deviation of critical parameters ]
+	{\small The average and standard deviation of critical parameters: Region R4} 
+	\label{fig:mean and std of nets}
+\end{figure*}
 

### 2019-10-17 analysis/rapid_review.py

@@ -55,6 +55,7 @@ def ci_ac(X, n, a):
     X_tilde = X + k**2/2
     n_tilde = n + k**2
     p_tilde = X_tilde / n_tilde
+    
     q_tilde = 1-p_tilde
 
     ci = k* np.sqrt(p_tilde*q_tilde/n_tilde) 
@@ -82,7 +83,7 @@ class ScreenScenario:
         self.s = sample
         self.irrelevant_heuristic = irrelevant_heuristic # how many irrelevant do we have to see in a row in order to stop
         self.results = []
-        self.iteration_size = 40
+        self.iteration_size = 20
         self.iterations = 0
         self.recall_track = []
         self.work_track = []
@@ -252,9 +253,6 @@ class ScreenScenario:
                     if self.r_seen > self.r_predicted and self.wss_bir is None:
                         self.wss_bir = 1 - self.seen_docs / self.N
                         self.recall_bir = self.get_recall()
-                    X = 0
-                    max_min_recall = 0
-                    self.max_prob_recall = 0
                     if nrs is True:# (self.wss_nrs is None or self.wss_hyper is None) and nrs is True:
                         self.n_remaining=self.N - self.seen_docs
                         Xs = np.cumsum(np.array(self.ratings[::-1]))

### 2019-10-22 images/flow.tex
parallelised

@@
 
-	\node (sform) [form, right=of screenr2, xshift=-15mm] { Estimate $\hat{p}U$ };
+    	\node (e1) [form, right=of stopscreen, xshift=-15mm] { $p < 1-\frac{\alpha}{2}$ };
 
-  \node (rform) [form, right=of recall, xshift=-15mm] {$ 1 - \frac{\hat{p}U}{\hat{p}P} \frac{U}{P}$};
+	%\node (sform) [form, right=of screenr2, xshift=-15mm] { Estimate $\hat{p}U$ };
+
+  \node (rform) [form, right=of recall, xshift=-15mm] { $p < 1-\alpha $ };
 

### 2019-10-22 manuscript/main.tex
parallelised

@@
 
-	Machine learning for evidence synthesis is a growing field, where machine learning enabled interventions are introduced at various points in the workflow, in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
+	Machine learning for evidence synthesis is an rapidly emerging field [cite].
+	Machine learning enabled interventions have been proposed at various points in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
 	A major strand of the literature works on screening - the identification of relevant documents in a set of documents whose relevance is uncertain \cite{OMara-Eves2015}. 
@@
 	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (recall) - before having seen all the documents in the corpus.
-	The proportion of documents not yet seen by the human when they reach the given recall threshold are referred to as the work saved.
+	The proportion of documents not yet seen by the human when they reach the given recall threshold are referred to as the work saved, as this represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.
 
@@
 	We argue that the inadequacy lies in the unreliability - particularly across different domains, or datasets with different properties \todo{ref on cross-domain reliability} - both of the work saved and the recall achieved.
-Not achieving the desired level of recall is a particular challenge for systematic review screening.
-In systematic reviews, the larger the proportion of the relevant literature is not considered, the greater the risk of publication bias \todo{ref cochrane handbook}.
 	Without the reliable or reportable achievement of a desired level of recall, AL systems in live reviews remain challenging.
@@
 	This study proposes a system for estimating the recall based on random sampling of remaining documents. 
-	We draw on the literature on binomial distributions \cite{Brown2001}, to illustrate how AL users can predefine a threshold in terms of uncertainty and recall, and use this to transparently save work with machine learning, while making a statement like ``There is a <5\% chance that we achieve a recall under 95\%''.
+	We use basic statistics to test the hypothesis that less than a given recall threshold has been achieved.
+	This allows AL users to predefine a threshold in terms of uncertainty and recall, so that they can make transparent, understandable statements like ``There is a <5\% chance that we achieve a recall under 95\%''.
 
@@
 
-	We start by explaining the sampling and heuristic based stopping criteria before showing with toy data how the criteria fall short. 
-	Then we introduce our own suggested stopping criteria, and show its benefits with toy data, before testing all criteria on real world datasets.
+	We start by explaining the sampling and heuristic based stopping criteria and showing how their methodological weaknesses. 
+	Then we introduce our own suggested stopping criteria before testing all criteria on real world datasets.
 
@@
 
-	However, the estimation of the BIR fails to take into account the sampling uncertainty. 
+	However, the estimation of the BIR fails to take into account sampling uncertainty. 
 	Figure \ref{bir-sampling}.a shows for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant), the predicted number of documents after each document drawn. 
@@
 	\begin{itemize}
-		\item much higher than the true value, meaning that the stopping criterion would never be reached - and no work could be saved
+		\item much higher than the true value, meaning that the stopping criterion would never be reached - and no work could be saved, or;
 		\item much lower than the true value, meaning that the stopping criterion would be reached before the desired level of recall was actually achieved
@@
 
-	We can estimate the uncertainty of a binomial outcome using an Agresti-Coull confidence interval (which has been shown to be conservative at levels near 0 \todo{this is less relevant here, go back to literature and see if another is better} \cite{Brown2001}).
-	Figure \ref{bir-sampling}.b shows the 95\% confidence interval (brown) and its upper bound (green) in the same set of random draws. 
-	Where the green line is above the above the lower (upper) grey line, identification of 100\% (95\%) of the estimated number of studies would be impossible, and no work would be saved.
-
-	Figure \ref{bir-variation} explores different scenarios of true relevance and probability thresholds, showing that the higher the true proportion of relevant documents, the fewer documents one would have to sample to make the identification of a given fraction (the true number of relevant results divided by the upper estimate) possible.
-	\todo{Why does it behave like this, is there a theoretical maximum for each? something to do with absolute values? a Bug? Is uncertainty a function of being near to 0?} 
-	Even with a relatively high proportion of relevant documents, and a sample of 500, all relevant documents would have to be identified to have a 95\% likelihood of achieving 95\% recall. 
-	With low proportions of relevant documents, only an 80\% recall can be estimated with confidence, even after a sample of 2000, and subsequently identifying all relevant documents. 
-	N.B, these results ignore the random element of sample order, which explains the variation on figure \ref{bir-sampling}. 
-	They show upper confidence bounds for accurate estimates of relevance.
+	Where the yellow line is above the above the upper grey line, identification of 95\% of the estimated number of studies would be impossible, and no work would be saved.
+Where it is below, true recall, where recall is estimated to be 95\% would be less than 95\%.
 
 	\begin{figure}
-	\includegraphics[width=\linewidth]{../images/bir_sampling.png}
+	\includegraphics[width=\linewidth]{../images/bir_sampling_noci.png}
 	\caption{Uncertainty in estimating the baseline inclusion rate.
-a.
-shows the estimated BIR after drawing each document, for 50 random samples of from 20,000 documents, after each document is drawn.
-b.
-Shows the 95\% Confidence interval (brown) and the upper bound (green) for each of the same samples.
+For 50 random samples of 2,000 documents from 20,000 documents, we show the estimated number of relevant documents, after each document is drawn.
 The two lines correspond to 100\% of relevant documents, and $\frac{100}{0.95}$\% of relevant documents.}
-	\label{bir-sampling}
-\end{figure}
+	\label{bir-sampling}.
 
-\begin{figure}
-	\includegraphics[width=0.5\linewidth]{../images/bir_variation.png}
-	\caption{Maximum indentifiable fraction of relevant results under different levels of true relevance $p$ and confidence thresholds $\alpha$.
-Maximum identifiable fraction is the true number of relevant documents divided by the upper estimate.}
-	\label{bir-variation}
+	This shows how even after taking a large sample, baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
 \end{figure}
@@
 Figure \ref{unseen-proportions} illustrates this second point. 
-We show two scenarios with identical low proportions of relevant documents (note that estimating a proportion lower than 0.02 with 95\% confidence requires 160 consecutive irrelevant documents)\todo{demonstrate?}.
+We show two scenarios with identical low proportions of relevant documents observed in the unseen documents. 
 Recall is not simply a function of the relevance of unseen documents, but also of the number of unseen documents. 
 This also means that where machine learning has performed well (as in the top figure), low proportions of irrelevant documents in those that remain are indicative of lower recall than where ML has performed less well.
+Likewise, where the proportion of relevant documents is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low.
 
@@
 \cite{Jonnalagadda2013} note that ``the reviewer can elect to end the process of classifying documents at any point, recognizing that stopping before reviewing all documents involves a trade-off of lower recall for reduced workload'', although clearly the reviewer lacks information about probable recall.
-\cite{Yu2019} adopt a more complicated stopping criterion \todo{How does it actually work??} which allows the user to target a specific level of recall.
+\cite{Yu2019} adopt a more complicated stopping criterion which allows the user to target a specific level of recall.
 However, reviewers are not given the opportunity to specify a confidence level, and for two of the four datasets in which they tested their criteria, the median achieved recall at a stopping criteria targeting 95\% recall was below 95\%.
 \cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold.
-These last examples are promising developments, but [are not ready for live reviews because hard to understand/communicate?? do not give uncertainty about reaching a threshold]
+These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount.
 
@@
 	\end{equation}
-	with the number total number of seen relevant documents $\rho$, the number of seen relevant documents at the time we start random sampling $\hat{\rho}$, and the remaining number of relevant documents in the unseen sample at the time random sampling starts $\tilde{n}$.
+	with the total number of seen relevant documents $\rho$, the number of seen relevant documents at the time we start random sampling $\hat{\rho}$, and the remaining number of relevant documents in the unseen sample at the time random sampling starts $\tilde{n}$.
 	Reorganization of this equation gives
@@
 
-	\subsection*{Old description}
-
-	Figure \ref{flow} shows a workflow for the approach proposed in this paper. 
-	The random sampling occurs at the end of the process, and is used to estimate the number of relevant documents remaining and the total number of relevant documents. 
-	The rationale is to limit the uncertainty to a subsection of the dataset: that which has not yet been screened. 
-	As reviewers continue to draw random documents, the uncertainty range decreases, and the proportion of the data about which one is uncertain also decreases.
-
 	\begin{figure}
@@
 
-	\begin{table}
-		\begin{tabular}{p{0.15\linewidth} p{0.425\linewidth} p{0.425\linewidth}}
-			Parameter & Description & Estimation \\
-			\hline
-			$ P  $ &  Total number of studies & \textit{Observed} \\
-			$ S $ & Number of studies coded by humans & \textit{Observed} \\
-			$ U $ & Number of studies not yet coded by humans & \textit{Observed} ($P - S$) \\
-			$ \alpha $ & Acceptable uncertainty level & \textit{Given} \\
-			$ \tau $ & Target recall threshold & \textit{Given} \\
-			$ k $ & Number of relevant documents drawn & \\
-			$ N $ & Number of randomly drawn documents & \textit{Observed} \\
-			$ M $ & Number of remaining unseen documents before sampling & \\
-			$ n $ & Number of relevant documents in sample & \\
-			$ \hat{n} $ & Minimum number of relevant documents in sample that would mean the recall threshold had been reached  & \\
-		\end{tabular}
-		\caption{Parameters}
-		\label{parameters}
-	\end{table}
+	\subsection*{Analysis}
+%	
+%	\begin{table}
+%		\begin{tabular}{p{0.15\linewidth} p{0.425\linewidth} p{0.425\linewidth}}
+%			Parameter & Description & Estimation \\
+%			\hline
+%			$ P  $ &  Total number of studies & \textit{Observed} \\
+%			$ S $ & Number of studies coded by humans & \textit{Observed} \\
+%			$ U $ & Number of studies not yet coded by humans & \textit{Observed} ($P - S$) \\
+%			$ \alpha $ & Acceptable uncertainty level & \textit{Given} \\
+%			$ \tau $ & Target recall threshold & \textit{Given} \\
+%			$ k $ & Number of relevant documents drawn & \\
+%			$ N $ & Number of randomly drawn documents & \textit{Observed} \\
+%			$ M $ & Number of remaining unseen documents before sampling & \\
+%			$ n $ & Number of relevant documents in sample & \\
+%			$ \hat{n} $ & Minimum number of relevant documents in sample that would mean the recall threshold had been reached  & \\
+%		\end{tabular}
+%		\caption{Parameters}
+%		\label{parameters}
+%	\end{table}
 
-	Table \ref{parameters} shows the parameters, known, estimated, and given, available during the random sampling process and required to estimate recall. 
-	Expanding on what was stated before, recall - at a given confidence threshold is a function of 1) the upper estimate of the relevance of remaining documents, 2) The estimated relevance of all documents in the dataset, 3) The proportion of documents not yet seen.
-
-	Figure \ref{sample-recall} shows the minimum estimated recall for a set of confidence intervals, along with the actual recall (in grey) for a case where 800 out of 2,000 documents have been reviewed, 2\% of remaining documents are relevant, and 20\% of all documents are relevant.
-	We see that after 95\% recall has actually been achieved, but before 100\% of documents have been been seen, we can be confident at each of the given confidence levels that 95\% of relevant documents have been identified.
-	All estimates of recall only reach 100\% after all documents have been seen, as it is not possible to exclude the possibility, at any given confidence level, that the proportion of relevant documents is greater than 0 \todo{Actually maybe this is possible if we work with whole numbers?
+%
+%	
+%	Table \ref{parameters} shows the parameters, known, estimated, and given, available during the random sampling process and required to estimate recall. 
+%	Expanding on what was stated before, recall - at a given confidence threshold is a function of 1) the upper estimate of the relevance of remaining documents, 2) The estimated relevance of all documents in the dataset, 3) The proportion of documents not yet seen.
+%	
+%	Figure \ref{sample-recall} shows the minimum estimated recall for a set of confidence intervals, along with the actual recall (in grey) for a case where 800 out of 2,000 documents have been reviewed, 2\% of remaining documents are relevant, and 20\% of all documents are relevant.
+%	We see that after 95\% recall has actually been achieved, but before 100\% of documents have been been seen, we can be confident at each of the given confidence levels that 95\% of relevant documents have been identified.
+%	All estimates of recall only reach 100\% after all documents have been seen, as it is not possible to exclude the possibility, at any given confidence level, that the proportion of relevant documents is greater than 0 \todo{Actually maybe this is possible if we work with whole numbers?
 How does that change things?}.
-	Figure \ref{sample-recall-variation} shows the various trajectories of a 95\% minimum recall (blue) and actual recall (grey) given the same paramaters for 200 random samples.
-
-	\begin{figure}
-		\includegraphics[width=\linewidth]{../images/sample_recall}
-		\caption{Sample recall}
-		\label{sample-recall}
-	\end{figure}
-
-	\begin{figure}
-		\includegraphics[width=\linewidth]{../images/sample_recall_2}
-		\caption{Sample recall variation}
-		\label{sample-recall-variation}
-	\end{figure}
-
-	That the blue lines consistently meet a 95\% threshold to the right of the grey lines indicates that reviewers have to see more than 95\% of relevant documents in order to be confident that they have achieved their threshold.
-	We call the number of documents it is necessary to review to establish with confidence that reviewers have achieved 95\% recall after they have already achieved 95\% recall the \textit{additional burden}.
-
-	In figure \ref{additional-burden}, we investigate how the total number of documents and the proportion of relevant documents affect the additional burden \todo{Use the number of documents seen instead of p?}.
+%	Figure \ref{sample-recall-variation} shows the various trajectories of a 95\% minimum recall (blue) and actual recall (grey) given the same paramaters for 200 random samples.
+%
+%	That the blue lines consistently meet a 95\% threshold to the right of the grey lines indicates that reviewers have to see more than 95\% of relevant documents in order to be confident that they have achieved their threshold.
+%	We call the number of documents it is necessary to review to establish with confidence that reviewers have achieved 95\% recall after they have already achieved 95\% recall the \textit{additional burden}.
+%	
+%	In figure \ref{additional-burden}, we investigate how the total number of documents and the proportion of relevant documents affect the additional burden \todo{Use the number of documents seen instead of p?}.
 We demonstrate the additional burden decreases \todo{how?
@@
 This is because the proportion of documents not seen is greater, so uncertainty around the relevance of that proportion has a greater impact on the estimation of recall.
-
-	When to move to random sampling remains an open question.
+%
+%	When to move to random sampling remains an open question.
 Our current approach is to use the predicted relevance of the machine learning model as a guide.
@@
 If the estimated recall would be below the threshold, then we begin drawing random documents.
-
-	The stopping criteria explained here is
+%
+%	The stopping criteria explained here is
 
@@
 For simplicity, we use a basic SVM model [REFS], with 1-2 word n-grams in the document abstracts used as input data.
-We start with random samples of [100, 200, and 500] documents.
+We start with random samples of [200 and 500] documents.
 
 	As the intention is to evaluate stopping criteria, rather than machine learning approaches, we focus on additional burden and missed recall.
-These parameters correspond to the tradeoff inherent in stopping criteria: stopping too early means missing the recall target, stopping too late means undergoing additional burden than was actually required to meet the target.
+These parameters correspond to the tradeoff inherent in stopping criteria: stopping too early means missing the recall target, stopping too late means screening more documents than was actually required to meet the target.
 
@@
 
-	We use the collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}, along with PB COPD, and FASTREAD \todo{fill in gaps}.
+	We use the collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}, along with PB COPD, and FASTREAD [CITE].
 We feel that testing on datasets with different properties and from different domains is key to establishing criteria appropriate for general use.
@@
 	\section*{Results}
+
+	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criteria has been satisfied. 
+	For comparison, we also include the results that would have been achieved with perfect foresight, that is, the work saved when the 95\% recall target has been received.
+
+	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in [33]\% of cases, while the most common work saving is 0\%.
+This is in line with out expectations, that due to random sampling error, the expected number of document will often be over-estimated or under-estimated, resulting in poor work savings or poor recall.
+
+	The Heuristic based stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
+Although the
 
@@
 		{{\small Baseline Sampling}}    
-		\label{fig:mean and std of net14}
+		\label{fig:bir}
 	\end{subfigure}
@@
 		{{\small 50 consecutive irrelevant results}}    
-		\label{fig:mean and std of net24}
+		\label{fig:ih_50}
 	\end{subfigure}
@@
 		{{\small Hypergeometric sampling \\}}    
-		\label{fig:mean and std of net34}
+		\label{fig:hyper}
 	\end{subfigure}
@@
 		{{\footnotesize Pseudorandom hypergeometric sampling}}    
-		\label{fig:mean and std of net44}
+		\label{fig:nrs}
 	\end{subfigure}
-	\caption[ The average and standard deviation of critical parameters ]
-	{\small The average and standard deviation of critical parameters: Region R4} 
-	\label{fig:mean and std of nets}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_ih_200}
+		\caption[]%
+		{{\small 200 consecutive irrelevant results \\}}    
+		\label{fig:ih_200}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_pf}
+		\caption[]%
+		{{\footnotesize Perfect Foresight}}    
+		\label{fig:pf}
+	\end{subfigure}
+
+	\caption{\small Distribution of recall and work saved after each stopping criteria.} 
+	\label{recall-wss}
 \end{figure*}

### 2019-10-22 analysis/rapid_review.py

@@ -196,7 +196,10 @@ class ScreenScenario:
                 if n_last > 0.05*self.N:
                     n_last = int(round(0.05*self.N))
                 last_iteration = self.ratings[-n_last:]
-                last_iteration_relevance = np.sum(last_iteration)/len(last_iteration)
+                if len(last_iteration) < 1:
+                    last_iteration_relevance = 1
+                else:
+                    last_iteration_relevance = np.sum(last_iteration)/len(last_iteration)
                 n_last = 386
                 if self.N*0.1 > n_last:
                     n_last = round(self.N*0.1)
@@ -254,17 +257,18 @@ class ScreenScenario:
                         self.wss_bir = 1 - self.seen_docs / self.N
                         self.recall_bir = self.get_recall()
                     if nrs is True:# (self.wss_nrs is None or self.wss_hyper is None) and nrs is True:
-                        self.n_remaining=self.N - self.seen_docs
-                        Xs = np.cumsum(np.array(self.ratings[::-1]))
-                        ns = np.arange(len(self.ratings))
-                        hypotheticals = vgen_hypo(Xs, self.r_seen, self.recall_target)
-                        prob_target = hypergeom.cdf(Xs, self.n_remaining+ns, hypotheticals, ns+1)
-                        self.min_prob_target = prob_target.min()
-                        
-                        if self.min_prob_target < 1-self.alpha and self.wss_nrs is None:
-                            self.wss_nrs = 1 - self.seen_docs / self.N
-                            self.recall_nrs = self.get_recall()
-                        self.nr_prob_target_path.append(self.min_prob_target)
+                        if len(self.ratings)>0:
+                            self.n_remaining=self.N - self.seen_docs
+                            Xs = np.cumsum(np.array(self.ratings[::-1]))
+                            ns = np.arange(len(self.ratings))
+                            hypotheticals = vgen_hypo(Xs, self.r_seen, self.recall_target)
+                            prob_target = hypergeom.cdf(Xs, self.n_remaining+ns, hypotheticals, ns+1)
+                            self.min_prob_target = prob_target.min()
+
+                            if self.min_prob_target < 1-self.alpha and self.wss_nrs is None:
+                                self.wss_nrs = 1 - self.seen_docs / self.N
+                                self.recall_nrs = self.get_recall()
+                            self.nr_prob_target_path.append(self.min_prob_target)
                 time_checking=time.time()-t0
             if self.wss_nrs is None:
[…]

### 2019-10-22 analysis/run_scenarios.py

@@ -1,15 +1,20 @@
 import pandas as pd
 import numpy as np
 import rapid_review as rr
-import os
+import os, sys
 import matplotlib.pyplot as plt
 import importlib
 from sklearn.svm import SVC, OneClassSVM
 import argparse
+from functools import partial
+
 
 parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
 parser.add_argument('iterations', type=int)
+parser.add_argument('-p', type=int, default=0)
 args = parser.parse_args()
+
+print(args.p)
 
 ##########################################
 ## Pull the document metadata from the xml files from the Pubmed API
@@ -40,7 +45,6 @@ cohen_db = pd.read_csv(
 cohen_db['relevant'] = np.where(cohen_db['ab_relevant']=="I",1,0)
 cohen_db = cohen_db[["review","PMID","relevant"]]
 
-cohen_db.head()
 
 models = [
     SVC(kernel='linear',class_weight='balanced',probability=True)
@@ -50,6 +54,7 @@ iterations = args.iterations
 results = []
 rs_results = []
 paths = []
+
 for name, group in cohen_db.groupby('review'):
     df = pd.merge(
         group,
@@ -64,11 +69,21 @@ for name, group in cohen_db.groupby('review'):
         ss = rr.ScreenScenario(
             df, models, s, [50, 100, 200], name
         )
-        for i in range(iterations):
-            print(i)
-            r = ss.screen(i, True)
-            if r is not None:
-                results.append(r)
+
+        if args.p:
+            from multiprocessing import Pool
+            def simulate_screening_parallel(i,ss):
+                return ss.screen(i, True)
+            with Pool(args.p) as pool:
+                results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
+            print(results)
+            break
+        else:
+            for i in range(iterations):
+                print(i)
+                r = ss.screen(i, True)
+                if r is not None:
+                    results.append(r)
 
 results_df = pd.DataFrame.from_dict(results)
 results_df.to_csv('../results/results.csv', index=False)

### 2019-10-22 analysis/run_scenarios_copd_pb.py

@@ -6,9 +6,11 @@ import matplotlib.pyplot as plt
 import importlib
 from sklearn.svm import SVC, OneClassSVM
 import argparse
+from functools import partial
 
 parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
 parser.add_argument('iterations', type=int)
+parser.add_argument('-p', type=int, default=0)
 args = parser.parse_args()
 
 importlib.reload(rr)
@@ -54,11 +56,20 @@ for name, group in df.groupby('review'):
         ss = rr.ScreenScenario(
             group, models, s, [50, 100, 200], name
         )
-        for i in range(iterations):
-            print(i)
-            r = ss.screen(i, True)
-            if r is not None:
-                results.append(r)
-
+        if args.p:
+            from multiprocessing import Pool
+            def simulate_screening_parallel(i,ss):
+                return ss.screen(i, True)
+            with Pool(args.p) as pool:
+                results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
+            print(results)
+            break
+        else:
+            for i in range(iterations):
+                print(i)
+                r = ss.screen(i, True)
+                if r is not None:
+                    results.append(r)
+                    
 results_df = pd.DataFrame.from_dict(results)
 results_df.to_csv('../results/results_pb_copd.csv', index=False)

### 2019-10-22 analysis/run_scenarios_fr.py

@@ -4,11 +4,13 @@ import rapid_review as rr
 import os
 import matplotlib.pyplot as plt
 import importlib
+from functools import partial
 from sklearn.svm import SVC, OneClassSVM
 import argparse
 
 parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
 parser.add_argument('iterations', type=int)
+parser.add_argument('-p', type=int, default=0)
 args = parser.parse_args()
 
 datasets = ["Hall","Kitchenham","Radjenovic","Wahono"]
@@ -41,11 +43,20 @@ for name, group in frdf.groupby('review'):
         ss = rr.ScreenScenario(
             df, models, s, [50, 100, 200], name
         )
-        for i in range(iterations):
-            print(i)
-            r = ss.screen(i, True)
-            if r is not None:
-                results.append(r)
+        if args.p:
+            from multiprocessing import Pool
+            def simulate_screening_parallel(i,ss):
+                return ss.screen(i, True)
+            with Pool(args.p) as pool:
+                results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
+            print(results)
+            break
+        else:
+            for i in range(iterations):
+                print(i)
+                r = ss.screen(i, True)
+                if r is not None:
+                    results.append(r)
 
 results_df = pd.DataFrame.from_dict(results)
 results_df.to_csv('../results/results_fr.csv', index=False)

### 2019-10-22 analysis/run_scenarios.py
mpi

@@ -3,7 +3,6 @@ import numpy as np
 import rapid_review as rr
 import os, sys
 import matplotlib.pyplot as plt
-import importlib
 from sklearn.svm import SVC, OneClassSVM
 import argparse
 from functools import partial
@@ -12,6 +11,7 @@ from functools import partial
 parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
 parser.add_argument('iterations', type=int)
 parser.add_argument('-p', type=int, default=0)
+parser.add_argument('-mpi', type=int, default=0)
 args = parser.parse_args()
 
 print(args.p)
@@ -78,6 +78,16 @@ for name, group in cohen_db.groupby('review'):
                 results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
             print(results)
             break
+        elif args.mpi:
+            from mpi4py import MPI
+            comm = MPI.COMM_WORLD
+            num_procs = comm.Get_size()
+            rank = comm.Get_rank()
+            stat = MPI.Status()
+            r = ss.screen(rank, True)
+            results.append(r)
+            results_df = pd.DataFrame.from_dict(results)
+            results_df.to_csv(f'../results/results_{rank}.csv', index=False)
         else:
             for i in range(iterations):
                 print(i)

### 2019-10-22 analysis/run_scenarios.py
mpl

@@ -2,7 +2,6 @@ import pandas as pd
 import numpy as np
 import rapid_review as rr
 import os, sys
-import matplotlib.pyplot as plt
 from sklearn.svm import SVC, OneClassSVM
 import argparse
 from functools import partial

### 2019-10-22 analysis/run_scenarios.py
mpi all dbs

@@ -75,8 +75,6 @@ for name, group in cohen_db.groupby('review'):
                 return ss.screen(i, True)
             with Pool(args.p) as pool:
                 results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
-            print(results)
-            break
         elif args.mpi:
             from mpi4py import MPI
             comm = MPI.COMM_WORLD
@@ -85,8 +83,6 @@ for name, group in cohen_db.groupby('review'):
             stat = MPI.Status()
             r = ss.screen(rank, True)
             results.append(r)
-            results_df = pd.DataFrame.from_dict(results)
-            results_df.to_csv(f'../results/results_{rank}.csv', index=False)
         else:
             for i in range(iterations):
                 print(i)
@@ -94,5 +90,11 @@ for name, group in cohen_db.groupby('review'):
                 if r is not None:
                     results.append(r)
 
-results_df = pd.DataFrame.from_dict(results)
-results_df.to_csv('../results/results.csv', index=False)
+    break
+
+if args.mpi:
+    results_df = pd.DataFrame.from_dict(results)
+    results_df.to_csv(f'../results/results_{rank}.csv', index=False)
+else:                    
+    results_df = pd.DataFrame.from_dict(results)
+    results_df.to_csv('../results/results.csv', index=False)

### 2019-10-22 analysis/run_scenarios_copd_pb.py

@@ -11,6 +11,7 @@ from functools import partial
 parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
 parser.add_argument('iterations', type=int)
 parser.add_argument('-p', type=int, default=0)
+parser.add_argument('-mpi', type=int, default=0)
 args = parser.parse_args()
 
 importlib.reload(rr)
@@ -62,14 +63,24 @@ for name, group in df.groupby('review'):
                 return ss.screen(i, True)
             with Pool(args.p) as pool:
                 results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
-            print(results)
-            break
+        elif args.mpi:
+            from mpi4py import MPI
+            comm = MPI.COMM_WORLD
+            num_procs = comm.Get_size()
+            rank = comm.Get_rank()
+            stat = MPI.Status()
+            r = ss.screen(rank, True)
+            results.append(r)
         else:
             for i in range(iterations):
                 print(i)
                 r = ss.screen(i, True)
                 if r is not None:
                     results.append(r)
-                    
-results_df = pd.DataFrame.from_dict(results)
-results_df.to_csv('../results/results_pb_copd.csv', index=False)
+
+if args.mpi:
+    results_df = pd.DataFrame.from_dict(results)
+    results_df.to_csv(f'../results/results_{rank}.csv', index=False)
+else:                    
+    results_df = pd.DataFrame.from_dict(results)
+    results_df.to_csv('../results/results.csv', index=False)

### 2019-10-22 analysis/run_scenarios_fr.py

@@ -11,6 +11,7 @@ import argparse
 parser = argparse.ArgumentParser(description="run ML systematic review scenarios")
 parser.add_argument('iterations', type=int)
 parser.add_argument('-p', type=int, default=0)
+parser.add_argument('-mpi', type=int, default=0)
 args = parser.parse_args()
 
 datasets = ["Hall","Kitchenham","Radjenovic","Wahono"]
@@ -49,8 +50,14 @@ for name, group in frdf.groupby('review'):
                 return ss.screen(i, True)
             with Pool(args.p) as pool:
                 results.append(pool.map(partial(simulate_screening_parallel, ss=ss), list(range(iterations))))
-            print(results)
-            break
+        elif args.mpi:
+            from mpi4py import MPI
+            comm = MPI.COMM_WORLD
+            num_procs = comm.Get_size()
+            rank = comm.Get_rank()
+            stat = MPI.Status()
+            r = ss.screen(rank, True)
+            results.append(r)
         else:
             for i in range(iterations):
                 print(i)
@@ -58,5 +65,9 @@ for name, group in frdf.groupby('review'):
                 if r is not None:
                     results.append(r)
 
-results_df = pd.DataFrame.from_dict(results)
-results_df.to_csv('../results/results_fr.csv', index=False)
+if args.mpi:
+    results_df = pd.DataFrame.from_dict(results)
+    results_df.to_csv(f'../results/results_fr_{rank}.csv', index=False)
+else:                    
+    results_df = pd.DataFrame.from_dict(results)
+    results_df.to_csv('../results/results_fr.csv', index=False)

### 2019-10-23 analysis/run_scenarios.py
remove mpl

@@ -90,8 +90,6 @@ for name, group in cohen_db.groupby('review'):
                 if r is not None:
                     results.append(r)
 
-    break
-
 if args.mpi:
     results_df = pd.DataFrame.from_dict(results)
     results_df.to_csv(f'../results/results_{rank}.csv', index=False)

### 2019-10-23 analysis/run_scenarios_copd_pb.py

@@ -2,7 +2,6 @@ import pandas as pd
 import numpy as np
 import rapid_review as rr
 import os
-import matplotlib.pyplot as plt
 import importlib
 from sklearn.svm import SVC, OneClassSVM
 import argparse

### 2019-10-23 analysis/run_scenarios_fr.py

@@ -2,7 +2,6 @@ import pandas as pd
 import numpy as np
 import rapid_review as rr
 import os
-import matplotlib.pyplot as plt
 import importlib
 from functools import partial
 from sklearn.svm import SVC, OneClassSVM

### 2019-10-23 analysis/rapid_review.py
small sample

@@ -109,7 +109,7 @@ class ScreenScenario:
             #stop_words="english",tokenizer=tokenize
         ).fit_transform(df['x'])
 
-        clf = IsolationForest()
+        clf = IsolationForest(behaviour="new")
         clf.fit(self.X)
         self.df['outlying'] = clf.predict(self.X)
 
@@ -155,8 +155,12 @@ class ScreenScenario:
         self.reset()
         ## Do the random sample
         if s > self.df.shape[0]*0.5:
-            print(f"skipping sample {s}, as it is more than 50% of the data")
-            return
+            s = df.shape[0]*0.5
+            #print(f"skipping sample {s}, as it is more than 50% of the data")
+            #return {}
+            #ignore_fields = ["df", "X", "unseen_p"]
+            #result = {k: v for k, v in self.__dict__.items() if k not in ignore_fields}
+
         sids = random.sample(list(self.df.index), s)
         self.df.loc[sids,'seen'] = 1
         self.seen_docs = s

### 2019-10-23 analysis/run_scenarios.py

@@ -82,17 +82,19 @@ for name, group in cohen_db.groupby('review'):
             rank = comm.Get_rank()
             stat = MPI.Status()
             r = ss.screen(rank, True)
-            results.append(r)
+            if r is not None:
+                results.append(r)
         else:
             for i in range(iterations):
                 print(i)
                 r = ss.screen(i, True)
-                if r is not None:
-                    results.append(r)
+                if r is not None: 
+                   results.append(r)
 
 if args.mpi:
     results_df = pd.DataFrame.from_dict(results)
     results_df.to_csv(f'../results/results_{rank}.csv', index=False)
-else:                    
+else:
+    results = [r for r in results if r is not None]
     results_df = pd.DataFrame.from_dict(results)
     results_df.to_csv('../results/results.csv', index=False)

### 2019-10-23 analysis/run_scenarios_copd_pb.py

@@ -69,7 +69,8 @@ for name, group in df.groupby('review'):
             rank = comm.Get_rank()
             stat = MPI.Status()
             r = ss.screen(rank, True)
-            results.append(r)
+            if r is not None:
+                results.append(r)
         else:
             for i in range(iterations):
                 print(i)

### 2019-10-23 analysis/run_scenarios_fr.py

@@ -56,7 +56,8 @@ for name, group in frdf.groupby('review'):
             rank = comm.Get_rank()
             stat = MPI.Status()
             r = ss.screen(rank, True)
-            results.append(r)
+            if r is not None:
+                results.append(r)
         else:
             for i in range(iterations):
                 print(i)

### 2019-10-23 analysis/rapid_review.py
typos

@@ -109,7 +109,7 @@ class ScreenScenario:
             #stop_words="english",tokenizer=tokenize
         ).fit_transform(df['x'])
 
-        clf = IsolationForest(behaviour="new")
+        clf = IsolationForest(behaviour="new", contamination="auto")
         clf.fit(self.X)
         self.df['outlying'] = clf.predict(self.X)
 
@@ -155,7 +155,7 @@ class ScreenScenario:
         self.reset()
         ## Do the random sample
         if s > self.df.shape[0]*0.5:
-            s = df.shape[0]*0.5
+            s = self.df.shape[0]*0.5
             #print(f"skipping sample {s}, as it is more than 50% of the data")
             #return {}
             #ignore_fields = ["df", "X", "unseen_p"]
@@ -283,6 +283,9 @@ class ScreenScenario:
             if self.wss_bir_ci is None:
                 self.wss_bir_ci = 0
                 self.recall_bir_ci = 1
+            if self.wss_hyper is None:
+                self.wss_hyper = 0
+                self.recall_hyper = 1
             
             for ih in self.irrelevant_heuristic:
                 if getattr(self, f'wss_ih_{ih}') is None:

### 2019-10-23 analysis/rapid_review.py
rounding

@@ -155,7 +155,7 @@ class ScreenScenario:
         self.reset()
         ## Do the random sample
         if s > self.df.shape[0]*0.5:
-            s = self.df.shape[0]*0.5
+            s = int(round(self.df.shape[0]*0.5))
             #print(f"skipping sample {s}, as it is more than 50% of the data")
             #return {}
             #ignore_fields = ["df", "X", "unseen_p"]

### 2019-11-19 manuscript/main.tex
0 draft

@@
 \usepackage{subcaption}
+\usepackage{booktabs}
 %\def\includegraphic{}
@@
 				Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. 
-				Machines and humans  work together, with the human providing training data, and the machine optimising the documents a human screens, so that they can identify all relevant documents after viewing only a fraction of the total documents.
+				Machines and humans  work together, with humans providing training data, and the machine optimising the documents that the humans screen.
+This enables the identification of all relevant documents after viewing only a fraction of the total documents.
 				However, current approaches lack robust stopping criteria, so that reviewers do not know when they have seen all or a certain proportion of relevant documents.
@@
 
-	Machine learning for evidence synthesis is an rapidly emerging field [cite].
-	Machine learning enabled interventions have been proposed at various points in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
-	A major strand of the literature works on screening - the identification of relevant documents in a set of documents whose relevance is uncertain \cite{OMara-Eves2015}. 
-	For this task, several papers have developed or evaluated active learning.
+	Evidence synthesis technology is a rapidly emerging field that promises to change the practice of evidence synthesis work \cite{Westgate2018}.
+	Interventions have been proposed at various points in order to reduce the human effort required to produce systematic reviews and other forms of evidence synthesis.
+	A major strand of the literature works on screening: the identification of relevant documents in a set of documents whose relevance is uncertain \cite{OMara-Eves2015}. 
+	This is a time consuming and repetitive task, and in a research environment with constrained resources and increasing amounts of literature, this may limit the scope of the evidence synthesis projects undertaken.
+	Several papers have developed Active Learning (AL) approaches \cite{miwa2014, Wallace2010a, Wallace2010, Jonnalagadda2013, Przybya2018} to reduce the time required to screen documents.
+This paper sets out how current approaches are  unsuitable in practice, and outlines and evaluates a small modification that would make AL systems ready for live reviews.
 
 	Active learning is an iterative process where documents screened by humans are used to train a machine learning model to predict the relevance of unseen papers \cite{Settles2009}.
-	The most relevant studies are passed back to the human and rated, generating more labels to feed back to the machine. 
+	The algorithm chooses which studies will next be screened by humans, often those which are likely to be relevant or about which the model is uncertain, in order to generate more labels to feed back to the machine.
 	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (recall) - before having seen all the documents in the corpus.
-	The proportion of documents not yet seen by the human when they reach the given recall threshold are referred to as the work saved, as this represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.
+	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved, as this represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.
 
-	In live review settings, however, the recall remains unknown until all documents have been screened. 
-	In order for work to really be saved, reviewers have to stop screening while uncertain about the recall.
-This is particularly problematic in systematic reviews because low recall increases the risk of publication bias %\cite{Lefebrve in Cochrane}
+	Machine learning applications are often evaluated using sets of documents from already completed systematic reviews for which inclusion or exclusion labels already exist. 
+	As all human labels are known \textit{a priori}, it is possible to simulate the screening process, recording when a given recall target has been achieved.
+	In live review settings, however, recall remains unknown until all documents have been screened. 
+	In order for work to really be saved, reviewers have to stop screening while uncertain about recall. 
+	This is particularly problematic in systematic reviews because low recall increases the risk of bias \cite{Lefebvre2011}.
 	The lack of appropriate stopping criteria has therefore been identified as a research gap \cite{bannach-brown2019}, although some approaches have been suggested.
-These fall into the following categories
+These fall into the following categories:
 	\begin{itemize}
@@
 They stop when this number, or a given proportion of it, has been reached \cite{Shemilt2014}
-		\item \textbf{Heuristics:} Reviewers stop when a given number of irrelevant articles are seen in a row \cite{Przybya2018}. 
+		\item \textbf{Heuristics:} Reviewers stop when a given number of irrelevant articles are seen in a row \cite{Jonnalagadda2013, Przybya2018}. 
 		\item \textbf{Pragmatic criteria:} Reviewers stop when they run out of time \cite{miwa2014}. 
@@
 
-	We show in this paper how the first two criteria are inadequate.
-	We argue that the inadequacy lies in the unreliability - particularly across different domains, or datasets with different properties \todo{ref on cross-domain reliability} - both of the work saved and the recall achieved.
-	Without the reliable or reportable achievement of a desired level of recall, AL systems in live reviews remain challenging.
+	We show in this paper how existing criteria are inadequate.
+	We demonstrate with several previously used datasets that their inadequacy lies in the unreliability - particularly across different domains, or datasets with different properties \cite{OMara-Eves2015} - both of the work saved and the recall achieved.
+	Without the reliable or reportable achievement of a desired level of recall, deployment of AL systems in live reviews remains challenging.
 
 	This study proposes a system for estimating the recall based on random sampling of remaining documents. 
-	We use basic statistics to test the hypothesis that less than a given recall threshold has been achieved.
-	This allows AL users to predefine a threshold in terms of uncertainty and recall, so that they can make transparent, understandable statements like ``There is a <5\% chance that we achieve a recall under 95\%''.
+	We use a simple statistical method to iteratively test a null hypothesis that the recall achieved is less than a given target recall.
+	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``There is a <5\% chance that we achieve a recall under 95\%''.
 
@@
 
-	We evaluate this stopping criteria on real-world systematic review datasets on which active learning systems have previously been tested.
+	We evaluate our stopping criteria, along with heuristic and sampling based criteria on real-world systematic review datasets on which AL systems have previously been tested \cite{Cohen2006, Yu2019, Terasawa2009, Castaldi2009}.
 
@@
 
+	We start by explaining the sampling and heuristic based stopping criteria and showing their methodological weaknesses. 
+Then we introduce our own suggested stopping criteria before testing all criteria on real world datasets.
+
 	\subsection*{Existing Stopping Criteria for Active Learning}
-
-	We start by explaining the sampling and heuristic based stopping criteria and showing how their methodological weaknesses. 
-	Then we introduce our own suggested stopping criteria before testing all criteria on real world datasets.
 
@@
 	However, the estimation of the BIR fails to take into account sampling uncertainty. 
-	Figure \ref{bir-sampling}.a shows for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant), the predicted number of documents after each document drawn. 
+	Figure \ref{bir-sampling} shows the predicted number of documents after each document drawn for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant). 
 	Depending on the luck of the draw, one might estimate the true number of documents as being
-	\begin{itemize}
+	\begin{enumerate}
 		\item much higher than the true value, meaning that the stopping criterion would never be reached - and no work could be saved, or;
 		\item much lower than the true value, meaning that the stopping criterion would be reached before the desired level of recall was actually achieved
-	\end{itemize}
+	\end{enumerate}
 
@@
 For 50 random samples of 2,000 documents from 20,000 documents, we show the estimated number of relevant documents, after each document is drawn.
-The two lines correspond to 100\% of relevant documents, and $\frac{100}{0.95}$\% of relevant documents.}
-	\label{bir-sampling}.
-
+The two horizontal grey  lines correspond to the actual number of relevant documents, and to $\frac{100}{0.95}$\% of relevant documents.
 	This shows how even after taking a large sample, baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
+		\label{bir-sampling}
+	}
 \end{figure}
-
-\subsubsection*{Heuristic Based Stopping Criteria}
-
-Some studies give the example of heuristic based stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Przybya2018} [+REFS]. 
-We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low. 
-We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; and 2) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.
-
-Figure \ref{unseen-proportions} illustrates this second point. 
-We show two scenarios with identical low proportions of relevant documents observed in the unseen documents. 
-Recall is not simply a function of the relevance of unseen documents, but also of the number of unseen documents. 
-This also means that where machine learning has performed well (as in the top figure), low proportions of irrelevant documents in those that remain are indicative of lower recall than where ML has performed less well.
-Likewise, where the proportion of relevant documents is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low.
 
@@
 The top bar shows a random distribution of relevant documents (green) and irrelevant documents (red) at a given proportion of relevance.
-The bottom bar  shows distributions of relevant and irrelevant documents in hypothetical sets of seen (right) and unseen (left - transparent) documents.}
+The bottom bar  shows distributions of relevant and irrelevant documents in hypothetical sets of seen (right) and unseen (left - transparent) documents.}Jonnalagadda2013
 	\label{unseen-proportions}
 \end{figure}
+
+\subsubsection*{Heuristic Based Stopping Criteria}
+
+Some studies give the example of heuristic based stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
+We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low. 
+We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; and 2) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.
+
+Figure \ref{unseen-proportions} illustrates this second point. 
+We show two scenarios with identical low proportions of relevant documents observed in the unseen documents.
+In the top figure, machine learning (ML) has performed well, and 74\% of documents seen were relevant. 
+In the bottom figure,  ML has performed less well, and only 26\% of documents seen were relevant.
+In both cases, only 2\% of unseen documents are relevant, but 2\% of a larger number means more relevant documents are missed.
+Recall is not simply a function of the relevance of unseen documents, but also of the number of unseen documents. 
+This also means that where ML has performed well (as in the top figure), a low proportion of relevant documents in those not yet checked is indicative of lower recall than where ML has performed less well.
+Likewise, where the proportion of relevant documents in the whole corpus is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low.
 
@@
 However, reviewers are not given the opportunity to specify a confidence level, and for two of the four datasets in which they tested their criteria, the median achieved recall at a stopping criteria targeting 95\% recall was below 95\%.
-\cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold.
+\cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold, rather than above it in a reliable proportion of cases.
 These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount.
@@
 
-	Our aim is to formulate a stopping criterion such that a recall of $\tau$ has been reached with confidence level $\alpha$.
-	To achieve this, we formulate a hypothesis test:
-	We first calculate the number $\hat{n}$ of relevant documents that can maximally remain in the unseen documents to achieve a given recall target.
+	After using machine learning to select which documents are screened by humans as described above, we begin drawing a random sample from the remaining documents (when this happens is described below).
+
+	We use random sampling to estimate the probability that a target recall $\tau$ has been achieved.
+Because we are sampling a binary outcome without replacement, we can use the hypergeometric distribution to formulate a statistical test.
+The hypergeometric distribution tells us the statistical significance of observing $k$ successes (relevant documents) in $n$ draws from a finite population of $N$ documents with $K$ successes.
+
+	\begin{equation}
+		k \sim Hypergeometric(N, K, n)
+	\end{equation}
+
+	In our case, we know $k$, $n$ and $N$ after each draw, but $K$ is unknown.
+We therefore substitute a hypothetical value $\hat{K}$: the minimum number of relevant documents in the sample had the recall target been missed.
+
+	Recall $R$ is given by the number of relevant documents that have been seen $\dot{\rho}$ over the number of relevant documents in the whole dataset $\rho$
+
+	\begin{equation}
+		R = \frac{\dot{\rho}}{\rho}
+	\end{equation}
+
+	The number of relevant documents in the whole dataset is composed of $\hat{\rho}$ relevant documents seen before random sampling began and $\tilde{K}$ relevant documents unseen at the start of random sampling.
+We can therefore express $R$ as
+
+	\begin{equation}
+		R = \frac{\dot{\rho}}{\hat{\rho} + \tilde{K}},
+	\end{equation}
+
+	Substituting the target recall $\tau$ for $R$, reorganising and rounding up to the next integer, we can, after each draw, calculate $\hat{K}$, which is the minimum number of relevant documents that could have been remaining when random sampling started, if recall were lower than the target.
+
+	\begin{equation}
+		\hat{K} = \lceil \frac{\rho}{\tau} - \hat{\rho} \rceil
+	\end{equation}
+
+	We use the cumulative distribution function to estimate the probability $p$ of having observed $k$ or fewer relevant documents in the sample given $\hat{K}$, or, in other words, the probability of observing as many relevant documents in our random sample as we did, if our recall target had not been achieved.
+If this is below our confidence level $1 - \alpha$, we can reject the null hypothesis that the recall target was not achieved.
+
+	%%%%%%%%%%%
+%	
+%	We calculate a hypothetical value $\hat{K}$
+%
+%	Our aim is to formulate a stopping criterion such that a recall of $\tau$ has been reached with confidence level $\alpha$.
+%	To achieve this, we formulate a hypothesis test:
+%	We first calculate the number $\hat{n}$ of relevant documents that can maximally remain in the unseen documents to achieve a given recall target.
 The recall is defined as the ratio of seen relevant documents over all relevant documents
-	\begin{equation}
-		\tau = \frac{\rho}{\hat{\rho} + \tilde{n}},
-	\end{equation}
-	with the total number of seen relevant documents $\rho$, the number of seen relevant documents at the time we start random sampling $\hat{\rho}$, and the remaining number of relevant documents in the unseen sample at the time random sampling starts $\tilde{n}$.
-	Reorganization of this equation gives
-	\begin{equation}
-		\tilde{n} = \frac{\rho}{\tau} - \hat{\rho}.
-	\end{equation}
-	Because we need to use integers, we get the next largest integer given by the ceiling function $\lceil \cdot \rceil$:
-	\begin{equation}
-		\hat{n} = \lceil \frac{\rho}{\tau} - \hat{\rho} \rceil
-	\end{equation}
-	The null-hypothesis to reject can then be formulated as:
-	\begin{quote}
-	$H_0$: There are $\hat{n}$ or more relevant documents remaining in the set of unseen documents.
-	\end{quote}
-
-	% notation here follows notation for scipy description of hypergeometric distribution
-	% better use another notation more in line with the literature? -> literature seems to use very different notations
-
-	In theory, we draw a random sample of $N$ documents, evaluate how many of them are relevant, and denote this number with $k$.
+%
+%	with the total number of seen relevant documents $\rho$, the number of seen relevant documents at the time we start random sampling $\hat{\rho}$, and the remaining number of relevant documents in the unseen sample at the time random sampling starts $\tilde{n}$.
+%	Reorganization of this equation gives
+%	\begin{equation}
+%		\tilde{n} = \frac{\rho}{\tau} - \hat{\rho}.
+%	\end{equation}
+%	Because we need to use integers, we get the next largest integer given by the ceiling function $\lceil \cdot \rceil$:
+%
+%	The null-hypothesis to reject can then be formulated as:
+%	\begin{quote}
+%	$H_0$: There are $\hat{n}$ or more relevant documents remaining in the set of unseen documents.
+%	\end{quote}
+%	
+%	% notation here follows notation for scipy description of hypergeometric distribution
+%	% better use another notation more in line with the literature? -> literature seems to use very different notations
+%	
+%	In theory, we draw a random sample of $N$ documents, evaluate how many of them are relevant, and denote this number with $k$.
 We can estimate the probability that we observe $k$ or fewer relevant documents in a random draw of $N$ documents without replacement from a pool of $M$ documents, of which $n$ are relevant using the cumulative distribution function of the hypergeometric distribution.
-This function is defined as the sum over the probability mass function of the hypergeometric distribution $p(x, M, n, N)$:
-	\begin{equation}	
-		 cdf(k, M, n, N) = \sum_{x \leq k} p(x, M, n, N) = \sum_{x \leq k} \frac{\binom{n}{x} \binom{M - n}{N - x}} {\binom{M}{N}},
-		 \label{pfunc}
-	\end{equation}
-	where $\binom{a}{b}$ denote binomial coefficients. In our case, $M$ is the number of unseen documents before we start random sampling.
-
-	We can reject the null hypothesis if the p-value $p = cdf(k,M,\hat{n},N)$ is smaller than $1 - \alpha$ [references to hypergeometric testing?, needs further argumentation?].
+%This function is defined as the sum over the probability mass function of the hypergeometric distribution $p(x, M, n, N)$:
+%	\begin{equation}	
+%		 cdf(k, M, n, N) = \sum_{x \leq k} p(x, M, n, N) = \sum_{x \leq k} \frac{\binom{n}{x} \binom{M - n}{N - x}} {\binom{M}{N}},
+%		 \label{pfunc}
+%	\end{equation}
+%	where $\binom{a}{b}$ denote binomial coefficients. In our case, $M$ is the number of unseen documents before we start random sampling.
+%	
+%	We can reject the null hypothesis if the p-value $p = cdf(k,M,\hat{n},N)$ is smaller than $1 - \alpha$ [references to hypergeometric testing?, needs further argumentation?].
 Because we are testing for the lower tail, this implies that we can be confident with a level $\alpha$ that the number of unseen relevant documents is lower than $\hat{n}$.
-
-	In practice, we can draw random unseen documents repeatedly. After each draw, we calculate the p-value using the number of relevant documents $k$ and the number of randomly sampled documents $N$.
-	We can stop the screening once a p-value $p < 1 - \alpha$ is reached.
+%	
+%	In practice, we can draw random unseen documents repeatedly. After each draw, we calculate the p-value using the number of relevant documents $k$ and the number of randomly sampled documents $N$.
+%	We can stop the screening once a p-value $p < 1 - \alpha$ is reached.
 
@@
 
-	After reviewing each document, $S$ documents are screened, and $U$ documents are yet to be seen.
-We treat $i = 1 \dots S$ of the previously screened documents as a random sample, and calculate $p$ as before for each sample, taking the minimum across all samples $p_{min}$.
+	After reviewing each document, $S$ documents have been screened, and $U$ documents are yet to be seen.
+We treat $i = 1 \dots S$ of the previously screened documents as a random sample, and calculate $p$, using the method above, for each sample, taking the minimum across all samples $p_{min}$.
 If $p_{min}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling.
-We also calculate $p_{min}$ for the remaining values as if we had not switched to random sampling and present these in the results below.
+We also calculate $p_{min}$ for the remaining documents as if we had not switched to random sampling and record the recall and work saved when $p_{min} < 1 - \alpha$.
+We present these in the results below as the psuedo-random sampling criterion.
 
@@
 
-	We evaluate each of the three criteria discussed on real world test data. With each method, we run 100 runs and record 
+	We evaluate each of the criteria discussed, operationalising the heuristic stopping criteria with 50, 100, and 200 consecutive irrelevant records, on real world test data.
+We run 100 iterations on each dataset and record
 	\begin{itemize}
-		\item \textbf{Actual Recall}: The recall when the stopping criteria is met
-		\item \textbf{WS-SC}: Work saved when the stopping criteria is met \footnote{we do not use work saved over sampling, is it is not axiomatic that seeing 95\% of a random sample would achieve 95\% recall}
-		\item \textbf{Additional Burden}: The proportion of the literature that need to be screened in order to \textit{be confident} that 95\% recall has been achieved, after 95\% of documents have actually been identified.
-		\item \textbf{Missed Recall}: The amount by which, if any, the recall target was missed - in percentage points
+		\item \textbf{Actual Recall}: The recall when the stopping criteria was met
+		\item \textbf{WS-SC}: Work saved when the stopping criteria was met \footnote{We do not use work saved over sampling, is it is not axiomatic that seeing 95\% of a random sample would achieve 95\% recall}
+		\item \textbf{Additional Burden}: the work saved when the criterion was triggered subtracted from the work saved when the recall target was actually achieved.
+
 	\end{itemize}
-	We operationalise the heuristic stopping criteria with 50, 100, and 200 irrelevant records in a row.
-For simplicity, we use a basic SVM model [REFS], with 1-2 word n-grams in the document abstracts used as input data.
-We start with random samples of [200 and 500] documents.
-
-	As the intention is to evaluate stopping criteria, rather than machine learning approaches, we focus on additional burden and missed recall.
-These parameters correspond to the tradeoff inherent in stopping criteria: stopping too early means missing the recall target, stopping too late means screening more documents than was actually required to meet the target.
+	For simplicity, we use a basic SVM model \cite{Cortes95, Pedregosa2011}, with 1-2 word n-grams taken from the document abstracts used as input data.
+We start with random samples of 200 documents.
+Subsequently, we ``screen'', that is, we reveal the labels of, batches of the 20 documents with the highest predicted relevance scores, retraining the model after each batch.
+Each criterion is evaluated after each document is ``screened''.
+	 For our criteria, we set the target recall value to 95\% and the confidence level to 95\%.
 
 	%\subsubsection*{Evaluation Data}
+	\begin{table}
+		\input{../tables/datasets.tex}
+		\caption{Dataset properties}
+		\label{tab:data}
+	\end{table}
 
-	We use the collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}, along with PB COPD, and FASTREAD [CITE].
-We feel that testing on datasets with different properties and from different domains is key to establishing criteria appropriate for general use.
+	The systematic review datasets used for testing are described in table \ref{tab:data}.
+We use the seminal collection of systematic reviews used to develop machine learning applications for document screening by Aaron Cohen and co-authors in 2006 \cite{Cohen2006}, along with the widely used Proton Beam \cite{Terasawa2009} and COPD \cite{Castaldi2009} datasets, and computer science datasets used to test FASTREAD \cite{Yu2019}.
+Testing on datasets with different properties and from different domains is key to establishing criteria appropriate for general use.
+Choosing as broad as possible data also prevents us from being able to ``tune'' our machine learning approach in ways that may work well for specific datasets but not generalise well.
+Work savings, even maximum work savings are therefore below the state of the art recorded for each of these datasets.
 In this way we can show how well the criteria perform even when the model performs badly.
+
+	%As a last step, we explain the factors that increase or decrease the performance of our criteria, using a simple regression model.
 
@@
 	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criteria has been satisfied. 
-	For comparison, we also include the results that would have been achieved with perfect foresight, that is, the work saved when the 95\% recall target has been received.
+	For comparison, we also include the results that would have been achieved with \textit{a priori} knowledge of the data, that is, the work saved when the 95\% recall target was actually reached.
+In a live systematic review, reviewers would never know when this had been reached, but these are the work savings most often reported in machine learning for systematic review screening studies.
 
-	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in [33]\% of cases, while the most common work saving is 0\%.
-This is in line with out expectations, that due to random sampling error, the expected number of document will often be over-estimated or under-estimated, resulting in poor work savings or poor recall.
-
-	The Heuristic based stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
-Although the
+	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
+This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.
 
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_pf}
+		\includegraphics[width=\textwidth]{../images/jointplot_pf.pdf}
 		\caption[]%
-		{{\footnotesize Perfect Foresight}}    
+		{{\footnotesize \textit{a priori} knowledge}}   
 		\label{fig:pf}
@@
 
-	\caption{\small Distribution of recall and work saved after each stopping criteria.} 
+	\caption{\small Distribution of recall and work saved after each stopping criteria.
+Green dots show results for datasets with less than 1,000 documents, orange dots show datasets with 1,000 - 2,000 documents, and blue dots show datasets with more than 2,000 documents}
 	\label{recall-wss}
@@
 
+\begin{figure*}
+	\centering
+	\begin{subfigure}[b]{0.475\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_bir}
+		\caption[Network2]%Example-Image
+		{{\small Baseline Sampling}}    
+		\label{fig:bir_ab}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}  
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_ih_50}
+		\caption[]%
+		{{\small 50 consecutive irrelevant results}}    
+		\label{fig:ih_50_ab}
+	\end{subfigure}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_hyper}
+		\caption[]%
+		{{\small Hypergeometric sampling \\}}    
+		\label{fig:hyper_ab}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_ih_100}
+		\caption[]%
+		{{\small 100 consecutive irrelevant results }}    
+		\label{fig:ih_200_ab}
+
+	\end{subfigure}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_nrs.pdf}
+		\caption[]%
+		{{\footnotesize Pseudorandom hypergeometric sampling}}    
+		\label{fig:nrs_ab}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_ih_200.pdf}
+		\caption[]%
+		{{\footnotesize 200 consecutive irrelevant results }}    
+		\label{fig:ih_100_ab}
+	\end{subfigure}
+
+	\caption{\small Distribution of recall and additional burden after each stopping criteria.
+Additional burden is the work saved when the criteria was trigerred subtracted from the work saved when the target was reached.}
+	\label{recall-burden}
+\end{figure*}
+
+	The Heuristic based stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
+Although the mean work saved for IH50 is 41\%, the target is missed in 39\% of cases.
+The cases below the horizontal grey line indicate instances where work has been saved at the expense of achieving the recall target.
+
+	Both the random sampling and the pseudorandom sampling criteria achieve the target threshold of 95\% in more than 95\% of cases.
+In fact, the pseudorandom sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings.
+In theory, the pseudorandom sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random.
+Based on our experiments, this assumption seems reasonable, and accounts for the higher recall.
+Because the pseudorandom sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data.
+This accounts for the higher work savings.
+
+	\begin{figure}
+		\includegraphics[width=0.9\linewidth]{../images/wss_nrs}
+		\caption{Work saved for the pseudorandom sampling method in each dataset.
+Labels show the number of relevant documents and the total number of documents.
+The datasets are presented in order of the number of documents.
+The whiskers represent the 5th and 95th percentiles.}
+		\label{wss}
+	\end{figure}
+
+ 	\begin{figure*}
+		\centering
+		\begin{subfigure}[b]{0.475\textwidth}
+			\centering
+			\includegraphics[width=\textwidth]{../images/h0_paths_Radjenovic.pdf}
+			\caption[Network2]%Example-Image
+			{{\small Radjenovic}}    
+			\label{fig:Radjenovic}
+		\end{subfigure}
+		\hfill
+		\begin{subfigure}[b]{0.475\textwidth}  
+			\centering 
+			\includegraphics[width=\textwidth]{../images/h0_paths_ProtonBeam.pdf}
+			\caption[]%
+			{{\small ProtonBeam}}    
+			\label{fig:ProtonBeam}
+		\end{subfigure}
+	\vskip\baselineskip
+		\begin{subfigure}[b]{0.475\textwidth}
+			\centering
+			\includegraphics[width=\textwidth]{../images/h0_paths_Statins.pdf}
+			\caption[Network2]%Example-Image
+			{{\small Statins}}    
+			\label{fig:Statins}
+		\end{subfigure}
+		\hfill
+		\begin{subfigure}[b]{0.475\textwidth}  
+			\centering 
+			\includegraphics[width=\textwidth]{../images/h0_paths_Triptans.pdf}
+			\caption[]%
+			{{\small Triptans}}    
+			\label{fig:Triptans}
+		\end{subfigure}
+
+		\caption{\small The path of recall (yellow) and the p-value of H0 for four different datasets} 
+		\label{H0paths}
+	\end{figure*}
+
+	In figure \ref{recall-burden} we rescale the x axis, calling it additional burden, which is simply the work saved when the criterion is triggered subtracted from the work saved when the recall target was actually achieved.
+This measure indicates whether the stopping criterion was triggered too early (positive values), or too late (negative values).
+Here we directly highlight the tradeoffs involved in deciding when to stop screening.
+
+	To help explain the different work savings that were observed in our experiments, we show the distribution of work savings from our pseudorandom criterion for each dataset in figure \ref{wss}.
+In general, higher work savings are possible when the total number of documents is larger.
+However, in datasets with a low proportion of relevant documents, smaller work savings are possible.
+
+	Figure \ref{H0paths} shows the recall and the probability of the null hypothesis for the best performing iteration of four datasets.
+Although the 95\% recall target is achieved very quickly in the Radjenovic dataset, the null hypothesis cannot be excluded until much later.
+This is because the dataset has only 47 relevant documents out of a population of 5,999.
+After the 95\% recall target was achieved, 45 out of 47 relevant documents had been seen and 5,029 documents remained.
+The null hypothesis was therefore that 3 or more of these 5,029 documents were relevant, which requires a lot of evidence to disprove.
+The burden of proof was smaller in the case of the Proton Beam dataset, where the null hypothesis at the point that the 95\% recall threshold was reached was that 13 out of 3,369 remaining documents were relevant.
+
+	The Statins and Triptans datasets show how the criterion performs when the machine learning model has performed poorly in predicting relevant results.
+In each case, 95\% recall is achieved with close to 20\% of documents remaining.
+With fewer documents remaining, it takes less time to rule out the possibility that the number of relevant documents left is greater than that which would be compatible with the achievement of the recall target.
+
 	\section*{Discussion}
+
+	Our results show that it is possible to use machine learning to achieve a given level of recall with a given level of uncertainty.
+The tradeoff for achieving recall reliably is that the work saving achieved is less than the maximum possible work saving.
+However, for large datasets with a significant proportion of relevant documents, the additional effort required to satisfy the criterion will be smaller.
+This makes the approach well suited to broad topics with lots of literature.
+In other words, it is precisely where machine learning will be most useful that the additional effort will be smaller.
+
+	Different use cases for machine learning enhanced screening may also carry different requirements for recall, or different tolerances for uncertainty.
+These can be flexibly accommodated within our stopping criterion.
+Importantly, the ability to make probabilistic statements about the chance of achieving a given recall target makes it possible to clearly communicate the implications of using machine learning enhanced screening to readers and reviewers who are not machine learning specialists.
+This is extremely important in live systematic reviews.
+
+	Our criteria has the further advantage that it is independent of the modelling choice or performance.
+If a machine learning model performs badly at discerning relevant from irrelevant results, the only consequence will be that the work saved will be low.
+With other criteria this may result in poor recall.
+	When using machine learning for screening, poor recall can result in biased results, while low work savings represent no loss to the reviewer as compared to not using machine learning.
+
+	So far, systematic review standards have no way of accommodating screening with machine learning. 
+	We hope that the reliability and clarity of reporting offered by our stopping criteria make them suitable for incorporation into standards, so that machine learning for systematic review screening can fulfil its promise of reducing workload and making more ambitious reviews tractable.
 
@@
 
-	We argue that the stopping criteria proposed here has several important advantages.
-
-	\begin{itemize}
-		\item The accuracy of results is independent of modelling decisions. When a model fits badly, you have to work more, but you don't miss results without knowing it.
-		\item A minimum threshold with uncertainty is understandable and communicable in a live review. It is based solely on arithmetic and simple statistics.
-		\item A minimum threshold better takes into account the undesirability of missed recall.
-	\end{itemize}
+	This paper demonstrates the unsuitability of existing stopping criteria for machine learning approaches to document screening, and proposes a simple method that delivers reliable recall, independent of machine learning approach or model performance.
+Our robust statistical stopping criteria allow users to easily communicate the implications of their use of machine learning, making machine learning enhanced screening ready for live reviews.
 

### 2019-11-19 analysis/run_scenarios_copd_pb.py

@@ -80,7 +80,7 @@ for name, group in df.groupby('review'):
 
 if args.mpi:
     results_df = pd.DataFrame.from_dict(results)
-    results_df.to_csv(f'../results/results_{rank}.csv', index=False)
+    results_df.to_csv(f'../results/results_pb_copd_{rank}.csv', index=False)
 else:                    
     results_df = pd.DataFrame.from_dict(results)
-    results_df.to_csv('../results/results.csv', index=False)
+    results_df.to_csv('../results/results_pb_copd.csv', index=False)

### 2019-11-19 analysis/run_scenarios_fr.py

@@ -57,7 +57,7 @@ for name, group in frdf.groupby('review'):
             stat = MPI.Status()
             r = ss.screen(rank, True)
             if r is not None:
-                results.append(r)
+                pythonresults.append(r)
         else:
             for i in range(iterations):
                 print(i)

### 2019-11-19 manuscript/main.tex
undo typo

@@
 The top bar shows a random distribution of relevant documents (green) and irrelevant documents (red) at a given proportion of relevance.
-The bottom bar  shows distributions of relevant and irrelevant documents in hypothetical sets of seen (right) and unseen (left - transparent) documents.}Jonnalagadda2013
+The bottom bar  shows distributions of relevant and irrelevant documents in hypothetical sets of seen (right) and unseen (left - transparent) documents.}
 	\label{unseen-proportions}

### 2019-11-26 manuscript/main.tex

@@
 %\RequirePackage{natbib}
-%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
+\RequirePackage{natbib}% uncomment this for author-year bibliography
 %\RequirePackage{hyperref}
@@
 			\author[
-			addressref={aff1},
+			addressref={aff1, aff3},
 			email={mueller-hansen@mcc-berlin.net}
@@
 				\cny{United Kingdom}
+			}
+			\address[id=aff3]{%
+				\orgname{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association},
+				\street{P.O. Box 60 12 03},
+				\postcode{D-14412}
+				\city{Potsdam},
+				\cny{Germany}
 			}
@@
 This means that such systems are hard to implement in live reviews.
-				This paper introduces a workflow for working with robust and flexible statistical stopping criteria, that offer real work reductions on the basis of a given confidence level of reaching a given recall.
-				The stopping criteria is shown on test datasets to achieve a reliable level of recall, while still providing consistent work reductions, while other methods proposed are shown to provide inconsistent recall and work reductions across datasets.
+				This paper introduces a workflow with robust and flexible statistical stopping criteria, which offer real work reductions on the basis of a given confidence level of reaching a given recall.
+				The stopping criteria are shown on test datasets to achieve a reliable level of recall, while still providing consistent work reductions, while other methods proposed are shown to provide inconsistent recall and work reductions across datasets.
 
@@
 	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (recall) - before having seen all the documents in the corpus.
-	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved, as this represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.
+	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved.
+This represents the proportion of documents that they do not have to screen, which they would have had to without machine learning.
 
@@
 	We use a simple statistical method to iteratively test a null hypothesis that the recall achieved is less than a given target recall.
+If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
 	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``There is a <5\% chance that we achieve a recall under 95\%''.
@@
 
-	We evaluate our stopping criteria, along with heuristic and sampling based criteria on real-world systematic review datasets on which AL systems have previously been tested \cite{Cohen2006, Yu2019, Terasawa2009, Castaldi2009}.
+	In the remainder of the paper, we first discuss in detail the shortcomings of existing stopping criteria.
+Then, we introduce our new criterion based on a hypergeometric test.
+We evaluate our stopping criteria, and compare their performance with heuristic and sampling based criteria on real-world systematic review datasets on which AL systems have previously been tested \cite{Cohen2006, Yu2019, Terasawa2009, Castaldi2009}.
 
@@
 
+	%\subsection*{Existing Stopping Criteria for Active Learning}
+
 	We start by explaining the sampling and heuristic based stopping criteria and showing their methodological weaknesses.
-Then we introduce our own suggested stopping criteria before testing all criteria on real world datasets.
 
-	\subsection*{Existing Stopping Criteria for Active Learning}
+	\subsection*{Sampling Based Stopping Criteria}
 
-	\subsubsection*{Sampling Based Stopping Criteria}
-
-	The stopping criterion suggested by \cite{Shemilt2014} involves establishing the Baseline Inclusion Rate (BIR), by taking a random sample at the beginning of screening. 
-	This is used to estimate the number of relevant documents in the whole dataset. 
+	The stopping criterion suggested by Shemilt et al. \cite{Shemilt2014} involves establishing the Baseline Inclusion Rate (BIR), by taking a random sample at the beginning of screening. 
+	The BIR is used to estimate the number of relevant documents in the whole dataset. 
 	Reviewers continue to screen until this number, or a proportion of it corresponding to the desired level of recall, is reached.
 
-	However, the estimation of the BIR fails to take into account sampling uncertainty. 
+	However, the estimation of the BIR fails to correctly take into account sampling uncertainty \footnote{Although Shemilt et al.
+\cite{Shemilt2014} employ a method  to choose a sample size based on uncertainty, they fail to acknowledge the potential implications for recall of their choices.
+Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 2,012$ relevant results.
+To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results}.
 	Figure \ref{bir-sampling} shows the predicted number of documents after each document drawn for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant). 
@@
 
-	Where the yellow line is above the above the upper grey line, identification of 95\% of the estimated number of studies would be impossible, and no work would be saved.
-Where it is below, true recall, where recall is estimated to be 95\% would be less than 95\%.
+	Where the yellow line is above the upper grey line, identification of 95\% of the estimated number of studies would be impossible, and no work would be saved.
+In the example, this is the case for about 47\% of random draws of 2000 documents.
+Where it is below the lower grey line, the estimated number of relevant documents is estimated to be less than 95\% of the true number, such that the stopping criterion would give a lower recall than desired.
+For the example, this would occurr in about 32\% of random draws.
+	This shows how even after taking a large sample, baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
 
@@
 The two horizontal grey  lines correspond to the actual number of relevant documents, and to $\frac{100}{0.95}$\% of relevant documents.
-	This shows how even after taking a large sample, baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
 		\label{bir-sampling}
@@
 
-\subsubsection*{Heuristic Based Stopping Criteria}
+\subsubsection*{Heuristic Stopping Criteria}
 
-Some studies give the example of heuristic based stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
+Some studies give the example of heuristic stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
 We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low. 
@@
 We show two scenarios with identical low proportions of relevant documents observed in the unseen documents.
-In the top figure, machine learning (ML) has performed well, and 74\% of documents seen were relevant. 
-In the bottom figure,  ML has performed less well, and only 26\% of documents seen were relevant.
+In the top figure, machine learning (ML) has performed well, and 74\% of the seen documents were relevant. 
+In the bottom figure,  ML has performed less well, and only 26\% of the seen documents were relevant.
 In both cases, only 2\% of unseen documents are relevant, but 2\% of a larger number means more relevant documents are missed.
@@
 
-\cite{Wallace2010a} Develop a ``simple, operational stopping criterion'': stopping after half the documents have been screened.
+Wallace et al.
+\cite{Wallace2010a} develop a ``simple, operational stopping criterion'': stopping after half the documents have been screened.
 Although the criterion worked in their experiment, it is unclear how this could be generalised, and its development depended on knowledge of the true relevance values.
-\cite{Jonnalagadda2013} note that ``the reviewer can elect to end the process of classifying documents at any point, recognizing that stopping before reviewing all documents involves a trade-off of lower recall for reduced workload'', although clearly the reviewer lacks information about probable recall.
-\cite{Yu2019} adopt a more complicated stopping criterion which allows the user to target a specific level of recall.
+Jonnalagadda and Petitti \cite{Jonnalagadda2013} note that ``the reviewer can elect to end the process of classifying documents at any point, recognizing that stopping before reviewing all documents involves a trade-off of lower recall for reduced workload'', although clearly the reviewer lacks information about probable recall.
+Yu and Menzies \cite{Yu2019} adopt a more complicated stopping criterion which allows the user to target a specific level of recall.
 However, reviewers are not given the opportunity to specify a confidence level, and for two of the four datasets in which they tested their criteria, the median achieved recall at a stopping criteria targeting 95\% recall was below 95\%.
-\cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold, rather than above it in a reliable proportion of cases.
+Di Nunzio \cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold, rather than above it in a reliable proportion of cases.
 These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount.
@@
 
-	\subsection*{Random Sampling}
+	\subsubsection*{Random Sampling}
 
@@
 Because we are sampling a binary outcome without replacement, we can use the hypergeometric distribution to formulate a statistical test.
-The hypergeometric distribution tells us the statistical significance of observing $k$ successes (relevant documents) in $n$ draws from a finite population of $N$ documents with $K$ successes.
+The hypergeometric distribution tells us the probability of observing $k$ relevant documents in $n$ draws from a finite population of $N$ documents with $K$ relevant documents.
 
@@
 
-	Recall $R$ is given by the number of relevant documents that have been seen $\dot{\rho}$ over the number of relevant documents in the whole dataset $\rho$
+	Recall $R$ is given by the number of relevant documents that have been seen $\rho_{s}$ over the number of relevant documents in the whole dataset $\rho_{tot}$
 
 	\begin{equation}
-		R = \frac{\dot{\rho}}{\rho}
+		R = \frac{\rho_{seen}}{\rho_{tot}}
 	\end{equation}
 
-	The number of relevant documents in the whole dataset is composed of $\hat{\rho}$ relevant documents seen before random sampling began and $\tilde{K}$ relevant documents unseen at the start of random sampling.
+	The number of relevant documents in the whole dataset is the sum of $\hat{\rho}$ relevant documents seen before random sampling began and $\tilde{K}$ relevant documents unseen at the start of random sampling.
 We can therefore express $R$ as
@@
 
-	We use the cumulative distribution function to estimate the probability $p$ of having observed $k$ or fewer relevant documents in the sample given $\hat{K}$, or, in other words, the probability of observing as many relevant documents in our random sample as we did, if our recall target had not been achieved.
+	We use the cumulative distribution function of the hypergeometric distribution to estimate the probability $p$ of having observed $k$ or fewer relevant documents in the sample given $\hat{K}$.
+This function gives us an upper bound on the probability of observing maximally the number of relevant documents in our random sample as we did, if our recall target had not been achieved.
 	If this is below our confidence level $1 - \alpha$, we can reject the null hypothesis that the recall target was not achieved.
@@
 
-	\subsection*{Pseudo-random sampling}
+	\subsubsection*{Pseudo-random sampling}
+
 	In order to decide when to begin a random sample, we employ pseudo-random sampling, where we treat previously screened documents as a random sample.
@@
 
-	\subsection*{Analysis}
+	%\subsection*{Analysis}
 %	
@@
 
-	We evaluate each of the criteria discussed, operationalising the heuristic stopping criteria with 50, 100, and 200 consecutive irrelevant records, on real world test data.
-We run 100 iterations on each dataset and record
+	We evaluate each of the criteria discussed on real world test data, operationalising the heuristic stopping criterion with 50, 100, and 200 consecutive irrelevant records.
+We run 100 iterations on each dataset and record the following measures.
 	\begin{itemize}
-		\item \textbf{Actual Recall}: The recall when the stopping criteria was met
-		\item \textbf{WS-SC}: Work saved when the stopping criteria was met \footnote{We do not use work saved over sampling, is it is not axiomatic that seeing 95\% of a random sample would achieve 95\% recall}
+		\item \textbf{Actual Recall}: The recall when the stopping criterion was met
+		\item \textbf{WS-SC}: Work saved when the stopping criterion was met \footnote{We do not use work saved over sampling, it is not axiomatic that seeing 95\% of a random sample would achieve 95\% recall}
 		\item \textbf{Additional Burden}: the work saved when the criterion was triggered subtracted from the work saved when the recall target was actually achieved.
@@
 We start with random samples of 200 documents.
-Subsequently, we ``screen'', that is, we reveal the labels of, batches of the 20 documents with the highest predicted relevance scores, retraining the model after each batch.
+Subsequently, we ``screen'', that is, we reveal the labels of batches of the 20 documents with the highest predicted relevance scores, retraining the model after each batch.
 Each criterion is evaluated after each document is ``screened''.
@@
 
-	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criteria has been satisfied. 
+	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criterion has been satisfied. 
 	For comparison, we also include the results that would have been achieved with \textit{a priori} knowledge of the data, that is, the work saved when the 95\% recall target was actually reached.
@@
 
-	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
+	The baseline sampling criterion (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
 This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.
@@
 
-	\caption{\small Distribution of recall and work saved after each stopping criteria.
-Green dots show results for datasets with less than 1,000 documents, orange dots show datasets with 1,000 - 2,000 documents, and blue dots show datasets with more than 2,000 documents}
+	\caption{\small Distribution of recall and work saved after each stopping criterion.
+Green dots show results for datasets with less than 1,000 documents, orange dots show datasets with 1,000 - 2,000 documents, and blue dots show datasets with more than 2,000 documents.}
 	\label{recall-wss}
@@
 
-	\caption{\small Distribution of recall and additional burden after each stopping criteria.
-Additional burden is the work saved when the criteria was trigerred subtracted from the work saved when the target was reached.}
+	\caption{\small Distribution of recall and additional burden after each stopping criterion.
+Additional burden is the work saved when the criterion was triggered minus the work saved when the target was reached.
+Coloring of data points as in Fig.
+\ref{recall-wss}.}
 	\label{recall-burden}
@@
 
-	The Heuristic based stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
+	The Heuristic stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
 Although the mean work saved for IH50 is 41\%, the target is missed in 39\% of cases.
@@
 This measure indicates whether the stopping criterion was triggered too early (positive values), or too late (negative values).
-Here we directly highlight the tradeoffs involved in deciding when to stop screening.
+The figure directly highlights the tradeoffs involved in deciding when to stop screening.
 
@@
 In general, higher work savings are possible when the total number of documents is larger.
-However, in datasets with a low proportion of relevant documents, smaller work savings are possible.
+However, in datasets with a low proportion of relevant documents, many documents need to be screened to achieve a high confidence that there are only few relevant documents remaining in the unseen ones.
+Therefore, smaller work savings are possible.
 
@@
 The null hypothesis was therefore that 3 or more of these 5,029 documents were relevant, which requires a lot of evidence to disprove.
-The burden of proof was smaller in the case of the Proton Beam dataset, where the null hypothesis at the point that the 95\% recall threshold was reached was that 13 out of 3,369 remaining documents were relevant.
+The burden of proof was smaller in the case of the Proton Beam dataset: at the point that the 95\% recall threshold was reached, the null hypothesis to disprove was that 13 out of 3,369 remaining documents were relevant.
 
@@
 In each case, 95\% recall is achieved with close to 20\% of documents remaining.
-With fewer documents remaining, it takes less time to rule out the possibility that the number of relevant documents left is greater than that which would be compatible with the achievement of the recall target.
+With fewer documents remaining, it takes less screening to rule out the possibility that the number of relevant documents left is incompatible with the achievement of the recall target.
 
@@
 
-	Our results show that it is possible to use machine learning to achieve a given level of recall with a given level of uncertainty.
+	Our results show that it is possible to use machine learning to achieve a given level of recall with a given level of confidence.
 The tradeoff for achieving recall reliably is that the work saving achieved is less than the maximum possible work saving.
-However, for large datasets with a significant proportion of relevant documents, the additional effort required to satisfy the criterion will be smaller.
+However, for large datasets with a significant proportion of relevant documents, the additional effort required to satisfy the criterion will be small.
 This makes the approach well suited to broad topics with lots of literature.
-In other words, it is precisely where machine learning will be most useful that the additional effort will be smaller.
+In other words, it is precisely where machine learning will be most useful that the additional effort will be small.
 
@@
 
-	Our criteria has the further advantage that it is independent of the modelling choice or performance.
-If a machine learning model performs badly at discerning relevant from irrelevant results, the only consequence will be that the work saved will be low.
+	Our criterion has the further advantage that it is independent of the choice or performance of the machine learning model.
+If a model performs badly at discerning relevant from irrelevant results, the only consequence will be that the work saved will be low.
 With other criteria this may result in poor recall.
@@
 	So far, systematic review standards have no way of accommodating screening with machine learning. 
-	We hope that the reliability and clarity of reporting offered by our stopping criteria make them suitable for incorporation into standards, so that machine learning for systematic review screening can fulfil its promise of reducing workload and making more ambitious reviews tractable.
+	We hope that the reliability and clarity of reporting offered by our stopping criterion make them suitable for incorporation into standards, so that machine learning for systematic review screening can fulfil its promise of reducing workload and making more ambitious reviews tractable.
 
@@
 	\section*{Author's contributions}
-	Text for this section \ldots
+	MC designed the research and conducted the experiments. FMH contributed to the development of the statistical basis for the stopping criterion. Both authors wrote and edited the manuscript.
 
@@
 
-	\bibliography{../mendeley}
+	\bibliography{mendeley}
 	\bibliographystyle{vancouver}

### 2019-11-28 manuscript/main.tex

@@
 	This is particularly problematic in systematic reviews because low recall increases the risk of bias \cite{Lefebvre2011}.
-	The lack of appropriate stopping criteria has therefore been identified as a research gap \cite{bannach-brown2019}, although some approaches have been suggested.
+	The lack of appropriate stopping criteria has therefore been identified as a research gap \cite{bannach-brown2019, Marshall2019}, although some approaches have been suggested.
 These fall into the following categories:
@@
 \cite{Shemilt2014} employ a method  to choose a sample size based on uncertainty, they fail to acknowledge the potential implications for recall of their choices.
-Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 2,012$ relevant results.
-To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results}.
+Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 251$ relevant results.
+To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results.
+See the notebook https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb that accompanies this paper for a detailed discussion}.
 	Figure \ref{bir-sampling} shows the predicted number of documents after each document drawn for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant). 
@@
 We show two scenarios with identical low proportions of relevant documents observed in the unseen documents.
-In the top figure, machine learning (ML) has performed well, and 74\% of the seen documents were relevant. 
-In the bottom figure,  ML has performed less well, and only 26\% of the seen documents were relevant.
+In the top figure, machine learning (ML) has performed well, and 74\% of the screened documents were relevant. 
+In the bottom figure,  ML has performed less well, and only 26\% of the screened documents were relevant.
 In both cases, only 2\% of unseen documents are relevant, but 2\% of a larger number means more relevant documents are missed.
@@
 	We use the cumulative distribution function of the hypergeometric distribution to estimate the probability $p$ of having observed $k$ or fewer relevant documents in the sample given $\hat{K}$.
-This function gives us an upper bound on the probability of observing maximally the number of relevant documents in our random sample as we did, if our recall target had not been achieved.
+This function gives us an upper bound on the probability of observing no more than the number of relevant documents in our random sample that we did, if our recall target had not been achieved.
 	If this is below our confidence level $1 - \alpha$, we can reject the null hypothesis that the recall target was not achieved.
@@
 
-	We evaluate each of the criteria discussed on real world test data, operationalising the heuristic stopping criterion with 50, 100, and 200 consecutive irrelevant records.
+	We evaluate each of the criteria discussed on real world test data, operationalising the heuristic stopping criteria with 50, 100, and 200 consecutive irrelevant records.
 We run 100 iterations on each dataset and record the following measures.
 	\begin{itemize}
-		\item \textbf{Actual Recall}: The recall when the stopping criterion was met
-		\item \textbf{WS-SC}: Work saved when the stopping criterion was met \footnote{We do not use work saved over sampling, it is not axiomatic that seeing 95\% of a random sample would achieve 95\% recall}
+		\item \textbf{Actual Recall}: The recall when the stopping criteria was met
+		\item \textbf{WS-SC}: Work saved when the stopping criteria was met
 		\item \textbf{Additional Burden}: the work saved when the criterion was triggered subtracted from the work saved when the recall target was actually achieved.
@@
 	For simplicity, we use a basic SVM model \cite{Cortes95, Pedregosa2011}, with 1-2 word n-grams taken from the document abstracts used as input data.
-We start with random samples of 200 documents.
-Subsequently, we ``screen'', that is, we reveal the labels of batches of the 20 documents with the highest predicted relevance scores, retraining the model after each batch.
+We start with random samples of 200 documents (we do not employ Shemilt et al's methods for identifying the ``optimal'' sample size, as we showed these in the methods section to be unhelpful).
+Subsequently, we ``screen'', that is, we reveal the labels of, batches of the 20 documents with the highest predicted relevance scores, retraining the model after each batch.
 Each criterion is evaluated after each document is ``screened''.
@@
 
-	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criterion has been satisfied. 
+	Figure \ref{recall-wss} shows the actual recall and work savings achieved when each stopping criteria has been satisfied. 
 	For comparison, we also include the results that would have been achieved with \textit{a priori} knowledge of the data, that is, the work saved when the 95\% recall target was actually reached.
@@
 
-	The baseline sampling criterion (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
+	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
 This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.
@@
 
-	\caption{\small Distribution of recall and work saved after each stopping criterion.
+	\caption{\small Distribution of recall and work saved after each stopping criteria.
 Green dots show results for datasets with less than 1,000 documents, orange dots show datasets with 1,000 - 2,000 documents, and blue dots show datasets with more than 2,000 documents.}

### 2019-11-28 manuscript/main.tex

@@
 	\begin{figure}
-		\includegraphics[width=0.5\linewidth]{../images/flow}
+		\includegraphics[width=0.7\linewidth]{../images/flow}
 		\caption{A workflow for active learning in screening with a statistical stopping criterion}
@@
 		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_hyper}
+		\caption[]%
+		{{\small Hypergeometric sampling \\}}    
+		\label{fig:hyper}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_nrs}
+		\caption[]%
+		{{\footnotesize Pseudorandom hypergeometric sampling}}    
+		\label{fig:nrs}
+	\end{subfigure}
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}
+		\centering
 		\includegraphics[width=\textwidth]{../images/jointplot_bir}
@@
 		\label{fig:ih_50}
-	\end{subfigure}
-	\vskip\baselineskip
-	\begin{subfigure}[b]{0.475\textwidth}   
-		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_hyper}
-		\caption[]%
-		{{\small Hypergeometric sampling \\}}    
-		\label{fig:hyper}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[b]{0.475\textwidth}   
-		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_nrs}
-		\caption[]%
-		{{\footnotesize Pseudorandom hypergeometric sampling}}    
-		\label{fig:nrs}
 	\end{subfigure}
@@
 		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_hyper}
+		\caption[]%
+		{{\small Hypergeometric sampling \\}}    
+		\label{fig:hyper_ab}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/jointplot_burden_nrs.pdf}
+		\caption[]%
+		{{\footnotesize Pseudorandom hypergeometric sampling}}    
+		\label{fig:nrs_ab}
+	\end{subfigure}
+	\hfill
+	\vskip\baselineskip
+	\begin{subfigure}[b]{0.475\textwidth}
+		\centering
 		\includegraphics[width=\textwidth]{../images/jointplot_burden_bir}
@@
 	\vskip\baselineskip
-	\begin{subfigure}[b]{0.475\textwidth}   
-		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_hyper}
-		\caption[]%
-		{{\small Hypergeometric sampling \\}}    
-		\label{fig:hyper_ab}
-	\end{subfigure}
-	\hfill
+
 	\begin{subfigure}[b]{0.475\textwidth}   
@@
 
-	\end{subfigure}
-	\vskip\baselineskip
-	\begin{subfigure}[b]{0.475\textwidth}   
-		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_nrs.pdf}
-		\caption[]%
-		{{\footnotesize Pseudorandom hypergeometric sampling}}    
-		\label{fig:nrs_ab}
 	\end{subfigure}
@@
 
-	The Heuristic stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
-Although the mean work saved for IH50 is 41\%, the target is missed in 39\% of cases.
-The cases below the horizontal grey line indicate instances where work has been saved at the expense of achieving the recall target.
-
 	Both the random sampling and the pseudorandom sampling criteria achieve the target threshold of 95\% in more than 95\% of cases.
-In fact, the pseudorandom sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings.
+In fact, the pseudorandom sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases.
 In theory, the pseudorandom sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random.
@@
 This accounts for the higher work savings.
+
+	The Heuristic stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
+Although the mean work saved for IH50 is 41\%, the target is missed in 39\% of cases.
+The cases below the horizontal grey line indicate instances where work has been saved at the expense of achieving the recall target.
 
@@
 The datasets are presented in order of the number of documents.
-The whiskers represent the 5th and 95th percentiles.}
+The whiskers represent the 5th and 95th percentiles.
+The grey line shows work savings of 5\%.
+}
 		\label{wss}
@@
 The null hypothesis was therefore that 3 or more of these 5,029 documents were relevant, which requires a lot of evidence to disprove.
-The burden of proof was smaller in the case of the Proton Beam dataset: at the point that the 95\% recall threshold was reached, the null hypothesis to disprove was that 13 out of 3,369 remaining documents were relevant.
+The burden of proof was smaller in the case of the Proton Beam dataset: at the point that the 95\% recall threshold was reached, the null hypothesis to disprove was that a minimum of 13 out of 3,369 remaining documents were relevant.
 
@@
 In each case, 95\% recall is achieved with close to 20\% of documents remaining.
-With fewer documents remaining, it takes less screening to rule out the possibility that the number of relevant documents left is incompatible with the achievement of the recall target.
+With fewer documents remaining, it takes fewer screening decisions to rule out the possibility that the number of relevant documents left is incompatible with the achievement of the recall target.
 
@@
 
-	Our criterion has the further advantage that it is independent of the choice or performance of the machine learning model.
+	Our criteria have the further advantage that they are independent of the choice or performance of the machine learning model.
 If a model performs badly at discerning relevant from irrelevant results, the only consequence will be that the work saved will be low.
@@
 	So far, systematic review standards have no way of accommodating screening with machine learning. 
-	We hope that the reliability and clarity of reporting offered by our stopping criterion make them suitable for incorporation into standards, so that machine learning for systematic review screening can fulfil its promise of reducing workload and making more ambitious reviews tractable.
+	We hope that the reliability and clarity of reporting offered by our stopping criteria make them suitable for incorporation into standards, so that machine learning for systematic review screening can fulfil its promise of reducing workload and making more ambitious reviews tractable.
 

### 2019-11-28 manuscript/main.tex
Updates from Overleaf

@@
 See the notebook https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb that accompanies this paper for a detailed discussion}.
+	This uncertainty is crucial, as errors can have severe consequences.
+If the estimated number of relevant documents is even one unit above the true value, then no work savings will be achieved.
+If the
 	Figure \ref{bir-sampling} shows the predicted number of documents after each document drawn for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant). 
@@
 
-	\begin{figure}
-	\includegraphics[width=\linewidth]{../images/bir_sampling_noci.png}
-	\caption{Uncertainty in estimating the baseline inclusion rate.
-For 50 random samples of 2,000 documents from 20,000 documents, we show the estimated number of relevant documents, after each document is drawn.
-The two horizontal grey  lines correspond to the actual number of relevant documents, and to $\frac{100}{0.95}$\% of relevant documents.
-		\label{bir-sampling}
-	}
-\end{figure}
+	\begin{figure*}
+	\centering
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/bir_errors}
+		\caption[]%
+		{{\small Error types after 1,000 documents \\}}    
+		\label{fig:bir_error}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.475\textwidth}   
+		\centering 
+		\includegraphics[width=\textwidth]{../images/bir_error_distribution}
+		\caption[]%
+		{{\footnotesize Error types across sample sizes}}    
+		\label{fig:bir_error_distribution}
+	\end{subfigure}
+
+	\caption{\small Distribution of recall and work saved after each stopping criteria.
+Green dots show results for datasets with less than 1,000 documents, orange dots show datasets with 1,000 - 2,000 documents, and blue dots show datasets with more than 2,000 documents.}
+	\label{recall-wss}
+\end{figure*}
 

### 2019-11-29 manuscript/main.tex

@@
 
-			\title{Statistical Stopping Criteria for Active Learning in Systematic Review Screening}
+			\title{Robust Statistical Stopping Criteria for Automated Screening in Systematic Reviews}
 
@@
 If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
-	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``There is a <5\% chance that we achieve a recall under 95\%''.
+	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of 95\%''.
 
@@
 	However, the estimation of the BIR fails to correctly take into account sampling uncertainty \footnote{Although Shemilt et al.
-\cite{Shemilt2014} employ a method  to choose a sample size based on uncertainty, they fail to acknowledge the potential implications for recall of their choices.
-Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 251$ relevant results.
+\cite{Shemilt2014} employ a method  to choose a sample size based on uncertainty, they fail to acknowledge the potential implications for recall of their choice.
+Their margin of error of 0.0025 and observed proportion of relevant studies of 0.0005 translate to estimates of $400 \pm 451$ relevant results.
 To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results.
-See the notebook https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb that accompanies this paper for a detailed discussion}.
+See the notebook \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb} that accompanies this paper for a detailed discussion}.
 	This uncertainty is crucial, as errors can have severe consequences.
 If the estimated number of relevant documents is even one unit above the true value, then no work savings will be achieved.
-If the
-	Figure \ref{bir-sampling} shows the predicted number of documents after each document drawn for 50 random samples of 2,000 documents from 20,000 documents (where 5\% of documents are relevant). 
-	Depending on the luck of the draw, one might estimate the true number of documents as being
-	\begin{enumerate}
-		\item much higher than the true value, meaning that the stopping criterion would never be reached - and no work could be saved, or;
-		\item much lower than the true value, meaning that the stopping criterion would be reached before the desired level of recall was actually achieved
-	\end{enumerate}
+If the number of relevant documents is underestimated, then the recall achieved will be less than 100\%.
 
-	Where the yellow line is above the upper grey line, identification of 95\% of the estimated number of studies would be impossible, and no work would be saved.
-In the example, this is the case for about 47\% of random draws of 2000 documents.
-Where it is below the lower grey line, the estimated number of relevant documents is estimated to be less than 95\% of the true number, such that the stopping criterion would give a lower recall than desired.
-For the example, this would occurr in about 32\% of random draws.
-	This shows how even after taking a large sample, baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
+	The number of relevant documents drawn without replacement from a finite sample of documents follows the hypergeometric distribution. 
+	Figure \ref{fig:bir_error} shows the distribution of the predicted number of documents after drawing 1,000 documents from a total of 20,000 documents, where 500 documents (2.5\%) are relevant.
+The left shaded portion of the graph shows all the cases where the recall will be less than 95\%.
+This occurs 35.91\% of the time.
+The right shaded portion of the graph shows the cases where the number of relevant documents is overestimated and no work savings could be made.
+This occurs 46.24\% of the time.
+In only 17.85\% of cases can work savings be achieved while still achieving a recall of at least 95\%.
+
+	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size. Even in very large samples both types of error remain frequent).
+	This shows how baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
 
@@
 
-	\caption{\small Distribution of recall and work saved after each stopping criteria.
-Green dots show results for datasets with less than 1,000 documents, orange dots show datasets with 1,000 - 2,000 documents, and blue dots show datasets with more than 2,000 documents.}
-	\label{recall-wss}
+	\caption{\small Distribution of under- or over-estimation errors using the BIR sampling method in a dataset of 20,000 documents of which 500 are relevant.
+Panel (a) shows the probability distribution of the estimated number of relevant documents after a sample of 1,000 documents.
+Panel (b) shows the probability of each type of error according to the sample size.}
+	\label{bir-sampling}
 \end{figure*}
@@
 
+	All computational steps required to reproduce this analysis are documented online at \url{https://github.com/mcallaghan/rapid-screening}.
+
 	\section*{Results}
@@
 In a live systematic review, reviewers would never know when this had been reached, but these are the work savings most often reported in machine learning for systematic review screening studies.
-
-	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
-This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.
 
@@
 
+	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
+This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.
+
 	The Heuristic stopping criteria, both for 50 consecutive irrelevant results (Figure \ref{fig:ih_50} - IH50), and for 200 irrelevant results (Figure \ref{fig:ih_200}) also perform unreliably.
@@
 
-	In figure \ref{recall-burden} we rescale the x axis, calling it additional burden, which is simply the work saved when the criterion is triggered subtracted from the work saved when the recall target was actually achieved.
-This measure indicates whether the stopping criterion was triggered too early (positive values), or too late (negative values).
-The figure directly highlights the tradeoffs involved in deciding when to stop screening.
+	In figure \ref{recall-burden} we rescale the x axis, calling it additional burden, which is simply the work saved when the criterion is triggered minus the work saved when the recall target was actually achieved.
+This measure indicates whether the stopping criterion was triggered too early (negative values), or too late (positive values).
+The figure directly highlights the tradeoffs involved in deciding when to stop screening: For our criteria, there is mostly a small additional burden which comes with the necessity to make sure the desired recall target has been reached and reject the null hypothesis that this has not been the case.
+For the other criteria, there are many cases in which additional burden is negative, i.e.
+the criterion has been triggered too early.
+In these cases, however, the desired recall is hardly ever reached.
 
@@
 The tradeoff for achieving recall reliably is that the work saving achieved is less than the maximum possible work saving.
-However, for large datasets with a significant proportion of relevant documents, the additional effort required to satisfy the criterion will be small.
+However, for large datasets with a significant proportion of relevant documents, the additional effort required to satisfy the criterion will be small compared to the work saved by using machine learning.
 This makes the approach well suited to broad topics with lots of literature.
@@
 	\section*{Acknowledgements}
-	Max Callaghan is supported by a PhD scholarship from the Heinrich Böll Foundation
+	Max Callaghan is supported by a PhD scholarship from the Heinrich Böll Foundation. Finn M\"{u}ller-Hansen is supported by the BMBF project Strategic Scenario Analysis (START).
 

### 2019-12-02 manuscript/main.tex
final comments merged before submission

@@
 				This paper introduces a workflow with robust and flexible statistical stopping criteria, which offer real work reductions on the basis of a given confidence level of reaching a given recall.
-				The stopping criteria are shown on test datasets to achieve a reliable level of recall, while still providing consistent work reductions, while other methods proposed are shown to provide inconsistent recall and work reductions across datasets.
+				The stopping criteria are shown on test datasets to achieve a reliable level of recall, while still providing work reductions of on average 17\%.
+Other methods proposed previously are shown to provide inconsistent recall and work reductions across datasets.
 
@@
 
-	We show in this paper how existing criteria are inadequate.
-	We demonstrate with several previously used datasets that their inadequacy lies in the unreliability - particularly across different domains, or datasets with different properties \cite{OMara-Eves2015} - both of the work saved and the recall achieved.
+	We review these methods in the following section and discuss their theoretical limitations. They are then tested on several previous systematic review datasets.
+	We demonstrate theoretically and with our experimental results, that these methods can not deliver consistent levels of work savings or recall - particularly across different domains, or datasets with different properties \cite{OMara-Eves2015}.
 	Without the reliable or reportable achievement of a desired level of recall, deployment of AL systems in live reviews remains challenging.
@@
 
-	\section*{Methods}
+	\section*{Methods Review}
 
-	%\subsection*{Existing Stopping Criteria for Active Learning}
-
-	We start by explaining the sampling and heuristic based stopping criteria and showing their methodological weaknesses.
+	We start by explaining the sampling and heuristic based stopping criteria and discussing their methodological limitations.
 
@@
 
+\subsection*{Methods}
+
 \subsection*{A Statistical Stopping Criterion for Active Learning}
@@
 
-	%%%%%%%%%%%
-%	
-%	We calculate a hypothetical value $\hat{K}$
-%
-%	Our aim is to formulate a stopping criterion such that a recall of $\tau$ has been reached with confidence level $\alpha$.
-%	To achieve this, we formulate a hypothesis test:
-%	We first calculate the number $\hat{n}$ of relevant documents that can maximally remain in the unseen documents to achieve a given recall target.
-The recall is defined as the ratio of seen relevant documents over all relevant documents
-%
-%	with the total number of seen relevant documents $\rho$, the number of seen relevant documents at the time we start random sampling $\hat{\rho}$, and the remaining number of relevant documents in the unseen sample at the time random sampling starts $\tilde{n}$.
-%	Reorganization of this equation gives
-%	\begin{equation}
-%		\tilde{n} = \frac{\rho}{\tau} - \hat{\rho}.
-%	\end{equation}
-%	Because we need to use integers, we get the next largest integer given by the ceiling function $\lceil \cdot \rceil$:
-%
-%	The null-hypothesis to reject can then be formulated as:
-%	\begin{quote}
-%	$H_0$: There are $\hat{n}$ or more relevant documents remaining in the set of unseen documents.
-%	\end{quote}
-%	
-%	% notation here follows notation for scipy description of hypergeometric distribution
-%	% better use another notation more in line with the literature? -> literature seems to use very different notations
-%	
-%	In theory, we draw a random sample of $N$ documents, evaluate how many of them are relevant, and denote this number with $k$.
-We can estimate the probability that we observe $k$ or fewer relevant documents in a random draw of $N$ documents without replacement from a pool of $M$ documents, of which $n$ are relevant using the cumulative distribution function of the hypergeometric distribution.
-%This function is defined as the sum over the probability mass function of the hypergeometric distribution $p(x, M, n, N)$:
-%	\begin{equation}	
-%		 cdf(k, M, n, N) = \sum_{x \leq k} p(x, M, n, N) = \sum_{x \leq k} \frac{\binom{n}{x} \binom{M - n}{N - x}} {\binom{M}{N}},
-%		 \label{pfunc}
-%	\end{equation}
-%	where $\binom{a}{b}$ denote binomial coefficients. In our case, $M$ is the number of unseen documents before we start random sampling.
-%	
-%	We can reject the null hypothesis if the p-value $p = cdf(k,M,\hat{n},N)$ is smaller than $1 - \alpha$ [references to hypergeometric testing?, needs further argumentation?].
-Because we are testing for the lower tail, this implies that we can be confident with a level $\alpha$ that the number of unseen relevant documents is lower than $\hat{n}$.
-%	
-%	In practice, we can draw random unseen documents repeatedly. After each draw, we calculate the p-value using the number of relevant documents $k$ and the number of randomly sampled documents $N$.
-%	We can stop the screening once a p-value $p < 1 - \alpha$ is reached.
-
 	\subsubsection*{Pseudo-random sampling}
@@
 	\end{figure}
-
-	%\subsection*{Analysis}
-%	
-%	\begin{table}
-%		\begin{tabular}{p{0.15\linewidth} p{0.425\linewidth} p{0.425\linewidth}}
-%			Parameter & Description & Estimation \\
-%			\hline
-%			$ P  $ &  Total number of studies & \textit{Observed} \\
-%			$ S $ & Number of studies coded by humans & \textit{Observed} \\
-%			$ U $ & Number of studies not yet coded by humans & \textit{Observed} ($P - S$) \\
-%			$ \alpha $ & Acceptable uncertainty level & \textit{Given} \\
-%			$ \tau $ & Target recall threshold & \textit{Given} \\
-%			$ k $ & Number of relevant documents drawn & \\
-%			$ N $ & Number of randomly drawn documents & \textit{Observed} \\
-%			$ M $ & Number of remaining unseen documents before sampling & \\
-%			$ n $ & Number of relevant documents in sample & \\
-%			$ \hat{n} $ & Minimum number of relevant documents in sample that would mean the recall threshold had been reached  & \\
-%		\end{tabular}
-%		\caption{Parameters}
-%		\label{parameters}
-%	\end{table}
-
-%
-%	
-%	Table \ref{parameters} shows the parameters, known, estimated, and given, available during the random sampling process and required to estimate recall. 
-%	Expanding on what was stated before, recall - at a given confidence threshold is a function of 1) the upper estimate of the relevance of remaining documents, 2) The estimated relevance of all documents in the dataset, 3) The proportion of documents not yet seen.
-%	
-%	Figure \ref{sample-recall} shows the minimum estimated recall for a set of confidence intervals, along with the actual recall (in grey) for a case where 800 out of 2,000 documents have been reviewed, 2\% of remaining documents are relevant, and 20\% of all documents are relevant.
-%	We see that after 95\% recall has actually been achieved, but before 100\% of documents have been been seen, we can be confident at each of the given confidence levels that 95\% of relevant documents have been identified.
-%	All estimates of recall only reach 100\% after all documents have been seen, as it is not possible to exclude the possibility, at any given confidence level, that the proportion of relevant documents is greater than 0 \todo{Actually maybe this is possible if we work with whole numbers?
-How does that change things?}.
-%	Figure \ref{sample-recall-variation} shows the various trajectories of a 95\% minimum recall (blue) and actual recall (grey) given the same paramaters for 200 random samples.
-%
-%	That the blue lines consistently meet a 95\% threshold to the right of the grey lines indicates that reviewers have to see more than 95\% of relevant documents in order to be confident that they have achieved their threshold.
-%	We call the number of documents it is necessary to review to establish with confidence that reviewers have achieved 95\% recall after they have already achieved 95\% recall the \textit{additional burden}.
-%	
-%	In figure \ref{additional-burden}, we investigate how the total number of documents and the proportion of relevant documents affect the additional burden \todo{Use the number of documents seen instead of p?}.
-We demonstrate the additional burden decreases \todo{how?
-Non-linearly?} with increasing total numbers of documents.
-With decreasing relevance of the remaining documents (holding the recall constant at 95\%, this means that fewer documents remain unseen), the additional burden increases.
-This implies that where the machine learning has more successfully identified documents and reduced burden, the additional burden required to establish the level of recall is higher.
-This is because the proportion of documents not seen is greater, so uncertainty around the relevance of that proportion has a greater impact on the estimation of recall.
-%
-%	When to move to random sampling remains an open question.
-Our current approach is to use the predicted relevance of the machine learning model as a guide.
-We calculate the [minimum/maximum] estimated recall based on observing the predicted relevance level in [x] random documents.
-If the estimated recall would be below the threshold, then we begin drawing random documents.
-%
-%	The stopping criteria explained here is
 
@@
 	\section*{Acknowledgements}
-	Max Callaghan is supported by a PhD scholarship from the Heinrich Böll Foundation. Finn M\"{u}ller-Hansen is supported by the BMBF project Strategic Scenario Analysis (START).
+	Max Callaghan is supported by a PhD scholarship from the Heinrich Böll Foundation.
+Finn M\"{u}ller-Hansen acknowledges funding from the German Federal Ministry of Research and Education within the Strategic Scenario Analysis (START) project (grant reference: 03EK3046B).
 

### 2019-12-17 images/flow.tex

@@
 	},
+	end/.style={
+		base,
+		shape=ellipse,
+		fill={rgb:red,1;green,2;blue,5},
+		text=white
+	}
 }
@@
   \node (screenrel)  [screen, below of =sample]  {Screen most \\ relevant documents};
-  \node (stopscreen) [decision, join] {Irrelevant?};
+  \node (stopscreen) [decision, join] {Recall unlikely $<$ threshold?};
 
@@
 
-  \node (recall) [decision, join] {Recall $>$ threshold?};
+  \node (recall) [decision, join] {Recall unlikely $<$ threshold?};
 
-    	\node (e1) [form, right=of stopscreen, xshift=-15mm] { $p < 1-\frac{\alpha}{2}$ };
+    	\node (e1) [form, right=of stopscreen, xshift=-10mm] { $p < 1-\frac{\alpha}{2}$ };
 
@@
 
-  \node (rform) [form, right=of recall, xshift=-15mm] { $p < 1-\alpha $ };
+  \node (rform) [form, right=of recall, xshift=-10mm] { $p < 1-\alpha $ };
 
-  \node (reldocs) [multidoc, below of=recall, xshift=20mm, yshift=-8mm]  {Relevant documents};
-  \node (irreldocs) [multidoc, fill=blue!5, below of=recall, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
+  \node (stop) [end, below of=recall, yshift=-10mm] {Stop};
+
+  \node (reldocs) [multidoc, below of=stop, xshift=20mm, yshift=-8mm]  {Relevant documents};
+  \node (irreldocs) [multidoc, fill=blue!5, below of=stop, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
   %\node
@@
 
- \path (stopscreen.west) to node [very near start, yshift=1em] {$n$} (c1);
+ \path (stopscreen.west) to node [very near start, yshift=1em] {$no$} (c1);
 
@@
 
-  \path (recall.west) to node [very near start, yshift=1em] {$n$} (c3);
+  \path (recall.west) to node [very near start, yshift=1em] {$no$} (c3);
+
+    \path (stopscreen.south) to node [very near start, yshift=-0.5em, xshift=1em] {$yes$} (screenr2.north);
+
+    \path (recall.south) to node [very near start, yshift=-0.5em, xshift=1em] {$yes$} (stop.north);
 
@@
 
- \draw [o->,norm] (recall.south) -- (reldocs);
- \draw [o->,norm] (recall.south) -- (irreldocs);
+ \draw [o->,norm] (recall.south) -- (stop);
+
+ \draw [o->,norm] (stop.south) -- (reldocs);
+ \draw [o->,norm] (stop.south) -- (irreldocs);
 

### 2020-06-11 manuscript/main.tex
Revisions d0

@@
 %\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass{bmcart}
+\documentclass[linenumbers]{bmcart}
 
@@
 	The algorithm chooses which studies will next be screened by humans, often those which are likely to be relevant or about which the model is uncertain, in order to generate more labels to feed back to the machine.
-	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (recall) - before having seen all the documents in the corpus.
+	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (described by recall: the number of relevant studies identified divided by the total number of relevant studies) - before having seen all the documents in the corpus.
 	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved.
@@
 	The lack of appropriate stopping criteria has therefore been identified as a research gap \cite{bannach-brown2019, Marshall2019}, although some approaches have been suggested.
-These fall into the following categories:
+These have most commonly fallen into the following categories:
 	\begin{itemize}
@@
 		\item \textbf{Pragmatic criteria:} Reviewers stop when they run out of time \cite{miwa2014}. 
+		\item \textbf{Novel automatic stopping criteria:} Recent papers have proposed more complicated novel systems for automatically deciding when to stop screening \cite{Yu2019, DiNunzio2018, Howard2020}
 	\end{itemize}
 
-	We review these methods in the following section and discuss their theoretical limitations. They are then tested on several previous systematic review datasets.
-	We demonstrate theoretically and with our experimental results, that these methods can not deliver consistent levels of work savings or recall - particularly across different domains, or datasets with different properties \cite{OMara-Eves2015}.
+	We review the first three classes of these methods in the following section and discuss their theoretical limitations. They are then tested on several previous systematic review datasets.
+	We demonstrate theoretically and with our experimental results, that these three classes of methods can not deliver consistent levels of work savings or recall - particularly across different domains, or datasets with different properties \cite{OMara-Eves2015}.
+We also discuss the limitations of novel automatic stopping criteria, which have all demonstrated promising results, but do not achieve a given level of recall in a reliable or reportable way.
 	Without the reliable or reportable achievement of a desired level of recall, deployment of AL systems in live reviews remains challenging.
@@
 	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of 95\%''.
-
-	The information retrieval literature discusses similar stopping criteria for ranking algorithms like BM25 and variants \cite{DiNunzio2018, Yu2019}.
-However, the estimators they use to determine the recall rely on the specific ranking functions and depend on their search input.
-Therefore, the quality of the estimation depends on the adequacy of the model.
-Our approach, on the contrary, is independent of model choice or model performance.
 
@@
 
-	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size. Even in very large samples both types of error remain frequent).
+	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size.
+Even in very large samples both types of error remain frequent, and the risk of saving no work at all remains close to 50\%.
 	This shows how baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/bir_errors}
+		\includegraphics[width=\textwidth]{2_figs_bir_errors.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/bir_error_distribution}
+		\includegraphics[width=\textwidth]{2_figs_bir_error_distribution.pdf}
 		\caption[]%
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{../images/proportions_1.png}
-	\includegraphics[width=\linewidth]{../images/proportions_2.png}
+	\includegraphics[width=\linewidth]{2_figs_proportions_1.png}
+	\includegraphics[width=\linewidth]{2_figs_proportions_2.png}
 	\caption{Similar low proportions of relevant documents in unseen documents with different consequences for recall.
@@
 Some studies give the example of heuristic stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
-We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low. 
-We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; and 2) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.
+We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low, as the probability of observing 0 relevant documents in a given sample (analogous to a set of consecutive irrelevant results) is inversely related to the number of relevant documents in the population.
+We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; 
+2) it lacks a formal definition; and 3) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.
 
-Figure \ref{unseen-proportions} illustrates this second point. 
+Figure \ref{unseen-proportions} illustrates this third point. 
 We show two scenarios with identical low proportions of relevant documents observed in the unseen documents.
@@
 Likewise, where the proportion of relevant documents in the whole corpus is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low. 
+This shows us that even a perfect estimator of the relevance of unseen documents is insufficient on its own to provide sufficient information about when to stop screening.
 
-\subsubsection*{Other stopping criteria}
+\subsubsection*{Pragmatic stopping criteria}
 
@@
 Jonnalagadda and Petitti \cite{Jonnalagadda2013} note that ``the reviewer can elect to end the process of classifying documents at any point, recognizing that stopping before reviewing all documents involves a trade-off of lower recall for reduced workload'', although clearly the reviewer lacks information about probable recall.
-Yu and Menzies \cite{Yu2019} adopt a more complicated stopping criterion which allows the user to target a specific level of recall.
+
+\subsubsection*{Novel automatic stopping criteria}
+Two examples come from the information retrieval literature.
+Di Nunzio \cite{DiNunzio2018} presents a novel automatic stopping criterion based on BM25, although recall reported is ``often between 0.92 and 0.94 and consistently over 0.7''.
+Yu and Menzies \cite{Yu2019} also present a stopping criterion based on BM25 which allows the user to target a specific level of recall. 
 However, reviewers are not given the opportunity to specify a confidence level, and for two of the four datasets in which they tested their criteria, the median achieved recall at a stopping criteria targeting 95\% recall was below 95\%.
-Di Nunzio \cite{DiNunzio2018} also present an innovative stopping criteria, but it does not take into account uncertainty, and produces results \textit{near} a target recall threshold, rather than above it in a reliable proportion of cases.
-These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount.
+In each case, the reliability of the estimate is dependent on the performance of the model.
+
+Finally, Howard et al.
+\cite{Howard2020} present a method to estimate recall based on the the number of irrelevant documents $D$ observed in a list of documents since the $\delta th$ previous relevant document.
+They reason that this should follow the negative binomial distribution based on the proportion of remaining relevant documents $p$, and use this information to estimate $\hat{p}$, and with this, the total number of relevant articles and the estimated recall.
+However, since screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial.
+Further the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$.
+Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally.
+Lastly, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours).
+
+These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance.
+In the following, we explain our own method, which provides clearly communicable estimates of recall, which manage uncertainty in a way robust to model performance.
 
@@
 
+	In our screening setup, we start off with $N_T$ documents that are potentially relevant.
+$\rho_{tot}$ of these documents are actually relevant, but we don't know this value \textit{a priori}.
+As we screen relevant documents we include them, so $\rho_{seen}$ represents the number of relevant documents screened, and $\tau$ recall is given by
+
+	\begin{equation}
+		\tau = \frac{\rho_{seen}}{\rho_{tot}}
+	\end{equation}
+
+	We set a target recall $\tau_{tar}$ and a confidence level $\alpha$. 
+	We want to keep screening until $\tau \geq \tau_{tar}$, and devise a hypothesis test to estimate whether this is the case with a given level of confidence. 
+	We do this	based on interrupting the active-learning process and drawing a random sample from the remaining unseen documents. 
+	We first describe this test, before showing how a variation on the test can be used to decide when to begin drawing a random sample.
+
 	\subsubsection*{Random Sampling}
 
-	After using machine learning to select which documents are screened by humans as described above, we begin drawing a random sample from the remaining documents (when this happens is described below).
-
-	We use random sampling to estimate the probability that a target recall $\tau$ has been achieved.
-Because we are sampling a binary outcome without replacement, we can use the hypergeometric distribution to formulate a statistical test.
-The hypergeometric distribution tells us the probability of observing $k$ relevant documents in $n$ draws from a finite population of $N$ documents with $K$ relevant documents.
+	At the start of the sample, $N_{AL}$ is the number of documents seen during the active learning process, and $N_s$ is the number of documents remaining, so that
 
 	\begin{equation}
-		k \sim Hypergeometric(N, K, n)
+	N_s = N_T - N_{AL}
 	\end{equation}
 
-	In our case, we know $k$, $n$ and $N$ after each draw, but $K$ is unknown.
-We therefore substitute a hypothetical value $\hat{K}$: the minimum number of relevant documents in the sample had the recall target been missed.
-
-	Recall $R$ is given by the number of relevant documents that have been seen $\rho_{s}$ over the number of relevant documents in the whole dataset $\rho_{tot}$
+	We refer to the number  of relevant documents seen during active learning as $\rho_{AL}$, and the number of remaining relevant documents as $K$.
+We do not know the value of $K$ but know that it is given by the total number of relevant documents minus the number of relevant documents seen during active learning.
 
 	\begin{equation}
-		R = \frac{\rho_{seen}}{\rho_{tot}}
+	K = \rho_{tot} - \rho_{AL}
 	\end{equation}
 
-	The number of relevant documents in the whole dataset is the sum of $\hat{\rho}$ relevant documents seen before random sampling began and $\tilde{K}$ relevant documents unseen at the start of random sampling.
-We can therefore express $R$ as
+	With each draw, $n$ is the number of documents drawn and $k$ is the number of relevant documents drawn.
+The number of relevant documents seen is updated by adding the number of relevant documents seen since sampling began to the number of relevant documents seen during active learning.
 
 	\begin{equation}
-		R = \frac{\dot{\rho}}{\hat{\rho} + \tilde{K}},
+	\rho_{seen} = \rho_{ML} + k
 	\end{equation}
 
-	Substituting the target recall $\tau$ for $R$, reorganising and rounding up to the next integer, we can, after each draw, calculate $\hat{K}$, which is the minimum number of relevant documents that could have been remaining when random sampling started, if recall were lower than the target.
+	We proceed to form a null hypothesis that the true value of recall is less than our target recall:
 
 	\begin{equation}
-		\hat{K} = \lceil \frac{\rho}{\tau} - \hat{\rho} \rceil
+	H_0 : \tau < \tau_{tar}
 	\end{equation}
 
-	We use the cumulative distribution function of the hypergeometric distribution to estimate the probability $p$ of having observed $k$ or fewer relevant documents in the sample given $\hat{K}$.
-This function gives us an upper bound on the probability of observing no more than the number of relevant documents in our random sample that we did, if our recall target had not been achieved.
-	If this is below our confidence level $1 - \alpha$, we can reject the null hypothesis that the recall target was not achieved.
+	Because we are sampling without replacement, we can use the hypergeometric distribution to find out the probability of observing $k$ relevant documents in a sample of $n$ documents from a population of $N$ documents of which $K$ are relevant.
+We know that $k$ is distributed hypergeometrically:
 
-	\subsubsection*{Pseudo-random sampling}
+	\begin{equation}
+	X \sim Hypergeometric(N, K, n)
+	\end{equation}
 
-	In order to decide when to begin a random sample, we employ pseudo-random sampling, where we treat previously screened documents as a random sample.
+	We introduce a hypothetical value for $K$, which we call $K_0$.
+This represents the minimum number of relevant documents remaining at the start of sampling compatible with our null hypothesis that recall is below our target.
+
+	\begin{equation}
+	K_0 = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
+	\end{equation}
+
+	Our null hypothesis can thereby reformulated: the true number of relevant documents in the sample is greater than our hypothetical value.
+
+	\begin{equation}
+	H_0 : K > K_{tar}
+	\end{equation}
+
+	We test this by calculating the probability of observing $k$ or fewer relevant from the hypergeometric distribution given by $K_{tar}$, using the cumulative distribution function.
+
+	\begin{equation}
+	p = P(X \leq k)$, where $X \sim Hypergeometric(N_s,K_{tar},n)
+	\end{equation}
+
+	Because $P(X \leq k)$, is strictly decreasing for values below $K_{tar}$, this gives the maximum probability of observing $k$ for all values of K compatible with our null hypothesis.
+
+	If the maximum probability of obtaining our observed results given our null hypothesis is below our $1-\alpha$, then we can reject our null hypothesis and stop screening.
+We can report the likelihood that we achieve a recall below our target as being less than $1-\alpha$.
+
+	\subsubsection*{Nonrandom sampling}
+
+	In order to decide when to begin a random sample, we employ nonrandom sampling, where we treat previously screened documents as random samples.
 The distribution of relevant documents among previously screened documents is clearly not random, as documents predicted to be relevant are prioritised.
@@
 
-	After reviewing each document, $S$ documents have been screened, and $U$ documents are yet to be seen.
-We treat $i = 1 \dots S$ of the previously screened documents as a random sample, and calculate $p$, using the method above, for each sample, taking the minimum across all samples $p_{min}$.
+	After reviewing each document, $N_{seen}$ documents have been screened, and $N_{unseen}$ documents are yet to be seen. 
+	We treat $i = 1 \dots N_{seen}$ of the previously screened documents as separate random samples, and calculate $p$, using the method above, for each sample. 
+	We take the minimum across all samples $p_{min}$ as the lowest probability that our null hypothesis is correct. 
 	If $p_{min}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling. 
-We also calculate $p_{min}$ for the remaining documents as if we had not switched to random sampling and record the recall and work saved when $p_{min} < 1 - \alpha$.
-We present these in the results below as the psuedo-random sampling criterion.
+	This method is an imprecise heuristic used to decide when to start random sampling. 
+	However, we also test its performance as a separate stopping criterion.
+	We calculate $p_{min}$ for the remaining documents as if we had not switched to random sampling and record the recall and work saved when $p_{min} < 1 - \alpha$. 
+	We present these in the results below as the nonrandom sampling criterion.
 
@@
 	\begin{figure}
-		\includegraphics[width=0.7\linewidth]{../images/flow}
+		\includegraphics[width=0.7\linewidth]{2_figs_flow.pdf}
 		\caption{A workflow for active learning in screening with a statistical stopping criterion}
@@
 Subsequently, we ``screen'', that is, we reveal the labels of, batches of the 20 documents with the highest predicted relevance scores, retraining the model after each batch.
+Theoretically, using smaller batch sizes could mean that the recall target is achieved more quickly, but this is a trade-off between computational time spent training, and the speed at which the algorithm can ``learn''.
+However this is a modelling choice which may affect work saved, but not recall.
 Each criterion is evaluated after each document is ``screened''.
@@
 	\begin{table}
-		\input{../tables/datasets.tex}
+		\input{1_tables_datasets.tex}
 		\caption{Dataset properties}
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_hyper}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_hyper.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_nrs}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_nrs.pdf}
 		\caption[]%
@@
 		\centering
-		\includegraphics[width=\textwidth]{../images/jointplot_bir}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_bir.pdf}
 		\caption[Network2]%Example-Image
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_ih_50}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_ih_50.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_ih_200}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_ih_200.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_pf.pdf}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_pf.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_hyper}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_hyper.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_nrs.pdf}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_nrs.pdf}
 		\caption[]%
@@
 		\centering
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_bir}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_bir.pdf}
 		\caption[Network2]%Example-Image
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_ih_50}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_ih_50.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_ih_100}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_ih_100.pdf}
 		\caption[]%
@@
 		\centering 
-		\includegraphics[width=\textwidth]{../images/jointplot_burden_ih_200.pdf}
+		\includegraphics[width=\textwidth]{2_figs_jointplot_burden_ih_200.pdf}
 		\caption[]%
@@
 	\begin{figure}
-		\includegraphics[width=0.9\linewidth]{../images/wss_nrs}
+		\includegraphics[width=0.9\linewidth]{2_figs_wss_nrs.pdf}
 		\caption{Work saved for the pseudorandom sampling method in each dataset.
@@
 			\centering
-			\includegraphics[width=\textwidth]{../images/h0_paths_Radjenovic.pdf}
+			\includegraphics[width=\textwidth]{2_figs_h0_paths_Radjenovic.pdf}
 			\caption[Network2]%Example-Image
@@
 			\centering 
-			\includegraphics[width=\textwidth]{../images/h0_paths_ProtonBeam.pdf}
+			\includegraphics[width=\textwidth]{2_figs_h0_paths_ProtonBeam.pdf}
 			\caption[]%
@@
 			\centering
-			\includegraphics[width=\textwidth]{../images/h0_paths_Statins.pdf}
+			\includegraphics[width=\textwidth]{2_figs_h0_paths_Statins.pdf}
 			\caption[Network2]%Example-Image
@@
 			\centering 
-			\includegraphics[width=\textwidth]{../images/h0_paths_Triptans.pdf}
+			\includegraphics[width=\textwidth]{2_figs_h0_paths_Triptans.pdf}
 			\caption[]%
@@
 
-	This paper demonstrates the unsuitability of existing stopping criteria for machine learning approaches to document screening, and proposes a simple method that delivers reliable recall, independent of machine learning approach or model performance.
+	This paper demonstrates the drawbacks of existing stopping criteria for machine learning approaches to document screening, particularly with regard to reliability.
+We propose a simple method that delivers reliable recall, independent of machine learning approach or model performance.
 Our robust statistical stopping criteria allow users to easily communicate the implications of their use of machine learning, making machine learning enhanced screening ready for live reviews.
@@
 \begin{backmatter}
+
+	\section*{Ethics approval}
+	Not applicable.
+
+	\section*{Consent for publication}
+	Not applicable.
+
+	\section*{Availability of data and materials} 
+	All computational steps required to reproduce this analysis are documented online at \url{https://github.com/mcallaghan/rapid-screening}.
 
@@
 
-	\bibliography{mendeley}
+	\bibliography{3_bib_mendeley}
 	\bibliographystyle{vancouver}

### 2020-06-11 analysis/run_scenarios.py

@@ -91,7 +91,7 @@ for name, group in cohen_db.groupby('review'):
                 if r is not None: 
                    results.append(r)
 
-if args.mpi:
+if args.mpi:a
     results_df = pd.DataFrame.from_dict(results)
     results_df.to_csv(f'../results/results_{rank}.csv', index=False)
 else:

### 2020-06-13 manuscript/main.tex
resubmission

@@
 %\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass[linenumbers]{bmcart}
+\documentclass{bmcart}
 
@@
 \usepackage{todonotes}
+\usepackage{bm}
 \usepackage{graphicx}
@@
 \usepackage{subcaption}
+\usepackage{caption}
 \usepackage{booktabs}
@@
 	Several papers have developed Active Learning (AL) approaches \cite{miwa2014, Wallace2010a, Wallace2010, Jonnalagadda2013, Przybya2018} to reduce the time required to screen documents.
-This paper sets out how current approaches are  unsuitable in practice, and outlines and evaluates a small modification that would make AL systems ready for live reviews.
+This paper sets out how current approaches are  unreliable in practice, and outlines and evaluates modifications that would make AL systems ready for live reviews.
 
@@
 	The algorithm chooses which studies will next be screened by humans, often those which are likely to be relevant or about which the model is uncertain, in order to generate more labels to feed back to the machine.
-	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies - or a given proportion of relevant studies (described by recall: the number of relevant studies identified divided by the total number of relevant studies) - before having seen all the documents in the corpus.
+	By prioritising those studies most likely to be relevant, a human reviewer most often identifies all relevant studies -- or a given proportion of relevant studies (described by recall: the number of relevant studies identified divided by the total number of relevant studies) -- before having seen all the documents in the corpus.
 	The proportion of documents not yet seen by the human when they reach the given recall threshold is referred to as the work saved.
@@
 If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
-	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of 95\%''.
+	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of more than 95\%''.
 
 	In the remainder of the paper, we first discuss in detail the shortcomings of existing stopping criteria.
-Then, we introduce our new criterion based on a hypergeometric test.
+Then, we introduce our new criteria based on a hypergeometric test.
 We evaluate our stopping criteria, and compare their performance with heuristic and sampling based criteria on real-world systematic review datasets on which AL systems have previously been tested \cite{Cohen2006, Yu2019, Terasawa2009, Castaldi2009}.
@@
 To reduce the margin of error to $\pm 5\%$ of estimated relevant studies, they would have had to screen 638,323 out of 804,919 results.
-See the notebook \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb} that accompanies this paper for a detailed discussion}.
+See the notebook \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/bir_theory.ipynb} that accompanies this paper for a detailed discussion.}.
 	This uncertainty is crucial, as errors can have severe consequences.
-If the estimated number of relevant documents is even one unit above the true value, then no work savings will be achieved.
-If the number of relevant documents is underestimated, then the recall achieved will be less than 100\%.
+Let us assume that users will stop screening when they have identified 95\% of the relevant number of documents.
+If the estimated number of relevant documents is more than the true number of relevant documents divided by 0.95, then the users will never see 95\% of the estimated number.
+This means that they will keep screening until they have seen all documents, and no work savings will be achieved.
+Conversely, if the number of relevant documents is underestimated by even a single unit, then the recall achieved will be lower than the target.
 
@@
 The left shaded portion of the graph shows all the cases where the recall will be less than 95\%.
-This occurs 35.91\% of the time.
-The right shaded portion of the graph shows the cases where the number of relevant documents is overestimated and no work savings could be made.
-This occurs 46.24\% of the time.
-In only 17.85\% of cases can work savings be achieved while still achieving a recall of at least 95\%.
+This occurs 48\% of the time.
+The right shaded portion of the graph shows the cases where the number of relevant documents is overestimated so much that no work savings could be made to achieve a target recall of 95\%.
+This occurs 29\% of the time.
+In only 23\% of cases can work savings be achieved while still achieving a recall of at least 95\%.
 
-	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size.
-Even in very large samples both types of error remain frequent, and the risk of saving no work at all remains close to 50\%.
-	This shows how baseline estimation inevitably offers poor reliability, both in terms of recall and in work saved.
+	Figure \ref{fig:bir_error_distribution} shows the probability distribution of these errors according to the sample size. Even in very large samples both types of error remain frequent.
+	This shows how baseline estimation inevitably offers poor reliability, either in terms of recall or in work saved.
 
@@
 	Some studies give the example of heuristic stopping criteria based on drawing a given number of irrelevant articles in a row \cite{Jonnalagadda2013, Przybya2018}. 
-We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low, as the probability of observing 0 relevant documents in a given sample (analogous to a set of consecutive irrelevant results) is inversely related to the number of relevant documents in the population.
+	We take this as a proxy for estimating that the proportion of documents remaining in the unseen documents is low, as the probability of observing 0 relevant documents in a given sample (analogous to a set of consecutive irrelevant results) is a decreasing function of the number of relevant documents in the population.
 	We find this a promising intuition, but argue that 1) it ignores uncertainty, as discussed in relation to the previous method; 
-2) it lacks a formal definition; and 3) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.
+	2) it lacks a formal description that would help to find a suitable threshold for the criterion; and 3) it misunderstands the significance of a low proportion of relevant documents in estimating the recall.
 
@@
 	In both cases, only 2\% of unseen documents are relevant, but 2\% of a larger number means more relevant documents are missed.
-Recall is not simply a function of the relevance of unseen documents, but also of the number of unseen documents. 
+	Recall is not simply a function of the proportion of unseen relevant documents, but also of the number of unseen documents. 
 	This also means that where ML has performed well (as in the top figure), a low proportion of relevant documents in those not yet checked is indicative of lower recall than where ML has performed less well.
 	Likewise, where the proportion of relevant documents in the whole corpus is low, a similarly low proportion of relevant documents is likely to be observed, even when true recall is low. 
-This shows us that even a perfect estimator of the relevance of unseen documents is insufficient on its own to provide sufficient information about when to stop screening.
+	This shows us that even a perfect estimator of the proportion of unseen relevant documents is insufficient on its own to provide sufficient information about when to stop screening.
+To estimate recall reliably, it is necessary to take into account the total number of unseen relevant documents (or their proportion times the number of unseen documents).
 
@@
 	Finally, Howard et al.
-\cite{Howard2020} present a method to estimate recall based on the the number of irrelevant documents $D$ observed in a list of documents since the $\delta th$ previous relevant document.
+\cite{Howard2020} present a method to estimate recall based on the number of irrelevant documents $D$ observed in a list of documents since the $\delta th$ previous relevant document.
 	They reason that this should follow the negative binomial distribution based on the proportion of remaining relevant documents $p$, and use this information to estimate $\hat{p}$, and with this, the total number of relevant articles and the estimated recall.
-However, since screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial.
+
+	However, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours).
+This is not guaranteed by the criterion itself but rather a finding of the simulation with example datasets.
 Further the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$.
 Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally.
-Lastly, their method does not quantify uncertainty, but can only claim that the method ``\textit{tends} to result in a conservative estimate of recall'' (emphasis ours).
+	Lastly, because screening is a form of sampling without replacement, the negative hypergeometric distribution should be preferred to the negative binomial, even though the latter can be a good approximation for cases with large numbers of documents.
 
-These last examples are promising developments, but fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance.
+	These last examples are promising developments, but they all fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance.
 In the following, we explain our own method, which provides clearly communicable estimates of recall, which manage uncertainty in a way robust to model performance.
@@
 
-	In our screening setup, we start off with $N_T$ documents that are potentially relevant.
+	In our screening setup, we start off with $N_{tot}$ documents that are potentially relevant.
 $\rho_{tot}$ of these documents are actually relevant, but we don't know this value \textit{a priori}.
-As we screen relevant documents we include them, so $\rho_{seen}$ represents the number of relevant documents screened, and $\tau$ recall is given by
+As we screen relevant documents we include them, so $\rho_{seen}$ represents the number of relevant documents screened, and recall $\tau$ is given by
 
@@
 	\tau = \frac{\rho_{seen}}{\rho_{tot}}
+	\label{eq:recall-def}
 	\end{equation}
@@
 
-	At the start of the sample, $N_{AL}$ is the number of documents seen during the active learning process, and $N_s$ is the number of documents remaining, so that
+	At the start of the sample, $N_{AL}$ is the number of documents seen during the active learning process, and $N$ is the number of documents remaining, so that
 
 	\begin{equation}
-	N_s = N_T - N_{AL}
+	N = N_{tot} - N_{AL}
 	\end{equation}
@@
 	K = \rho_{tot} - \rho_{AL}
+	\label{eq:K}
 	\end{equation}
 
-	With each draw, $n$ is the number of documents drawn and $k$ is the number of relevant documents drawn.
+	We now take random draws from the remaining $N$ documents, and denote the number of documents drawn with $n$ and the number of relevant documents drawn with $k$.
 The number of relevant documents seen is updated by adding the number of relevant documents seen since sampling began to the number of relevant documents seen during active learning.
@@
 	\begin{equation}
-	\rho_{seen} = \rho_{ML} + k
+	\rho_{seen} = \rho_{AL} + k
+	\label{eq:rho_seen}
 	\end{equation}
@@
 	H_0 : \tau < \tau_{tar}
+	\label{eq:null_hypothesis}
 	\end{equation}
@@
 	\begin{equation}
-	X \sim Hypergeometric(N, K, n)
+	k \sim Hypergeometric(N, K, n)
 	\end{equation}
 
-	We introduce a hypothetical value for $K$, which we call $K_0$.
+	We introduce a hypothetical value for $K$, which we call $K_{tar}$.
 This represents the minimum number of relevant documents remaining at the start of sampling compatible with our null hypothesis that recall is below our target.
@@
 	\begin{equation}
-	K_0 = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
+	K_{tar} = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
 	\end{equation}
 
-	Our null hypothesis can thereby reformulated: the true number of relevant documents in the sample is greater than our hypothetical value.
+	This equation is derived by combining Eqs.~\ref{eq:recall-def} and \ref{eq:rho_seen}.
+Because $k$ can only take integer values, $K_{tar}$ is the smallest integer that satisfies the inequality in Eq.~\ref{eq:null_hypothesis}.
+	With $K_{tar}$, we can reformulate our null hypothesis: the true number of relevant documents in the sample is greater than our hypothetical value.
 
@@
 
-	We test this by calculating the probability of observing $k$ or fewer relevant from the hypergeometric distribution given by $K_{tar}$, using the cumulative distribution function.
+	We test this by calculating the probability of observing $k$ or fewer relevant documents from the hypergeometric distribution given by $K_{tar}$, using the cumulative probability mass function.
 
 	\begin{equation}
-	p = P(X \leq k)$, where $X \sim Hypergeometric(N_s,K_{tar},n)
+	p = P(X \leq k)$, where $X \sim Hypergeometric(N,K_{tar},n)
+	\label{eq:p-value}
 	\end{equation}
 
-	Because $P(X \leq k)$, is strictly decreasing for values below $K_{tar}$, this gives the maximum probability of observing $k$ for all values of K compatible with our null hypothesis.
+	Because the cumulative probability mass function $P(X \leq k)$ is decreasing with increasing $K$, this gives the maximum probability of observing $k$ for all values of $K$ compatible with our null hypothesis.
+Similar arguments have been made to derive confidence intervals for estimating the parameter $K$ in the hypergeometric distribution function \citep{Buonaccorsi1987, Sahai1995} and the derivation of an equivalent criterion could use the upper limit of such a confidence interval of an estimated $K$ from the observation of $k$.
 
-	If the maximum probability of obtaining our observed results given our null hypothesis is below our $1-\alpha$, then we can reject our null hypothesis and stop screening.
-We can report the likelihood that we achieve a recall below our target as being less than $1-\alpha$.
+	We can reject our null hypothesis and stop screening if the maximum probability of obtaining our observed results given our null hypothesis $p$ is below $1-\alpha$.
+This means, we can report the likelihood that we achieve a recall above our target as being more than $\alpha$ \footnote{The notebook, \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteria_theory.ipynb}, in the github repository accompanying this paper contains a step by step explanation of this method with code and examples}.
 
-	\subsubsection*{Nonrandom sampling}
+	\subsubsection*{Ranked quasi-sampling}
 
-	In order to decide when to begin a random sample, we employ nonrandom sampling, where we treat previously screened documents as random samples.
-The distribution of relevant documents among previously screened documents is clearly not random, as documents predicted to be relevant are prioritised.
-It is reasonable to assume, though, that the density of relevant documents is greater among previously screened documents than among remaining unseen documents.
-This would make the following estimates conservative.
+	We now proceed to describe a special case of the method described above which we (1) use as a heuristic in order to decide when to begin random sampling; and (2) test as an independent stopping criterion.
+The method works by treating batches of previously screened documents as if they were random samples.
 
-	After reviewing each document, $N_{seen}$ documents have been screened, and $N_{unseen}$ documents are yet to be seen. 
-	We treat $i = 1 \dots N_{seen}$ of the previously screened documents as separate random samples, and calculate $p$, using the method above, for each sample. 
-	We take the minimum across all samples $p_{min}$ as the lowest probability that our null hypothesis is correct. 
-	If $p_{min}$ is less than $1-\frac{\alpha}{2}$, we switch to random sampling. 
-	This method is an imprecise heuristic used to decide when to start random sampling. 
-	However, we also test its performance as a separate stopping criterion.
-	We calculate $p_{min}$ for the remaining documents as if we had not switched to random sampling and record the recall and work saved when $p_{min} < 1 - \alpha$. 
-	We present these in the results below as the nonrandom sampling criterion.
+	We calculate $p$ as above for subsets of the already screened documents.
+Concretely, we use subsets of documents $A_i$ by looking back to the last $i$ documents, $A_i = \{d_{N_{seen} - 1}, ..., d_{N_{seen} - i}\}$, where the documents $d$ are indexed in the order in which they have been screened.
+For a specific $i$, this corresponds to  random sampling beginning after seeing $i$ documents in the section above.
+	Thus, we set 
+	$N_{AL}$ to $i$, 
+	$n$ to $N_{seen}-i$, 
+	$\rho_{AL}$ to the number of relevant documents seen when $i$ documents had been seen,
+	and $k$ to the number of relevant documents seen since $i$ documents had been seen, and calculate $p$ according to Eq.~\ref{eq:p-value}.
+	We compute $p$ for all sets $A_i$ with $i \in {N_{seen}-1 \dots 1}$.
+	This gives us a vector $\bm{p}$, representing the values of $p$ which would have been estimated at each point at which we could have stopped active learning and began random sampling.
+The lowest probability of our null hypothesis being true that we would have thereby obtained is given by $p_{min}$.
+With the vectorized implementation included in our accompanying code, these calculations are completed in less than the time it would take a human to code the next document.
+
+	First, we use this method as a useful heuristic for deciding when to stop active learning, and switch to random sampling.
+For this, we choose a higher threshold for the likelihood, $p_{min} < 1-\frac{\alpha}{2}$.
+Second, we use the same ranked quasi-sampling as an independent stopping criterion, by continuing screening with active learning until $p_{min} < 1 - \alpha$.
+We present the results of this second procedure separately below.
+
+	Given that the documents seen during active learning are ranked according to predicted relevance, they do not in fact represent a random sample.
+This means that the test is unlikely to be accurate.
+It would be reasonable to assume that the proportion of relevant documents in each ranked quasi-sample is as high if not higher than the proportion of relevant documents in the unseen documents.
+This assumption would make this estimator conservative.
+As such it works in a similar way to the criterion proposed by Howard et al.
+\cite{Howard2020}, although it makes use of more information and provides hypothesis testing rather than just a point estimate of recall.
 
@@
 			\caption[]%
-		{{\footnotesize Pseudorandom hypergeometric sampling}}    
+			{{\footnotesize Ranked quasi-sampling}}    
 			\label{fig:nrs}
@@
 			\caption[]%
-		{{\footnotesize Pseudorandom hypergeometric sampling}}    
+			{{\footnotesize Ranked quasi-sampling}}    
 			\label{fig:nrs_ab}
@@
 
-	Both the random sampling and the pseudorandom sampling criteria achieve the target threshold of 95\% in more than 95\% of cases.
-In fact, the pseudorandom sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases.
-In theory, the pseudorandom sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random.
+	Both the random sampling and the ranked sampling criteria achieve the target threshold of 95\% in more than 95\% of cases. 
+	That this is greater than 95\% is accounted for by the fact that random sampling sometimes begins after the target has been achieved, in which case the null hypothesis would be \text{a priori} impossible.
+	The ranked quasi-sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases.
+In theory, the ranked sampling criteria is conservative if the assumption holds that documents chosen by machine learning are not less likely to be relevant than those chosen at random.
 Based on our experiments, this assumption seems reasonable, and accounts for the higher recall.
-Because the pseudorandom sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data.
+Because the ranked quasi-sampling criterion can flexibly choose its sample, whereas the random criterion has to wait for a random sample to be triggered, the criterion is also triggered earlier, as it can make use of more data.
 This accounts for the higher work savings.
@@
 		\includegraphics[width=0.9\linewidth]{2_figs_wss_nrs.pdf}
-		\caption{Work saved for the pseudorandom sampling method in each dataset.
+		\caption{Work saved for the ranked quasi-sampling method in each dataset.
 Labels show the number of relevant documents and the total number of documents.
@@
 
-	To help explain the different work savings that were observed in our experiments, we show the distribution of work savings from our pseudorandom criterion for each dataset in figure \ref{wss}.
+	To help explain the different work savings that were observed in our experiments, we show the distribution of work savings from our ranked quasi-sampling criterion for each dataset in figure \ref{wss}.
 In general, higher work savings are possible when the total number of documents is larger.
@@
 	When using machine learning for screening, poor recall can result in biased results, while low work savings represent no loss to the reviewer as compared to not using machine learning.
+
+	One caveat in the derivation of our criteria is that we did not address the potential problem of multiple testing formally.
+Such a derivation is mathematically challenging and beyond the scope of this paper.
+However, the performance of the criteria shows that this is of limited practical concern.
+Formally describing screening procedures with iterative testing should be a next step towards even more rigorous stopping criteria and should be fully worked out in future research.
 

### 2020-09-01 manuscript/main.tex
re-resubmission

@@
 
-			\title{Robust Statistical Stopping Criteria for Automated Screening in Systematic Reviews}
+			\title{Statistical Stopping Criteria for Automated Screening in Systematic Reviews}
 
@@
 This means that such systems are hard to implement in live reviews.
-				This paper introduces a workflow with robust and flexible statistical stopping criteria, which offer real work reductions on the basis of a given confidence level of reaching a given recall.
+				This paper introduces a workflow with flexible statistical stopping criteria, which offer real work reductions on the basis of rejecting a hypothesis of having missed a given recall target with a given level of confidence.
 				The stopping criteria are shown on test datasets to achieve a reliable level of recall, while still providing work reductions of on average 17\%.
@@
 If the hypothesis can be rejected, we conclude that the recall target has been achieved with a given confidence level and screening can be stopped.
-	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``A recall of more than 95\% was achieved with a confidence of more than 95\%''.
+	This allows AL users to predefine a target in terms of uncertainty and recall, so that they can make transparent, easily communicable statements like ``We reject the null hypothesis that we achieve a recall of less than 95\% with a significance level of 5\%''.
 
@@
 This is not guaranteed by the criterion itself but rather a finding of the simulation with example datasets.
-Further the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$.
+Further, the authors do not give sufficient information to reproduce their results, providing neither code (they describe their own proprietary software), nor an equation for $\hat{p}$.
 Additionally, the criterion requires a tuning parameter $\delta$, which users may have insufficient information to set optimally.
@@
 	These last examples are promising developments, but they all fail to take into account the needs of live systematic reviews, where the reliability of and ease of communication about recall are paramount, and the results are independent of model performance.
-In the following, we explain our own method, which provides clearly communicable estimates of recall, which manage uncertainty in a way robust to model performance.
+In the following, we explain our own method, which provides clearly communicable estimates of recall, and which manage uncertainty in a way robust to model performance.
 
@@
 
+	Accordingly, the alternative hypothesis is that recall is equal to or greater than our target:
+
+	\begin{equation}
+	H_1 : \tau \geq \tau_{tar}
+	\end{equation}
+
 	Because we are sampling without replacement, we can use the hypergeometric distribution to find out the probability of observing $k$ relevant documents in a sample of $n$ documents from a population of $N$ documents of which $K$ are relevant.
@@
 Because $k$ can only take integer values, $K_{tar}$ is the smallest integer that satisfies the inequality in Eq.~\ref{eq:null_hypothesis}.
-	With $K_{tar}$, we can reformulate our null hypothesis: the true number of relevant documents in the sample is greater than our hypothetical value.
+	With $K_{tar}$, we can reformulate our null hypothesis: the true number of relevant documents in the sample is greater than or equal to our hypothetical value.
 
 	\begin{equation}
-	H_0 : K > K_{tar}
+	H_0 : K \geq K_{tar}
 	\end{equation}
@@
 
-	We can reject our null hypothesis and stop screening if the maximum probability of obtaining our observed results given our null hypothesis $p$ is below $1-\alpha$.
-This means, we can report the likelihood that we achieve a recall above our target as being more than $\alpha$ \footnote{The notebook, \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteria_theory.ipynb}, in the github repository accompanying this paper contains a step by step explanation of this method with code and examples}.
+	We can reject our null hypothesis and stop screening if the maximum probability of obtaining our observed results given our null hypothesis $p$ is below $1-\alpha$ \footnote{The notebook, \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteria_theory.ipynb}, in the github repository accompanying this paper contains a step by step explanation of this method with code and examples}.
+To further investigate the accuracy of the test, we perform an experiment drawing 1 million random samples in 6 scenarios with different characteristics.
+We vary the value of $\rho_{AL}$ to simulate starting random sampling with different levels of recall achieved.
+
+	\begin{figure}
+		\centering
+		\includegraphics[width=\linewidth]{2_figs_hyper_validation.png}
+		\caption{The distribution of achieved recall values given our random sampling stopping criterion for 6 scenarios with different recall values at the start of sampling.}
+		\label{fig:hyper-valid}
+	\end{figure}
+
+	Figure \ref{fig:hyper-valid} shows that in each case, as long as recall is lower than the target recall when sampling begins, the percentage of times that the criteria is triggered too early is within two tenths of a percentage point of 5\% and the the 5th percentile of achieved recall values is within two tenths of a percentage point of the target recall 95\%.
 
@@
 	This gives us a vector $\bm{p}$, representing the values of $p$ which would have been estimated at each point at which we could have stopped active learning and began random sampling.
-The lowest probability of our null hypothesis being true that we would have thereby obtained is given by $p_{min}$.
+The point at which the p-value for our null hypothesis is lowest is given by $p_{min}$.
 With the vectorized implementation included in our accompanying code, these calculations are completed in less than the time it would take a human to code the next document.
@@
 	Both the random sampling and the ranked sampling criteria achieve the target threshold of 95\% in more than 95\% of cases. 
-	That this is greater than 95\% is accounted for by the fact that random sampling sometimes begins after the target has been achieved, in which case the null hypothesis would be \text{a priori} impossible.
+	That this is greater than 95\% is accounted for by the fact that random sampling sometimes begins after the target recall has been achieved, in which case the null hypothesis would be \text{a priori} impossible.
 	The ranked quasi-sampling criterion  outperforms the random sampling criterion with respect to both recall and work savings, saving a mean of 17\% of the work compared to 15\%, and missing the target in only 0.95\% compared to 3.29\% of cases.
@@
 
-	The baseline sampling criteria (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
+	The baseline sampling criteron (Figure \ref{fig:bir}) misses the 95\% recall target in 39.67\% of cases, while the most common work saving is 0\%.
 This is in line with our expectations that, due to random sampling error, the expected number of documents will often be over-estimated or under-estimated, resulting in zero work savings or poor recall.
@@
 
-	Figure \ref{H0paths} shows the recall and the probability of the null hypothesis for the best performing iteration of four datasets.
+	Figure \ref{H0paths} shows the recall and the p-value for the null hypothesis for the the iteration where the recall target is reached first for four datasets.
 Although the 95\% recall target is achieved very quickly in the Radjenovic dataset, the null hypothesis cannot be excluded until much later.
@@
 These can be flexibly accommodated within our stopping criterion.
-Importantly, the ability to make probabilistic statements about the chance of achieving a given recall target makes it possible to clearly communicate the implications of using machine learning enhanced screening to readers and reviewers who are not machine learning specialists.
+Importantly, the ability to make statements about the authors' confidence in achieving a given recall target makes it possible to clearly communicate the implications of using machine learning enhanced screening to readers and reviewers who are not machine learning specialists.
 This is extremely important in live systematic reviews.
@@
 
-	One caveat in the derivation of our criteria is that we did not address the potential problem of multiple testing formally.
+	One caveat in the derivation of our criteria is that we did not address the problem of multiple testing formally.
 Such a derivation is mathematically challenging and beyond the scope of this paper.
@@
 We propose a simple method that delivers reliable recall, independent of machine learning approach or model performance.
-Our robust statistical stopping criteria allow users to easily communicate the implications of their use of machine learning, making machine learning enhanced screening ready for live reviews.
+Our statistical stopping criteria allow users to easily communicate the implications of their use of machine learning, making machine learning enhanced screening ready for live reviews.
 

### 2020-09-08 images/flow_noeq_basic.tex
new presentation

--- 2019-07-03 images/flow.tex
@@
   \node (screenrel)  [screen, below of =sample]  {Screen most \\ relevant documents};
-  \node (stopscreen) [decision, join] {Irrelevant?};
+  \node (stopscreen) [decision, join] {Recall $\geq$ target?};
 
-  \node (screenr2) [screen, join]  {Screen random documents};
-
-  \node (recall) [decision, join] {Recall $>$ threshold?};
-
-	\node (sform) [form, right=of screenr2, xshift=-15mm] { Estimate $\hat{p}U$ };
-
-  \node (rform) [form, right=of recall, xshift=-15mm] {$ 1 - \frac{\hat{p}U}{\hat{p}P} \frac{U}{P}$};
-
-  \node (reldocs) [multidoc, below of=recall, xshift=20mm, yshift=-8mm]  {Relevant documents};
-  \node (irreldocs) [multidoc, fill=blue!5, below of=recall, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
+  \node (reldocs) [multidoc, below of=stopscreen, xshift=20mm, yshift=-8mm]  {Relevant documents};
+  \node (irreldocs) [multidoc, fill=blue!5, below of=stopscreen, xshift=-20mm, yshift=-8mm]  {Irrelevant documents};
   %\node
@@
 
-  \node [coord, left=of recall, xshift=-45mm] (c3)  {}; %\cmark{1}
-
  \path (stopscreen.west) to node [very near start, yshift=1em] {$n$} (c1);
@@
 
-  \path (recall.west) to node [very near start, yshift=1em] {$n$} (c3);
+ %\draw [o->,norm] (recall.west) -- (c3) |-  (c2) -- (model.north);
+% \draw [o->,norm] (recall.west) -- (c3) |-  (screenr2.west);
 
- %\draw [o->,norm] (recall.west) -- (c3) |-  (c2) -- (model.north);
- \draw [o->,norm] (recall.west) -- (c3) |-  (screenr2.west);
-
- \draw [o->,norm] (recall.south) -- (reldocs);
- \draw [o->,norm] (recall.south) -- (irreldocs);
+ \draw [o->,norm] (stopscreen.south) -- (reldocs);
+ \draw [o->,norm] (stopscreen.south) -- (irreldocs);
 

### 2020-09-08 pres/rapid-review-CEE-2020.tex
new presentation

\documentclass[9pt, aspectratio=169]{beamer}
\mode<presentation>
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{tcolorbox}

\usetheme{Boadilla}

\definecolor{mcctext}{HTML}{0d77b0}
\definecolor{mcclogo}{HTML}{0067a1}
\definecolor{mcclight}{HTML}{D9E8F1}

\usecolortheme[named=mcclogo]{structure}

\usefonttheme{professionalfonts}

\newtheorem*{remark}{}

\bibliographystyle{apalike}

\title[Stopping Criteria]{Statistical Stopping Criteria for Automated Screening in Systematic Reviews}
\subtitle{}
\author{Max Callaghan, Finn Müller-Hansen }
\institute[MCC]{
	\includegraphics[height=1cm,width=2cm]{images/MCC_Logo_RZ_rgb.jpg} \hspace{3em}
	\includegraphics[height=1cm]{images/University-of-Leeds-logo.png}
}
\date{September 26, 2019}

\begin{document}

\setbeamertemplate{caption}{\raggedright\insertcaption\par}

	\begin{frame}
	\titlepage
\end{frame}

\begin{frame}{Context - the promise of work savings through machine learning}

\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{itemize}
			\item When doing a systematic review, or other evidence synthesis project, you often need to screen a lot of documents.
			\item Large literature on training machine learning algorithms to recognise relevant documents and give these to us first \cite{OMara-Eves2015}.
			\item The promise of this field is that we can stop before screening all documents, saving humans work.
		\end{itemize}

		\medskip

		\textbf{But,}

		\begin{itemize}
			\item How do we know what a good time to stop is?
			\item How can we report this?
		\end{itemize}

	\end{column}
	\begin{column}{0.382\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{images/example-recall}
			\caption{Source: \cite{Przybya2018}}
		\end{figure}
	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Workflows for reducing human effort in systematic review screening}

\begin{columns}
	\begin{column}{0.6\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/flow_noeq_basic.pdf}
		\end{figure}
	\end{column}
	\begin{column}{0.4\linewidth}
		Existing approaches to deciding when to stop screening are unsatisfactory \cite{bannach-brown2019, Marshall2019}, but fall into the following categories

		\medskip

		\begin{itemize}
			\item \textbf{BIR Sampling:} Estimate the number of relevant documents via sampling, stop when you have seen that number \cite{Shemilt2014}
			\item \textbf{Heuristics:} Stop after [x] consecutive irrelevant documents \cite{Jonnalagadda2013, Przybya2018}
			\item \textbf{Novel automatic stopping criteria:} More sophisticated systems for automatically deciding when to stop screening \cite{Yu2019, DiNunzio2018, Howard2020}
		\end{itemize}

	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Baseline Inclusion Rate (BIR) Sampling based criteria}
\begin{columns}
	\begin{column}{0.382\linewidth}
		\begin{enumerate}
			\item Sample a fraction of large set of documents
			\item Estimate the number of relevant documents based on the number seen in the sample
			\item Screen until this number of relevant documents (or a proportion corresponding to target recall) has been seen.
		\end{enumerate}

		\begin{remark}
			Sampling error is not accounted for and can have serious consequences
		\end{remark}
	\end{column}
	\begin{column}{0.618\linewidth}
		\begin{figure}
		\only<1>{
		\includegraphics[width=\linewidth]{../manuscript/2_figs_bir_errors.pdf}
		\caption{\small Distribution of errors after a sample of 1,000 using the BIR sampling method in a dataset of 20,000 documents of which 500 are relevant. }}
		\only<2>{
	\includegraphics[width=\linewidth]{../manuscript/2_figs_bir_error_distribution.pdf}
	\caption{\small Distribution of errors across sample sizes using the BIR sampling method in a dataset of 20,000 documents of which 500 are relevant. }}
		\end{figure}

	\end{column}
\end{columns}
\end{frame}

\begin{frame}{Heuristic criteria}
\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{itemize}
			\item If we see several consecutive irrelevant documents, we can surmise that the proportion of relevant documents remaining is low
		\end{itemize}

		\begin{remark}<2->
			This intuition is helpful, but misses the fact that the proportion of relevant documents in the set of unseen documents can mean very different things
		\end{remark}
	\end{column}
	\begin{column}{0.382\linewidth}

		\begin{figure}
			\only<1->{\includegraphics[width=\linewidth]{../manuscript/2_figs_proportions_2}}
			\only<2->{\includegraphics[width=\linewidth]{../manuscript/2_figs_proportions_1}}
		\end{figure}

	\end{column}
\end{columns}
\end{frame}

\begin{frame}{Novel automatic stopping criteria}

Several recent papers \cite{Yu2019, DiNunzio2018, Howard2020} have suggested more sophisticated stopping criteria.

\medskip

None of these have the same \textit{fundamental} problems as those discussed, but

\begin{itemize}
	\item None account properly for uncertainty
	\item None deliver \textit{reliable} meeting of recall targets
	\item None offer users the ability to communicate when they stopped
\end{itemize}

\begin{remark}
	If we want to use machine learning in live systematic reviews, we need criteria that can communicate the probable outcome of stopping early in a way that is clear and independent of machine learning approach and performance
\end{remark}

\end{frame}

\begin{frame}{A statistical stopping criteria for Active Learning}

We test a null hypothesis that a given recall target has not been achieved, starting a random sample at an arbitrary point.

\textbf{Definitions}

\begin{itemize}
	\item $N_{tot}$ is the total number of documents
	\item $\rho_{tot}$ is the total number of relevant documents
	\item $\rho_{seen}$ is the number of relevant documents seen by a screener
	\item $\tau$, or recall is $\frac{\rho_{seen}}{\rho_{tot}}$
	\item $\tau_{tar}$ is our recall target
	\item $N_{AL}$ is the number of documents seen after active learning has finished (at start of random sample)
	\item $N$ is the number of documents in the sample ($N_{tot} - N_{AL}$)
	\item $K$ is the number of relevant documents in the sample ($\rho_{tot} - \rho_{AL}$)
\end{itemize}

After each draw from the sample:

\begin{itemize}
	\item $n$ is the number of documents drawn
	\item $k$ is the number of relevant drawn
\end{itemize}

\end{frame}

\begin{frame}{A statistical stopping criteria for Active Learning - II}

We form a null hypothesis that the target recall has not been achieved

\begin{equation}
H_0 : \tau < \tau_{tar}
\label{eq:null_hypothesis}
\end{equation}

Accordingly, our alternative hypothesis is that recall is at least as large as our target:

\begin{equation}
H_1 : \tau \geq \tau_{tar}
\end{equation}

Because we are sampling without replacement, we know that $k$ is distributed hypergeometrically:

\begin{equation}
k \sim Hypergeometric(N, K, n)
\end{equation}

\end{frame}

\begin{frame}{A statistical stopping criteria for Active Learning - III}

We introduce a hypothetical value for $K$, which we call $K_{tar}$. This represents the lowest value for $K$ compatible with $H_0$

\begin{equation}
K_{tar} = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
\end{equation}

An testable analogue of $H_0$ is therefore

\begin{equation}
H_0 : K \geq K_{tar}
\end{equation}

The cumulative distribution function gives us the probability of observing what we observed, if our null hypothesis were true

\begin{equation}
p = P(X \leq k) \text{, where } X \sim Hypergeometric(N,K_{tar},n)
\label{eq:p-value}
\end{equation}

When $p < 1-\alpha$, we can stop screening, and report, for example, that we reject the null hypothesis that we achieve a recall below 95\% at the 5\% significance level

\end{frame}

\begin{frame}{A statistical stopping criteria for active learning - When to start a random sample?}

\begin{columns}
	\begin{column}{0.382\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../images/flow.pdf}
		\end{figure}
	\end{column}
	\begin{column}{0.618\linewidth}
		In the approach described, we need to stop machine learning at some point and switch to random sampling
		\begin{itemize}
			\item If we stop too early, it will take us a long time to get to our recall target
			\item If we stop too late, we will save less work
		\end{itemize}

		\medskip

		We define a subcriterion, ``ranked quasi-sampling'', which treats previously screened documents as random samples.

		\begin{itemize}
			\item Ranked quasi-sampling is always conservative, as long as the proportion of relevant documents given to the screener by the machine learning algorithm $\geq$ the proportion in the remaining documents
			\item We test this as an independent criteria, but also use it to decide when to stop screening
		\end{itemize}
	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Evaluation}
\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{table}
			\footnotesize
			\input{../manuscript/1_tables_datasets.tex}
			\caption{Dataset properties}
			\label{tab:data}
		\end{table}
	\end{column}
	\begin{column}{0.382\linewidth}

		\begin{itemize}
			\item We assemble a dataset of systematic reviews which other systems have been tested on
			\item We simulate 100 reviews on each of these
			\item In each review, a sample is drawn to begin with, then every 10 documents:
			\begin{itemize}
				\item a simple SVM is trained on documents already seen
				\item the relevance of unseen documents is predicted
				\item the real relevance values of the next 10 documents are revealed
			\end{itemize}
			\item We record when each criteria would have been achieved
		\end{itemize}

	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Criteria performance - A priori knowledge}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_pf.pdf}

		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		This is what would happen if we already knew how many relevant documents there were.

			\medskip

			The target is never missed, and we can achieve some very large work savings.

			\medskip

			These are the results most often reported!

	\end{column}
\end{columns}
\end{frame}

\begin{frame}{Criteria performance - Baseline estimation}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_bir.pdf}
		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
Estimating the number of relevant results based on a sample means we often (drastically!) miss our target, and we often save no work at all

	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Criteria performance - Heuristics}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\only<1>{\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_ih_50.pdf}}
			\only<2>{\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_ih_200.pdf}}
		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		\only<1-2>{Stopping after 50 consecutive irrelevant results sometimes works, but often results in awful recall or no work savings}
		\only<2>{
			\medskip

			A stricter heuristic (200 consecutive irrelevant results) less often leads to poor recall, but also means less work is saved)
		}
	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Criteria performance - Statistical stopping criteria (random sampling)}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_hyper.pdf}

		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		Our basic criterion makes modest work savings possible with a reliable achievement of the recall target

	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Criteria performance - Statistical stopping criteria (ranked quasi-random sampling)}

\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_nrs.pdf}

		\end{figure}
	\end{column}
	\begin{column}{0.5\linewidth}
		Our criterion using ranked quasi-sampling turns out to be more conservative, and to allow slightly larger work savings, but this is not robust to catastrophic machine learning failure!

	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Discussion}

\begin{columns}
	\begin{column}{0.618\linewidth}
		 \begin{itemize}
		 	\item We provide reliable stopping criteria that realise \textit{some} of the work savings in systematic review screening promised by machine learning
		 	\item Previously the outcome of a badly performing machine learning model could have been low recall levels (as low as a few percent). Now only work savings are at risk.
		 	\item We can use these in live reviews, because we can report the implications of our decision to stop early for recall
		 \end{itemize}
	\end{column}
	\begin{column}{0.382\linewidth}
		Work savings are modest but
		\begin{itemize}
			\item We used very general machine learning models and did not tune the parameters to perform better on individual datasets
			\item Larger savings are possible in larger datasets (where work savings are most helpful!)
			\item Forthcoming work will investigate how work savings depend on data features -> savings calculator for new projects
		\end{itemize}
	\end{column}
\end{columns}

\begin{center}
	\line(1,0){250}

	\medskip

	\textbf{Thanks!}

	\medskip

	Working paper: \url{https://doi.org/10.21203/rs.2.18218/v2}

	Contact: \url{callaghan@mcc-berlin.net}, Twitter: @MaxCallaghan5

	Code: \url{https://github.com/mcallaghan/rapid-screening}

\end{center}

\end{frame}

\begin{frame}{References}
\tiny
\bibliography{../manuscript/3_bib_mendeley}
\end{frame}

\end{document}

### 2020-09-24 analysis/hyper_criteriaR.Rmd
r function

---
title: "Active learning stopping criteria"
output: html_notebook
---

In an active learning pipeline for screening citations, humans screen documents, machines learn from these screening decisions what a relevant document looks like, and humans keep screening the documents predicted by the machine to be most relevant.

If machine-learning can identify relevant documents effectively, human screeners would screen all or most of the relevant documents in a set of documents before screening all or most of that set of documents. In principle they could stop at that point and save themselves some work.

Because we don't know *a priori* the number of relevant documents, human screeners won't know when they have reached a safe place to stop.

One way to approach this problem is to develop a null hypothesis that a given recall target has been missed, and to see how likely this is. If we get a p-value for this test that is below a given threshold, human screeners could stop at a level of risk of missing documents they are comfortable with.

We can calculate a p-value for this null hypothesis in R with the following information:

```{r}
# A vector of already screened documents
screened <- c(rep(1,20), rep(0,70))

# The total number of documents
N <- 100

# The recall target
recall_target <- 0.9

```

To test this hypothesis we use the hypergeometric distribution, which is based on an urn of randomly mixed red and green balls, which we draw out one by one, without replacing them. In this analogy, green marbles are relevant documents, red marbles are irrelevant documents.

What we need to do first is to imagine the smallest number of green marbles that could be in the urn that would mean that our recall target had not yet been reached.

```{r}

# How many relevant documents have been seen
r_seen <- sum(screened)

# let's treat the last n documents as a random sample
n <- 70
sample <- tail(screened, n)

# How many relevant documents have been in seen a random sample (or set of documents we are happy to assume does not have a systematically lower proportion of relevant documents)
x <- sum(sample)

# How many relevant documents were seen before this sample or psuedosample began
r_ML <- r_seen - x

# what's the minimum number of relevant documents there would have to be in the unseen documents + the sample in order for our target to be missed
get_khat <- function(r_ML, r_seen, recall_target){
  return(floor(r_seen/recall_target-r_ML+1))
}

k_hat <- get_khat(r_ML, r_seen, recall_target)

print(k_hat)

```

Now we can see what the chances are of pulling out the number of green marbles from the urn that we did in our sample, given an urn with the amount of green marbles that would mean that our target was missed

```{r}
p <- phyper(x, k_hat, N-(length(screened)-length(sample))-k_hat, length(sample))
print(p)
```

In a real application, we want to test this for all past urn drawings (the last 1 draws, the last 2 draws, the last 3 draws, etc.) and find the minimum probability of our null hypothesis, and the number of balls drawn. If we can assume that the proportion of green marbles in that draw is greater than or equal to the proportion of green marbles left in the urn, then this p-value is conservative.

```{r}
h0_p <- function(screened, N, recall_target) {

  # How many relevant documents have been seen
  r_seen <- sum(screened)

  # Sample size
  k_vec <- seq(1:length(screened))
  # number of relevant documents in sample
  x_vec <- cumsum(rev(screened))
  # number of relevant documents seen before sampling
  r_ML_vec <- r_seen - x_vec

  # k_hat over sample sizes
  k_hat_vec <- vapply(r_ML_vec, get_khat, numeric(1), recall_target=recall_target, r_seen=r_seen)

  # corresponding irrelevant documents in sample+unseen documents
  n_vec <- N-(length(screened)-k_vec)-k_hat_vec

  p_vec <- phyper(x_vec, k_hat_vec, n_vec, k_vec)

  n <- which.min(p_vec)

  return(list("min_p" = p_vec[n], "n" = n, "p_vec" = p_vec))
}

h0 <- h0_p(screened, N, recall_target)

print(paste0("Based on the last ",h0$n, " documents, the p-value for our null hypothesis that our recall is below ", recall_target, " is ", h0$min_p))

```

### 2020-09-24 analysis/hyper_criteriaR.Rmd
add missing function within functgion
rmd

@@
 title: "Active learning stopping criteria"
-output: html_notebook
+output: html_document
 ---
@@
 
+  # what's the minimum number of relevant documents there would have to be in the 
+  # unseen documents + the sample in order for our target to be missed
+  get_khat <- function(r_ML, r_seen, recall_target){
+    return(floor(r_seen/recall_target-r_ML+1))
+  }
+
   # Sample size

### 2021-01-19 pres/rapid-review-esmar.Rmd
esmarc presentation

---
author: Max Callaghan, Finn Müller-Hansen
date: January 21, 2021
output: 
    beamer_presentation:
        latex_engine: pdflatex
        includes:
            in_header: header.tex

bibliography: ../mendeley.bib
classoption: "aspectratio=169"

---

```{r setup, include=FALSE}
set.seed(5)
knitr::opts_chunk$set(cache = F)
knitr::opts_chunk$set(global.par = TRUE)
options(max.print = 20)
xlab <- "Documents Screened" 
ylab <- "Relevant documents identified"
```
##

\maketitle

## Introduction
- Identifying studies for an evidence synthesis project can be very time-consuming
- Machine learning (ML) can be trained using human decisions to predict the relevance of unseen documents
- Active learning means iterations of
    - Humans screen documents
    - ML is trained using human decisions
    - ML predicts relevance probability for all remaining documents
    - Documents are ordered in descending order of predicted relevance
- With active learning we can identify all relevant documents without having to view every single document.

## Demonstrable work savings

Numerous studies have used datasets from previous reviews to show that machine learning has the potential to save work in systematic review [@OMara-Eves2015; @Cohen2006; @Przybya2018].

However, most of the work savings are just *potential* work savings predicated on *a priori* knowledge of how many relevant studies we are looking for.

To achieve these savings, we need to know when to stop screening: and existing approaches to deciding this are insufficient [@Shemilt2014; @Howard2020; @Jonnalagadda2013]. We need to to confident that we don't miss studies, and we need to be able to communicate this confidence.

## Screening documents

We can set up a toy dataset with 1,000 documents ($N_{tot}$), of which 100 are relevant ($\rho_{tot}$)

```{r sample}
N_tot <- 2000 # total documents
r_tot <- 100 # total relevant documents

docs <- rep(0,N_tot)
docs[1:r_tot] <- 1
docs <- sample(docs)
docs

```

## Screening documents

At any point during screening, we can calculate recall ($\tau$) by dividing the number of relevant documents we have seen ($\rho_{seen}$) by the total number of relevant documents

\[ \tau = \frac{\rho_{seen}}{\rho_{tot}}\]

We usually accept that we won't identify every last document and aim for a recall target ($\tau_{tar}$) like 95%

## Screening documents

If we screen these at random, we will see 95% recall after seeing ~95% of all relevant documents

\columnsbegin

\column{0.5\textwidth}
\scriptsize
```{r randomPlotCode, eval=F, size="footnotesize"}
par(pty="s")
plot(
  cumsum(docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified",
  xlim=c(0,N_tot),
  ylim=c()
)

```
```{r}
tau_target=0.95
tau <- cumsum(docs)/r_tot
which(tau>tau_target)[1]/N_tot
```

\column{0.5\textwidth}
```{r randomPlotOut, echo=F, size="footnotesize"}
par(pty="s")
plot(
  cumsum(docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified",
  xlim=c(0,N_tot),
  ylim=c()
)
abline(h=r_tot*.95)

```

\columnsend

## Screening documents with machine learning

\columnsbegin

\column{0.5\textwidth}

\small
If machine learning means we are more likely to view a relevant document in each draw, we will achieve 95% recall much earlier
\scriptsize
```{r eval=F}
weights = rep(1,N_tot)
weights[which(docs==1)] <- seq(
  10,2,length.out=r_tot
)
ordered_docs <- sample(
  docs, prob=weights, replace=F
)
plot(cumsum(ordered_docs))
abline(h=r_tot*.95)
```
```{r echo=F}
weights = rep(1,N_tot)
weights[which(docs==1)] <- seq(15,2,length.out=r_tot)
ordered_docs <- sample(docs, prob=weights, replace=F)
```

```{r}
tau <- cumsum(ordered_docs)/r_tot
which(tau>=tau_target)[1]/N_tot
```

\column{0.5\textwidth}
```{r echo=F}
par(pty="s")
plot(
  cumsum(ordered_docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
abline(h=r_tot*.95)
```
\columnsend

## Screening documents with machine learning

\columnsbegin

\column{0.5\textwidth}

It's also easy to tell when 95% recall has been achieved when we know how many relevant documents there are

\column{0.5\textwidth}
```{r echo=F}
par(pty="s")
plot(
  cumsum(ordered_docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
abline(h=r_tot*.95)
```

\columnsend

## Screening documents with machine learning

But if we descale the y axis, it's not so easy to tell when we have achieved 95% recall

\columnsbegin
\column{0.5\textwidth}
\scriptsize
```{r eval=F}
r <- rle(ordered_docs) # calculate run lengths
cutoff <- cumsum(r$lengths)[
  r$lengths>40 & r$values==0
][1] # stop after 30 consecutive 0s
plot(
  cumsum(ordered_docs[1:cutoff]),
  xlim=c(1,N_tot),
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
```
\column{0.5\textwidth}
```{r echo=F}
par(pty="s")
r <- rle(ordered_docs) # calculate run lengths
cutoff <- cumsum(r$lengths)[r$lengths>40 & r$values==0][1]
plot(
  cumsum(ordered_docs[1:cutoff]),
  xlim=c(1,N_tot),
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
```

\columnsend

## Screening documents with machine learning

With a bit of maths, we can calculate when it is safe to stop, if we want to achieve 95% recall (or any other target) with a given level of confidence

\columnsbegin
\column{0.5\textwidth}

```{r echo=F}
par(pty="s")
plot(
  cumsum(ordered_docs), 
  xlab="Documents Screened", 
  ylab="Relevant documents identified"
)
abline(h=r_tot*.95)
```

\column{0.5\textwidth}

In a nutshell, we use the distribution of relevant documents in those we previously screened, to infer likely distributions of relevant documents in documents not yet seen.

\columnsend

## A statistical stopping criterion

Let's imagine we stop machine-learning prioritised screening at an arbitrary point and start drawing *at random* **without replacement** from the remaining documents.

\columnsbegin
\column{0.2\textwidth}
![](images/Keats_urn.jpg)

\column{0.8\textwidth}
In probability theory, we use the analogy of an urn with green marbles (relevant documents / successes) and red marbles (irrelevant documents / failures).

Using the hypergeometric distribution. we can calculate the probability of drawing

- $k$ relevant documents  
- in a sample of $n$ documents
- from an urn that had $N$ total documents
- of which $K$ were relevant

\columnsend

## A statistical stopping criterion

Let's return to a point that looked like a nice place to stop

\vspace{1em}

\tiny

\columnsbegin
\column{0.4\textwidth}
```{r echo=F}
par(pty="s")
r_al <- cumsum(ordered_docs[1:cutoff])[cutoff]
N_al <- cutoff
K <- r_tot-r_al 
N <- N_tot - N_al
```

```{r}
plot(
  cumsum(ordered_docs[1:cutoff]),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab
)
```
\column{0.6\textwidth}
$N_{AL}$ is the number of documents screened after active learning, and $\rho_{AL}$ is the number of relevant documents seen after active learning. So $N$ ($K$) is the number of (relevant) documents remaining "in the urn".
\tiny
```{r}
params <- list(
  N_al = cutoff,
  N = N_tot - N_al,
  r_al = cumsum(ordered_docs[1:cutoff])[cutoff],
  K = r_tot-r_al 
)
print(params)
```

\columnsend

## A statistical stopping criterion

We begin with a sample of 100 documents

\tiny

\columnsbegin
\column{0.5\textwidth}
```{r}
n <- 100
urndocs <- sample(ordered_docs[cutoff:N_tot], replace=F)
s_docs <- urndocs[1:n]
k <- cumsum(s_docs)[n]
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(s_docs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",n))
)

```
\column{0.5\textwidth}
$n$ is the number of documents in the sample (`r n`) and $k$ is the number of relevant documents drawn (`r k`)

\medskip

Using the hypergeometric distribution, we can calculate the probability of observing $k$ relevant documents in a sample of $n$ documents given an urn of $N$ documents of which $K$ are relevant.

```{r}
phyper(k, K, N, n)
```
However, although we do know $N$, we don't know how many documents are in the urn ($K$).

\columnsend

## A statistical stopping criterion - Hypothesis testing

\tiny

\columnsbegin
\column{0.5\textwidth}
```{r}
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(s_docs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",n))
)

```
\column{0.5\textwidth}
We form a null hypothesis that the target level of recall has not been achieved

\begin{equation}
H_0 : \tau < \tau_{tar}
\end{equation}

Accordingly, our alternative hypothesis is that recall is at least as large as our target:

\begin{equation}
H_1 : \tau \geq \tau_{tar}
\end{equation}

To operationalise this, we come up with a hypothetical value of $K$ which is the lowest value compatible with our null hypothesis

\begin{equation}
K_{tar} = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
\end{equation}

```{r}
get_ktar <- function(r_al, r_seen, recall_target){
  return(floor(r_seen/recall_target-r_al+1))
}
r_seen <- r_al+k
K_tar <- get_ktar(r_al, r_seen, .95)
```

In other words, if there were `r K_tar` or more relevant documents in the urn when sampling began, the `r r_al` relevant we identified before sampling, and the `r k` we drew from the urn would not be enough to meet our target recall level

\columnsend

## A statistical stopping criterion - Hypothesis testing

\tiny

\columnsbegin
\column{0.5\textwidth}
```{r}
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(s_docs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",n))
)

```
\column{0.5\textwidth}
The cumulative distribution function gives us the probability of observing what we observed, if our null hypothesis were true

\begin{equation}
p = P(X \leq k) \text{, where } X \sim Hypergeometric(N,K_{tar},n)
\label{eq:p-value}
\end{equation}

When $p < 1-\alpha$, we can stop screening, and report, for example, that we reject the null hypothesis that we achieve a recall below 95\% at the 5\% significance level

```{r}
p <- phyper(k, K_tar, N-K_tar, n)
print(k)
p
```
Based on what we observed in this sample, it is not unlikely at all that we have missed our target, so we keep screening

\columnsend

## A statistical stopping criterion - Hypothesis testing

\tiny

\columnsbegin
\column{0.5\textwidth}
```{r}
vec_K_tar <- vapply(
  cumsum(urndocs)+r_al, get_ktar, numeric(1), 
  r_al=r_al, recall_target=.95
)
p <- phyper(cumsum(urndocs), vec_K_tar, N-vec_K_tar, seq(1,N))
plot(
  c(cumsum(ordered_docs[1:cutoff]),cumsum(urndocs)+r_al),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab,
  col=c(rep("black",cutoff),rep("red",N))
)
abline(h=r_tot*.95)
abline(v=N_al+which(p<0.05)[1])

```
\column{0.5\textwidth}
The p score declines as we see more consecutive irrelevant documents, and as there are fewer and fewer documents left in the urn.

```{r}
plot(c(rep(NA,N_al),p), xlab="Documents screened")
abline(h=.05)
abline(v=N_al+which(p<0.05)[1])
```

\columnsend

## Ranked quasi sampling

\footnotesize

This is useful but not practical, because we don't know when to start a random sample, and doing so slows the whole process down.

-> We treat previously screened documents *as if* they were drawn from a random sample, which is conservative as long as the machine learning hasn't completely backfired.

\columnsbegin
\column{0.5\textwidth}
\tiny
```{r eval=F}
last_docs <- c(0,0,0,1,0,0,1,0,0,1,0,1,0,1,0,0,1,0,0,0)
cols <- ifelse(last_docs>0,"green","red")
plot(
  rep(1,20), col = cols, pch = 19,
  ylim=c(0.5,4.5), axes=FALSE, xlab="",ylab=""
)
for (i in seq(1,6)) {
  lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
  i_k <- cumsum(tail(last_docs,i))[i]
  text(
    20.5-i*0.5,0.75+i*0.4,
    paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25)
  )
}
```
\column{0.5\textwidth}
```{r echo=F}
last_docs <- c(0,0,0,1,0,0,1,0,0,1,0,1,0,1,0,0,1,0,0,0)
cols <- ifelse(last_docs>0,"green","red")
plot(rep(1,20), col = cols, pch = 19,ylim=c(0.5,4.5), axes=FALSE, xlab="",ylab="")
for (i in seq(1,6)) {
  lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
  i_k <- cumsum(tail(last_docs,i))[i]
  text(20.5-i*0.5,0.75+i*0.4,paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25))
}
```
\columnsend

## Ranked quasi sampling

\tiny

\columnsbegin
\column{0.5\textwidth}
```{r}
plot(
  cumsum(ordered_docs[1:cutoff]),
  xlim=c(1,N_tot), xlab=xlab, ylab=ylab
)
```
We can then calculate a p value for quasi-samples consisting of the last 1 documents to all screened documents

\column{0.5\textwidth}

```{r fig.height=5}
h0_p <- function(docs, N_tot, recall_target) {
  r_seen <- sum(docs)
  n_vec <- seq(1:length(docs))
  k_vec <- cumsum(rev(docs))
  r_al_vec <- r_seen - k_vec
  k_hat_vec <- vapply(r_al_vec, get_ktar, numeric(1), 
                    recall_target=recall_target, r_seen=r_seen)
  red_ball_vec <- N_tot-(length(docs)-n_vec)-k_hat_vec
  p_vec <- phyper(k_vec, k_hat_vec, red_ball_vec, n_vec) 
  n <- which.min(p_vec)
  return(list("min_p" = p_vec[n], "n" = n, "p_vec" = p_vec))
}
plot(h0_p(ordered_docs[1:cutoff],N_tot,.95)$p_vec,
     xlab="Documents in quasi sample",ylab="p")
```

\columnsend

## Ranked quasi sampling

\tiny

\columnsbegin
\column{0.5\textwidth}

We can calculate this after each document is screened and figure out when it is safe to stop

```{r warning=F, fig.height=5, fig.width=5, eval=F}
df <- data.frame(
  i=seq(1,N_tot), r_seen=cumsum(ordered_docs), 
  docs=ordered_docs, pmin=0
)
for (i in df$i) {
  p <- h0_p(df$docs[1:i], N_tot, .95)
  df$pmin[i] <- p$min_p
}
par(mfrow = c(2,1), mar = c(2, 2, 2, 2),pty="s")
plot(df$r_seen)
plot(df$pmin)
```
\column{0.5\textwidth}

```{r warning=F, fig.height=6, echo=F}
df <- data.frame(i=seq(1,N_tot), r_seen=cumsum(ordered_docs), docs=ordered_docs, pmin=0)
for (i in df$i) {
  p <- h0_p(df$docs[1:i], N_tot, .95)
  df$pmin[i] <- p$min_p
}
par(mfrow = c(2,1), mar = c(2, 2, 2, 2), pty="s")
plot(df$r_seen, ylab="Relevant documents identified")
plot(df$pmin, ylab="p H0", xlab="Documents screened")
```
\columnsend

## Evaluation

We tested this on 20 systematic review datasets, simulating 100 machine learning assisted reviews of each.

\columnsbegin
\column{0.382\textwidth}
![](../manuscript/2_figs_jointplot_nrs.pdf)
\column{0.618\textwidth}
Our criterion performed reliably. Although work savings were small, this is partly due to small datasets with few relevant documents. Because of the way the hypergeometric distribution works, the extra effort required to prove that it is safe to stop gets proportionally smaller.
\columnsend

## Conclusion

We provide a reliable way to estimate if it is safe to stop screening if you want to achieve a certain recall target. All you need are the R function in this presentation, a vector of 1s and 0s representing human screening decisions, and the number of documents in a database

\medskip

\centering

**Thanks!**

Paper: Callaghan, M.W., Müller-Hansen, F.. Statistical Stopping Criteria for Automated Screening in Systematic Reviews. \textit{Systematic Reviews}. \textbf{9} 273 (2020). \url{https://doi.org/10.1186/s13643-020-01521-4}.

Contact: \url{callaghan@mcc-berlin.net}, Twitter: \@MaxCallaghan5

Code: \url{https://github.com/mcallaghan/rapid-screening}

This presentation: \url{https://github.com/mcallaghan/rapid-screening/pres/rapid-review-esmar.Rmd}

## References

\scriptsize

### 2022-10-19 pres/WWGS.tex
WWGS presentation

\documentclass[9pt,aspectratio=169]{beamer}
\mode<presentation>
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usepackage{xmpmulti}
\usepackage{animate}
\usepackage{tcolorbox}
\usepackage{amsmath}
\usepackage{xcolor, soul}
\usepackage{gensymb}
\usepackage{appendixnumberbeamer}

%\definecolor{links}{HTML}{2A1B81}
%\hypersetup{urlcolor=links} % Does not apply color to href's
%\hypersetup{colorlinks,urlcolor=links}

\usetheme{Boadilla}
%\usecolortheme{seahorse}

\usefonttheme{professionalfonts}

\title[Stopping Criteria]{Statistical stopping criteria for automated screening in systematic reviews}
\date{20 October 2022}
\author{Finn Müller-Hansen, Max Callaghan}
\institute[MCC]{
	\includegraphics[height=1cm,width=2cm]{images/MCC_Logo_RZ_rgb.jpg}
}

\newtheorem*{remark}{}

\bibliographystyle{apalike}

\begin{document}
	\begin{frame}
	\titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%
% Motivation
%\begin{frame}{Motivation}
%
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%
% RITL
\begin{frame}{Researcher in the Loop process}

\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{itemize}
			\item<1-> A large literature \citep{omara-eves_using_2015} has developed ``human-in the loop'' machine learning applications which ``overcome the manual and time-consuming screening of large numbers of studies by prioritizing relevant studies via active learning''  \cite{van_de_schoot_open_2021}
			\item<2-> Identifying 95\% of relevant documents if screening manually would mean screening 95\% of all documents
			\item<3-> If we can use machine learning to prioritise documents likely to be relevant, we can achieve high levels of recall with low levels of effort.
			\item<4-> BUT, given we do not know \textit{a priori} the true number of relevant documents, we need strategies to stop screening and bank the work savings
			\item<5->Getting these wrong can mean missing our target
		\end{itemize}
	\end{column}
	\begin{column}{0.382\linewidth}
		\begin{figure}
			\only<2-3>{\includegraphics[width=\linewidth]{images/idealised_stopping.pdf}}
			\only<4>{\includegraphics[width=\linewidth]{images/potential_stopping.pdf}}
			\only<5>{\includegraphics[width=\linewidth]{images/early_stopping.pdf}}
%			\only<2>{\includegraphics[width=\linewidth]{images/random.pdf}}
%			\only<3>{\includegraphics[width=\linewidth]{images/ml_prio.pdf}}
%			\only<4>{\includegraphics[width=\linewidth]{images/ml_prio.pdf}}
		\end{figure}
	\end{column}
\end{columns}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%
% Urn
\begin{frame}{A stopping criterion}

\begin{columns}

	\begin{column}{0.382\linewidth}
		\begin{figure}
			\only<1-3>{\includegraphics[width=.8\linewidth]{images/Keats_urn.jpg}}
			\only<4->{\includegraphics[width=\linewidth]{../images/h0_paths_ProtonBeam.pdf}}

		\end{figure}
	\end{column}
	\begin{column}{0.618\linewidth}
		\begin{itemize}
			\item<1-> Our stopping criterin works by treating documents as if they were white (not relevant) and red (relevant) marbles drawn from an urn without replacement.
			\item<2-> The hypergeometric distribution tells us the probability of observing \textbf{k} red marbles in a sample of \textbf{n} marbles, given an urn with \textbf{N} marbles, of which \textbf{K} are relevant
			\item<3-> We reformulate this to generate a null hypothesis $H_0$ that a given recall target has been \textbf{missed}. 
			\item<4-> We calculate a p-score for our null hypothesis, and if this is low enough, we reject $H_0$ and stop screening.
		\end{itemize}
	\end{column}

\end{columns}

\medskip

\bigskip

\only<5->{Note, ML-prioritisation means documents are not drawn at random, which makes our test conservative as long as ML works as well as or better than random chance.}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%
% Results
\begin{frame}{Results}

We test our criteria against other commonly used criteria on 20 complete systematic review datasets

\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{itemize}
			\item<1-> \textit{Potential} work savings (if we already knew when to stop) varied widely (higher for larger datasets - blue dots)
			\item<2-> Existing stopping criteria can result in catastrophic errors
			\item<4-> Our criteria generated work savings with reliably conservative performance wrt our recall target.
		\end{itemize}
	\end{column}
	\begin{column}{0.382\linewidth}
		\begin{figure}
			\only<1>{\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_pf.pdf}
				\caption{A priori knowledge}
			}
			\only<2>{\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_ih_50.pdf}
				\caption{50 consecutive irrelevant articles}
			}
			\only<3>{\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_bir.pdf}
				\caption{Estimating baseline inclusion rate}
			}
			\only<4>{\includegraphics[width=\linewidth]{../manuscript/2_figs_jointplot_nrs.pdf}
				\caption{Our criterion}
			}

		\end{figure}
	\end{column}
\end{columns}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%
% Conclusion
\begin{frame}{Conclusion}

We provide a stopping criteria that works on any model, with any tool: \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteriaR.md}.

\medskip

Work savings in practice with large datasets are large!

\medskip

Future work will identify how biased our urn is, in order to use a noncentral hypergeometric distribution, which should give a more precise, less conservative criterion.

\begin{center}
	\line(1,0){250}

	\medskip

	\textbf{Thanks!}

	\medskip

	Contact: \url{mueller-hansen@mcc-berlin.net, callaghan@mcc-berlin.net}

	Twitter: \url{https://twitter.com/MaxCallaghan5}

\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%

\appendix

\begin{frame}{References}

\bibliography{rapid-review}

\end{frame}

\begin{frame}{Applications and extensions}

\begin{columns}

\column{0.33\textwidth}
\only<1->{
	\begin{figure}
		\includegraphics[width=\linewidth]{images/doebbeling_ws.png}
	\end{figure}
}

\column{0.33\textwidth}
\only<2->{
	\begin{figure}
		\includegraphics[width=\linewidth]{images/doebbeling_p.png}
	\end{figure}
}

\column{0.33\textwidth}

\small

\begin{itemize}
	\item<1-> We have used the stopping criteria to generate massive savings (77\%) in real projects
	\item<2-> If rejecting our $H_0$ was less labour intensive we could have saved around 82\%
	\item<3-> Using a biased urn could help create a more precise criterion
\end{itemize}

\end{columns}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%
% Other criteria
\begin{frame}{Baseline Inclusion Rate}

\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{itemize}
			\item<1-> If we try to estimate the baseline inclusion rate, we will get it wrong most of the time. Overestimating results in 0 work savings, while underestimating results in less than target recall.
			\item<2-> Wrongness decreases with larger sample sizes, but bad outcomes remain most frequent.
		\end{itemize}
	\end{column}
	\begin{column}{0.382\linewidth}
		\begin{figure}
			\only<1>{\includegraphics[width=\linewidth]{../manuscript/2_figs_bir_errors.pdf}
			}
			\only<2>{\includegraphics[width=\linewidth]{../manuscript/2_figs_bir_error_distribution.pdf}
}

		\end{figure}
	\end{column}
\end{columns}

\end{frame}

\begin{frame}{Theory I}
We form a null hypothesis that the target level of recall has not been achieved

\begin{equation}
H_0 : \tau < \tau_{tar}
\end{equation}

To operationalise this, we come up with a hypothetical value of $K$ which is the lowest value compatible with our null hypothesis

\begin{equation}
K_{tar} = \lfloor \frac{\rho_{seen}}{\tau_{tar}}-\rho_{AL}+1 \rfloor
\end{equation}

In other words, if there were $K_{tar}$ or more relevant documents in the urn when sampling began, the $\rho_{al}$ relevant we identified before sampling, and the $k$ we drew from the urn would not be enough to meet our target recall level.

\medskip

The cumulative distribution function gives us the probability of observing what we observed, if our null hypothesis were true

\begin{equation}
p = P(X \leq k) \text{, where } X \sim Hypergeometric(N,K_{tar},n)
\label{eq:p-value}
\end{equation}

\end{frame}

\end{document}

### 2022-10-19 pres/rapid-review-2022.Rmd
WWGS presentation

--- 2021-01-19 pres/rapid-review-esmar.Rmd
@@
 author: Max Callaghan, Finn Müller-Hansen
-date: January 21, 2021
+date: Wed 28 September 2022
 output: 
@@
 knitr::opts_chunk$set(global.par = TRUE)
-options(max.print = 20)
+options(max.print = 35)
 xlab <- "Documents Screened" 
@@
 To achieve these savings, we need to know when to stop screening: and existing approaches to deciding this are insufficient [@Shemilt2014; @Howard2020; @Jonnalagadda2013].
-We need to to confident that we don't miss studies, and we need to be able to communicate this confidence.
+We need to be confident that we don't miss studies, and we need to be able to communicate this confidence.
 
@@
 
-We can set up a toy dataset with 1,000 documents ($N_{tot}$), of which 100 are relevant ($\rho_{tot}$)
+We can set up a toy dataset with 2,000 documents ($N_{tot}$), of which 100 are relevant ($\rho_{tot}$)
 
@@
 docs[1:r_tot] <- 1
-docs <- sample(docs)
+docs <- sample(docs, replace=F)
 docs
+sum(docs)
 
@@
 
-If we screen these at random, we will see 95% recall after seeing ~95% of all relevant documents
+If we screen these at random, we will see 95% recall after seeing ~95% of all documents
 
@@
 
-\columnsbegin
-
-\column{0.5\textwidth}
-
-It's also easy to tell when 95% recall has been achieved when we know how many relevant documents there are
-
-\column{0.5\textwidth}
-```{r echo=F}
-par(pty="s")
-plot(
-  cumsum(ordered_docs), 
-  xlab="Documents Screened", 
-  ylab="Relevant documents identified"
-)
-abline(h=r_tot*.95)
-```
-
-\columnsend
-
-## Screening documents with machine learning
-
 But if we descale the y axis, it's not so easy to tell when we have achieved 95% recall
@@
 par(pty="s")
+cols = c(rep("black",cutoff),rep("red",N_tot-cutoff))
 plot(
@@
   xlab="Documents Screened", 
-  ylab="Relevant documents identified"
+  ylab="Relevant documents identified",
+  col=cols
 )
@@
 par(pty="s")
-r_al <- cumsum(ordered_docs[1:cutoff])[cutoff]
+r_al <- sum(ordered_docs[1:cutoff])
 N_al <- cutoff
@@
   N = N_tot - N_al,
-  r_al = cumsum(ordered_docs[1:cutoff])[cutoff],
+  r_al = sum(ordered_docs[1:cutoff]),
   K = r_tot-r_al 
@@
 s_docs <- urndocs[1:n]
-k <- cumsum(s_docs)[n]
+k <- sum(s_docs)
 plot(
@@
 \column{0.5\textwidth}
-```{r}
+```{r fig.height=6}
 vec_K_tar <- vapply(
@@
 
-```{r}
-plot(c(rep(NA,N_al),p), xlab="Documents screened")
+```{r fig.height=6}
+plot(c(rep(NA,N_al),p), xlab="Documents screened", ylab="p")
 abline(h=.05)
@@
   lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
-  i_k <- cumsum(tail(last_docs,i))[i]
+  i_k <- sum(tail(last_docs,i))
   text(
@@
   lines(seq(20.5-i,20.5),rep(0.75+i*0.4,i+1))
-  i_k <- cumsum(tail(last_docs,i))[i]
+  i_k <- sum(tail(last_docs,i))
   text(20.5-i*0.5,0.75+i*0.4,paste0("n=",i,", k=",i_k), adj=c(0.5,-0.25))
@@
 
+## Applications and extensions
+
+\columnsbegin
+
+\column{0.33\textwidth}
+\only<1->{
+\begin{figure}
+\includegraphics[width=\linewidth]{images/doebbeling_ws.png}
+\end{figure}
+}
+
+\column{0.33\textwidth}
+\only<2->{
+\begin{figure}
+\includegraphics[width=\linewidth]{images/doebbeling_p.png}
+\end{figure}
+}
+
+\column{0.33\textwidth}
+
+\small
+
+\begin{itemize}
+  \item<1-> We have used the stopping criteria to generate massive savings (77\%) in real projects
+  \item<2-> If rejecting our $H_0$ was less labour intensive we could have saved around 82\%
+  \item<3-> Assuming documents are drawn at random makes our criterion over-conservative -> making more realistic assumptions (perhaps using the predictions from the models) might make it more precise.
+\end{itemize}
+
+\columnsend
+
+## Open questions
+
+\begin{itemize}
+  \item<1-> Are the x\% we miss a representative sample?
+  \item<2-> Can we use predictions to generate a precise estimate?
+  \item<3-> Are we there yet?
+Can we infer the relationship between \textbf{Dataset size}, \textbf{Machine learning effectiveness}, \textbf{prevalence of relevant documents}, and \textbf{work savings}, and use this to make predictions about potential work savings?.
+\end{itemize}
+
 ## Conclusion
@@
 
-This presentation: \url{https://github.com/mcallaghan/rapid-screening/pres/rapid-review-esmar.Rmd}
+This presentation: \url{https://github.com/mcallaghan/rapid-screening/blob/master/pres/rapid-review-esmar.Rmd}
 

### 2022-10-19 analysis/hyper_criteria_scenarios.R

@@ -1,0 +1,223 @@
+library(foreach)
+library(doParallel)
+
+N_tot <- 60000 # total documents
+r_tot <- N_tot*0.01 # total relevant documents
+
+
+docs <- rep(0,N_tot)
+docs[1:r_tot] <- 1
+docs <- sample(docs, replace=F)
+
+par(pty="s")
+plot(
+  cumsum(docs), 
+  xlab="Documents Screened", 
+  ylab="Relevant documents identified",
+  xlim=c(0,N_tot),
+  ylim=c()
+)
+
+
+tau_target=0.95
+tau <- cumsum(docs)/r_tot
+which(tau>tau_target)[1]/N_tot
+
+dev.off()
+par(mfrow = c(1,3),pty="s")
+
+
+
+get_ktar <- function(r_al, r_seen, recall_target){
+  return(floor(r_seen/recall_target-r_al+1))
+}
+
+h0_p <- function(docs, N_tot, recall_target, verbose=FALSE) {
+  r_seen <- sum(docs)
+  n_vec <- seq(1:length(docs))
+  k_vec <- cumsum(rev(docs))
+  r_al_vec <- r_seen - k_vec
+  k_hat_vec <- vapply(r_al_vec, get_ktar, numeric(1), 
+                      recall_target=recall_target, r_seen=r_seen)
+  red_ball_vec <- N_tot-(length(docs)-n_vec)-k_hat_vec
+  p_vec <- phyper(k_vec, k_hat_vec, red_ball_vec, n_vec)
+  #p_vec <- phyperPeizer(k_vec, k_hat_vec, red_ball_vec, n_vec)
+  n <- which.min(p_vec)
+  if (verbose) {
+    return(list("min_p" = p_vec[n], "n" = n, "p_vec" = p_vec))
+  } else {
+    return(p_vec[n])
+  }
+}
+
+ptm <- proc.time()
+
+df <- data.frame(
+  i=seq(1,N_tot), r_seen=cumsum(ordered_docs), 
+  docs=ordered_docs, pmin=0
+)
+
+doc_sets <- list()
+
+cl <- makeCluster(4) #not to overload your computer
+registerDoParallel(cl)
+
+for (w in c(15,10,5)) {
+  weights = rep(1,N_tot)
+  weights[which(docs==1)] <- rep(
+    w,length.out=r_tot
+  )
+  ordered_docs <- sample(
[…]

### 2023-03-14 pres/stopping_2023.tex
biased pres

--- 2022-10-19 pres/WWGS.tex
@@
 \date{20 October 2022}
-\author{Finn Müller-Hansen, Max Callaghan}
+\author{Max Callaghan, Finn Müller-Hansen}
 \institute[MCC]{
@@
 
-%%%%%%%%%%%%%%%%%%%%%%
-% Conclusion
-\begin{frame}{Conclusion}
-
-We provide a stopping criteria that works on any model, with any tool: \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteriaR.md}.
-
-\medskip
-
-Work savings in practice with large datasets are large!
-
-\medskip
-
-Future work will identify how biased our urn is, in order to use a noncentral hypergeometric distribution, which should give a more precise, less conservative criterion.
-
-\begin{center}
-	\line(1,0){250}
-
-	\medskip
-
-	\textbf{Thanks!}
-
-	\medskip
-
-	Contact: \url{mueller-hansen@mcc-berlin.net, callaghan@mcc-berlin.net}
-
-	Twitter: \url{https://twitter.com/MaxCallaghan5}
-
-\end{center}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\begin{frame}{References}
-
-\bibliography{rapid-review}
-
-\end{frame}
-
 \begin{frame}{Applications and extensions}
@@
 	\end{columns}
+
+\end{frame}
+
+\begin{frame}{Biased urns}
+
+	\begin{columns}
+
+		\column{0.382\textwidth}
+		\only<1->{
+			\begin{figure}
+				\includegraphics[width=\linewidth]{../../biased-rapid-review/plots/bias.pdf}
+			\end{figure}
+		}
+
+		\column{0.6\textwidth}
+
+		\small
+
+		\begin{itemize}
+			\item<1-> Our original criteria treats documents as if they were drawn at random from an urn
+			\item<2-> In fact, they are drawn in descending order of their predicted relevance
+			\item<3-> Using a non-central hypergeometric distribution \citep{fog_calculation_2008}, we can input a \textbf{bias} parameter indicating how much more likely we are to draw a random relevant than a random non-relevant document.
+			\item<4-> Estimating this parameter is empirically non-trivial!
+		\end{itemize}
+
+	\end{columns}
+
+\end{frame}
+
+\begin{frame}{Realistic empirical evaluations for AI}
+
+\begin{remark}
+	To know when it is safe to use AI systems in future evidence synthesis projects, we need to evaluate them on past data \textbf{under realistic conditions}
+\end{remark}
+
+In living evidence applications:
+
+\begin{itemize}
+	\item<2-> How long can we trust classifier evaluations without labelling new data?
+	\item<3-> How does topic model fit decay over time?
+	\item<4-> How do we incorporate new topics?
+	\item<4-> How frequently \textit{would} LRs have been updated? Is this predictable?
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%
+% Conclusion
+\begin{frame}{Conclusion}
+
+We provide a stopping criteria that works on any model, with any tool: \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteriaR.md}.
+
+\medskip
+
+Work savings in practice with large datasets are large!
+
+\medskip
+
+Future work will identify how biased our urn is, in order to use a noncentral hypergeometric distribution, which should give a more precise, less conservative criterion.
+
+\begin{center}
+	\line(1,0){250}
+
+	\medskip
+
+	\textbf{Thanks!}
+
+	\medskip
+
+	Contact: \url{mueller-hansen@mcc-berlin.net, callaghan@mcc-berlin.net}
+
+\end{center}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%
+
+\appendix
+
+\begin{frame}{References}
+
+\bibliography{rapid-review}
 

### 2023-03-14 pres/WWGS.tex
update WWGS presentation

@@
 \date{20 October 2022}
-\author{Finn Müller-Hansen, Max Callaghan}
+\author{Max Callaghan, \underline{Finn Müller-Hansen}}
 \institute[MCC]{
@@
 % RITL
-\begin{frame}{Researcher in the Loop process}
+\begin{frame}{ML-assissted document screening}
 
@@
 		\begin{itemize}
-			\item<1-> A large literature \citep{omara-eves_using_2015} has developed ``human-in the loop'' machine learning applications which ``overcome the manual and time-consuming screening of large numbers of studies by prioritizing relevant studies via active learning''  \cite{van_de_schoot_open_2021}
-			\item<2-> Identifying 95\% of relevant documents if screening manually would mean screening 95\% of all documents
-			\item<3-> If we can use machine learning to prioritise documents likely to be relevant, we can achieve high levels of recall with low levels of effort.
-			\item<4-> BUT, given we do not know \textit{a priori} the true number of relevant documents, we need strategies to stop screening and bank the work savings
-			\item<5->Getting these wrong can mean missing our target
+			\item<1-> Growing number of ``researcher-in-the-loop'' machine learning applications for screening documents for systematic reviews \citep{omara-eves_using_2015, van_de_schoot_open_2021}.
+			\item<2-> By using machine learning to prioritise documents likely to be relevant, we can achieve high levels of recall without screening all documents.
+			\item<3-> BUT, given we do not know \textit{a priori} the true number of relevant documents, we need criteria when to stop screening.
+			\item<4->Stopping too early can lead to huge biases in reviews.
 		\end{itemize}
@@
 		\begin{figure}
-			\only<2-3>{\includegraphics[width=\linewidth]{images/idealised_stopping.pdf}}
-			\only<4>{\includegraphics[width=\linewidth]{images/potential_stopping.pdf}}
-			\only<5>{\includegraphics[width=\linewidth]{images/early_stopping.pdf}}
+			\only<2>{\includegraphics[width=\linewidth]{images/idealised_stopping.pdf}}
+			\only<3>{\includegraphics[width=\linewidth]{images/potential_stopping.pdf}}
+			\only<4>{\includegraphics[width=\linewidth]{images/early_stopping.pdf}}
 %			\only<2>{\includegraphics[width=\linewidth]{images/random.pdf}}
@@
 % Urn
-\begin{frame}{A stopping criterion}
+\begin{frame}{A statistical stopping criterion}
 
@@
 		\begin{itemize}
-			\item<1-> Our stopping criterin works by treating documents as if they were white (not relevant) and red (relevant) marbles drawn from an urn without replacement.
-			\item<2-> The hypergeometric distribution tells us the probability of observing \textbf{k} red marbles in a sample of \textbf{n} marbles, given an urn with \textbf{N} marbles, of which \textbf{K} are relevant
-			\item<3-> We reformulate this to generate a null hypothesis $H_0$ that a given recall target has been \textbf{missed}. 
-			\item<4-> We calculate a p-score for our null hypothesis, and if this is low enough, we reject $H_0$ and stop screening.
+			\item<1-> Our stopping criterion works by treating documents as if they were white (not relevant) and red (relevant) marbles drawn from an urn without replacement.
+			\item<2-> The hypergeometric distribution describes the probability of observing \textbf{k} red marbles in a sample of \textbf{n} marbles, given an urn with \textbf{N} marbles, of which \textbf{K} are red.
+			\item<3-> We formulate a null hypothesis $H_0$ that a given recall target (e.g. 95\% of relevant documents) has been \textbf{missed}. 
+			\item<4-> We calculate a p-score for $H_0$ and stop screening if this falls below a selected threshold.
 		\end{itemize}
@@
 
-\only<5->{Note, ML-prioritisation means documents are not drawn at random, which makes our test conservative as long as ML works as well as or better than random chance.}
+\only<5->{Note: ML-prioritisation means documents are not drawn at random, which makes our test conservative.}
 
@@
 
-We test our criteria against other commonly used criteria on 20 complete systematic review datasets
+We run simulations on 20 complete systematic review datasets to test our criterion.
 
@@
 		\begin{itemize}
-			\item<1-> \textit{Potential} work savings (if we already knew when to stop) varied widely (higher for larger datasets - blue dots)
-			\item<2-> Existing stopping criteria can result in catastrophic errors
-			\item<4-> Our criteria generated work savings with reliably conservative performance wrt our recall target.
+			\item<1-> \textit{Theoretically achievable} work savings in the datasets varies widely (higher for larger datasets - blue dots)
+			\item<2-> Existing stopping criteria can result in very low recall (examples: 50 consecutive irrelevant articles, using baseline inclusion rate)
+			\item<4-> Our criterion generated work savings with reliably conservative performance wrt our recall target.
 		\end{itemize}
@@
 
-We provide a stopping criteria that works on any model, with any tool: \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteriaR.md}.
+We provide a stopping criterion that works on any model and can be included in any tool: \url{https://github.com/mcallaghan/rapid-screening/blob/master/analysis/hyper_criteriaR.md}.
 
@@
 
-Work savings in practice with large datasets are large!
+In practice, we see huge work savings for large datasets!
 
@@
 
-Future work will identify how biased our urn is, in order to use a noncentral hypergeometric distribution, which should give a more precise, less conservative criterion.
+Future work: use noncentral hypergeometric distribution to generate a less conservative test and save more work.
 
@@
 
-	Contact: \url{mueller-hansen@mcc-berlin.net, callaghan@mcc-berlin.net}
+	Contact: \url{callaghan@mcc-berlin.net, mueller-hansen@mcc-berlin.net}
 
@@
 \begin{itemize}
-	\item<1-> We have used the stopping criteria to generate massive savings (77\%) in real projects
+	\item<1-> We have used the stopping criterion to generate massive savings (77\%) in real projects
 	\item<2-> If rejecting our $H_0$ was less labour intensive we could have saved around 82\%
